---
title: "iDeer Tool Function"
format: html
editor: visual

#User selects woodland
#Metrics are calculated for the focal woodland and all woodlands in surrounding 10km landscape
#Cut all rasters to 3km, buffer by 3km for spatial processes (5km total)
#Extract the metrics (roads, urban, connectivity etc)
#Run the BBN
#Loop this process for the 5 different landscapes
#Do this for the 5 DIA regions
#Write as a function
---

```{r}
library(plyr)
library(ggplot2)
library(dplyr)
library(sf)
library(raster)
library(terra)
library(here)
library(progress)
library(bnlearn)
library(RColorBrewer)
library(rasterVis)
library(ggmap)

here()
```

### Load spatial data derived for England and Wales

```{r}
bng <- 27700

#Load NFI/LCM2015 master map
map <- raster(here("edge_core_raster_all_tiles_EW.tif"))
crs(map) <- bng

#Load nearest road
#Nearest A road, B road or motorway
nearest_road <- raster(here("nearest_road_raster_2015_all_tiles_corrected.tif"))
crs(nearest_road) <- bng

#Load nearest urban
nearest_urban <- raster(here("nearest_urban_raster_2015_all_tiles_corrected.tif"))
crs(nearest_urban) <- bng
nearest_suburb <- raster(here("nearest_suburban_raster_2015_all_tiles_corrected.tif"))
crs(nearest_suburb) <- bng

#Load linear features at 1km scale
lf <- raster(here("lf_1000_mosaic_all_tiles_corrected.tif"))
crs(lf) <- bng

#Load MAX DAMS calculated at 1km scale
dams <- raster(here("focal.dams.1000.max.non_wood_EW.tif"))

#Read in woodland polygon dataset
#Combined NFI 2015 and LCM 2015 woodlands
#These polygons will be used to calculate the connectivity, they are unified
#All touching polygons have become unified
hab_patches_all <- st_read(here("nfi_lcm_overlaid_shapefile_woods_export.shp"))
st_crs(hab_patches_all) <- bng

head(hab_patches_all)

hab_patches_all <- hab_patches_all %>%
  rename(patch_ID = Id) %>%
  rename(patch_area = Shape_Area)

lcm_polys <- st_read(here("data/derived-data/nfi_lcm_2015_overlaid.shp"))
#Select for woodland polys only
wood_polys <- lcm_polys %>% filter(gridcode %in% c("1","2","22","23")) %>%
  rename(patch_ID = Id,
         class = gridcode)


```

### Choose woodlands

NEED TO CHANGE THIS TO BE ABLE TO DRAW A POLYGON INSIDE WOODLANDS SO THAT USER CAN SELECT THEIR LAND PARCEL

```{r}

#Choose five woodlands within the DIA priority areas to extract data for
#This is simulating the user selecting a woodland of interest

#Suffolk
#suffolk_wood <- hab_patches_all %>% subset(patch_ID == 260977)
#Cumbria
#cumbria_wood <- hab_patches_all %>% subset(patch_ID == 142765)
#South Wales 
#sw_wood <- hab_patches_all %>% subset(patch_ID == 340972)
#North Wales
#nw_wood <- hab_patches_all %>% subset(patch_ID == 188539)
#Oxford 
#ox_wood <- hab_patches_all %>% subset(patch_ID == 370647)

#select all woodlands together
#woods <- wood_polys %>%
#  filter(patch_ID %in% c("213709", "377707", "455598", "188539", "858585"))
#woods$Region <- c("North West","Cumbria","North Wales","Suffolk","Oxford")
#woods <- st_as_sf(woods)

#Just do one wood for now

woods <- wood_polys %>% filter(patch_ID %in% c("455598"))
woods$Region <- c("Suffolk")
woods <- st_as_sf(woods)



```

### Buffer woodlands for spatial data extraction

```{r}
#Buffer all woods by 3km
buffered_woods_3km <- st_as_sf(st_buffer(woods, dist = 3000))  # dist = 3km in meters
#Buffer all woods by 5km (3km landscape + 3km edge buffer)
buffered_woods_5km <- st_as_sf(st_buffer(woods, dist = 5000))  # dist = 5km in meters

#ensure crs of buffers is BNG
st_crs(buffered_woods_5km) <- bng
st_crs(buffered_woods_3km) <- bng

region_colors <- c("Cumbria" = "red","North Wales" = "blue",
                   "Suffolk"="magenta","North West"="yellow","Oxford"="orange")
buffered_woods_3km$color <- region_colors[buffered_woods_3km$Region]

# Plot the first geometry with no color
plot(st_geometry(buffered_woods_5km), col = NA, main = "Buffered Woods")

# Plot the second geometry with colors based on regions
plot(st_geometry(buffered_woods_3km), col = buffered_woods_3km$color, add = TRUE)

#Check points, woodlands buffers line up
plot(st_geometry(buffered_woods_5km), col = NA)
plot(st_geometry(buffered_woods_3km), col = buffered_woods_3km$color, add = TRUE)
plot(st_geometry(woods), col = "black", add = TRUE)
#plot(st_geometry(wood_points), add = TRUE, col = "red")

#Filter for woodland patches that intersect buffers
woods_buffers <- st_filter(wood_polys, buffered_woods_3km, .predicate = st_intersects)

```

### Extract raster variables for all woodlands in 10km buffer

Produce dataframes with pixel ID's from land cover map. This will allow calculation of pixel-level deer damage risk. Need to ensure there is a "patch_ID" column so that the connectivity values can be assigned to pixels within each patch (calculated at patch-level).

```{r}
#NEAREST URBAN EXTRACTION ####

extract_urban_rasters <- function(buffered_woods_3km, raster_layer, land_cover_raster) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_3km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_3km$geometry)) {
  buffer_geometry <- buffered_woods_3km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
 # Get the cell numbers within the extent
pixel_ID <- cellsFromExtent(land_cover_raster, buffer_extent)

# Get the coordinates for these cells
cell_coords <- as.data.frame(xyFromCell(land_cover_raster, pixel_ID))

#Extract pixel values
pixel_values <- extract(raster_layer, cell_coords)

    # Create a data frame with the extracted values and cell indices
    vals <- data.frame(nearest_urban = pixel_values, 
                       patch_ID = buffered_woods_3km$patch_ID[i],
                       pixel_ID = pixel_ID,
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    vals <- vals[!is.na(vals$nearest_urban), ]
    
    # Append to the results list
    extraction_results[[i]] <- vals
  }
  
  # Combine all results into a single dataframe
  result_df <- bind_rows(extraction_results)
  
  return(result_df)
}

nearest_urban_extraction <- extract_urban_rasters(buffered_woods_3km=buffered_woods_3km, 
raster_layer=nearest_urban,
land_cover_raster=map)

write.csv(nearest_urban_extraction, here("nearest_urban_extraction_pixelIDs_3km.csv"))
```

```{r}
#NEAREST SUBURBAN EXTRACTION ####
extract_suburban_rasters <- function(buffered_woods_3km, raster_layer, land_cover_raster) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_3km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_3km$geometry)) {
  buffer_geometry <- buffered_woods_3km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
 # Get the cell numbers within the extent
pixel_ID <- cellsFromExtent(land_cover_raster, buffer_extent)

# Get the coordinates for these cells
cell_coords <- as.data.frame(xyFromCell(land_cover_raster, pixel_ID))

#Extract pixel values
pixel_values <- extract(raster_layer, cell_coords)

    # Create a data frame with the extracted values and cell indices
    vals <- data.frame(nearest_suburban = pixel_values, 
                       patch_ID = buffered_woods_3km$patch_ID[i],
                       pixel_ID = pixel_ID,
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    vals <- vals[!is.na(vals$nearest_suburban), ]
    
    # Append to the results list
    extraction_results[[i]] <- vals
  }
  
  # Combine all results into a single dataframe
  result_df <- bind_rows(extraction_results)
  
  return(result_df)
}

nearest_suburban_extraction <- extract_suburban_rasters(buffered_woods_3km=buffered_woods_3km, 
raster_layer=nearest_suburb,
land_cover_raster=map)

write.csv(nearest_suburban_extraction, here("nearest_suburban_extraction_pixelIDs_3km.csv"))
```

```{r}
#MAX DAMS EXTRACTION AT PIXEL LEVEL ####

extract_dams_raster <- function(buffered_woods_3km, raster_layer, land_cover_raster) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_3km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  # List to store extracted data
  extraction_results <- list()
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_3km$geometry)) {
  buffer_geometry <- buffered_woods_3km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
 # Get the cell numbers within the extent
pixel_ID <- cellsFromExtent(land_cover_raster, buffer_extent)

# Get the coordinates for these cells
cell_coords <- as.data.frame(xyFromCell(land_cover_raster, pixel_ID))

#Extract pixel values
pixel_values <- extract(raster_layer, cell_coords)

    # Create a data frame with the extracted values and cell indices
    vals <- data.frame(max_dams = pixel_values, 
                       patch_ID = buffered_woods_3km$patch_ID[i],
                       pixel_ID = pixel_ID,
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    vals <- vals[!is.na(vals$max_dams), ]
    
    # Append to the results list
    extraction_results[[i]] <- vals
  }
  
  # Combine all results into a single dataframe
  result_df <- bind_rows(extraction_results)
  
  return(result_df)
}

dams_extraction <- extract_dams_raster(
buffered_woods_3km=buffered_woods_3km, 
raster_layer=dams,
land_cover_raster=map)


write.csv(dams_extraction, here("dams_extraction_pixelIDs_1000_3km.csv"))

#check pixels match up with other layers
dams_pixels <- dams_extraction$pixel_ID
urban_pixels <- nearest_urban_extraction$pixel_ID
matching_pixels <- intersect(dams_pixels, urban_pixels)
```

```{r}
#LINEAR FEATURES EXTRACTION AT PIXEL LEVEL ####
extract_lf_raster <- function(buffered_woods_3km, raster_layer, land_cover_raster) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_3km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_3km$geometry)) {
  buffer_geometry <- buffered_woods_3km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
   # Get the cell numbers within the extent
pixel_ID <- cellsFromExtent(land_cover_raster, buffer_extent)

# Get the coordinates for these cells
cell_coords <- as.data.frame(xyFromCell(land_cover_raster, pixel_ID))

#Extract pixel values
pixel_values <- extract(raster_layer, cell_coords)

    # Create a data frame with the extracted values and cell indices
    vals <- data.frame(lf = pixel_values, 
                       patch_ID = buffered_woods_3km$patch_ID[i],
                       pixel_ID = pixel_ID,
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    vals <- vals[!is.na(vals$lf), ]
    
    # Append to the results list
    extraction_results[[i]] <- vals
  }
  
  # Combine all results into a single dataframe
  result_df <- bind_rows(extraction_results)
  
  return(result_df)
}

lf_extraction <- extract_lf_raster(buffered_woods_3km=buffered_woods_3km, 
raster_layer=lf,
land_cover_raster=map)

write.csv(lf_extraction, here("linear_feature_extraction_pixelIDs_1000_3km.csv"))
```

```{r}
#NEAREST ROAD EXTRACTION AT PIXEL LEVEL ####

nearest_road_rasters <- function(buffered_woods_3km, raster_layer, land_cover_raster) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_3km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_3km$geometry)) {
  buffer_geometry <- buffered_woods_3km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
 # Get the cell numbers within the extent
pixel_ID <- cellsFromExtent(land_cover_raster, buffer_extent)

# Get the coordinates for these cells
cell_coords <- as.data.frame(xyFromCell(land_cover_raster, pixel_ID))

#Extract pixel values
pixel_values <- extract(raster_layer, cell_coords)

    # Create a data frame with the extracted values and cell indices
    vals <- data.frame(nearest_road = pixel_values, 
                       patch_ID = buffered_woods_3km$patch_ID[i],
                       pixel_ID = pixel_ID,
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    vals <- vals[!is.na(vals$nearest_road), ]
    
    # Append to the results list
    extraction_results[[i]] <- vals
  }
  
  # Combine all results into a single dataframe
  result_df <- bind_rows(extraction_results)
  
  return(result_df)
}

nearest_road_extraction <- nearest_road_rasters(buffered_woods_3km=buffered_woods_3km, 
raster_layer=nearest_road,
land_cover_raster=map)

write.csv(nearest_suburban_extraction, here("nearest_road_extraction_pixelIDs_3km.csv"))
```

```{r}
#LAND COVER EXTRACTION FOR 1km AROUND WOODLANDS ####
#Function written by Becks
#This function takes too much time to perform at the pixel scale
#Therefore we are calculating Nutritional Landscape Quality at the patch level.
#2mins for 1 patch in 3km landscape

  polyFragStats <- function(POLY, veg) {
  # Create a buffer around the polygon
  buffer_poly <- buffer(POLY, width = 1000, dissolve = TRUE)
  
  # Cut out the original polygon from the buffer
  buffer_area <- erase(buffer_poly, POLY)
  
  # Crop land cover map by the buffer area
  fcrop <- crop(veg, buffer_area)
  
  # Mask the cropped land cover map
  fmask <- mask(fcrop, buffer_area)
  
  # Extract class statistics
  res <- raster::extract(fmask, buffer_area, method = "simple")
  
  # Convert the results to a data frame and calculate statistics
  res <- as.data.frame(table(res))
  res <- res %>% rename(class=res)
  res$class <- as.numeric(as.character(res$class))
  res$class <- round(res$class, 3)
  res$patch_ID <- POLY$patch_ID
  
  res <- res %>%
    mutate(total_pix = as.numeric(sum(Freq)))%>%
    dplyr::select(patch_ID, class, Freq, total_pix) %>%
    mutate(perc_cov = Freq * 100 / total_pix, buff_area = area(buffer_area))
  
  return(res)
}

#Filter for woodland patches that intersect buffers
woods_buffers <- st_filter(wood_polys, buffered_woods_3km, .predicate = st_intersects)

POLY=as(woods_buffers, "Spatial")
veg = map

# Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = length(POLY),
  clear = FALSE,
  width = 60
)

respoly <- ldply(1:length(POLY), function(i) {  # for each polygon
  pb$tick()  # Update the progress bar
  out <- polyFragStats(POLY[i,],  veg=veg) # compute fragstats metrics 
  out$buffer_size <- "b1000"
  return(out)
})  

respoly

write.csv(respoly, here("landscapemetrics_1000_3km.csv"))

```

```{r}
#WOODLAND PIXEL EXTRACTION ####
#Do this at the pixel level

woodpixels <- function(POLY, veg, i) {
  # Crop land cover map by the woodland polygon
  fcrop <- crop(veg, extent(POLY[i, ]))
  
  # Mask the cropped land cover map
  fmask <- mask(fcrop, POLY[i, ])
  
  # Get extent of fmask
  fmask_extent <- extent(fmask)
  
  # Get the cell numbers within the extent
  pixel_ID <- cellsFromExtent(veg, fmask_extent)
  
  # Get the coordinates for these cells
  cell_coords <- as.data.frame(xyFromCell(veg, pixel_ID))
  
  # Extract pixel values
  pixel_values <- extract(fmask, cell_coords)
  
  # Ensure that there are pixel values before proceeding
  if (length(pixel_values) > 0) {
    # Create a data frame with the extracted values and cell indices
    vals <- data.frame(pixel_ID = pixel_ID,
                       lc_class = pixel_values,
                       patch_ID = POLY@data$patch_ID[i],
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    vals <- vals[!is.na(vals$lc_class), ]
  } else {
    # If no values, create an empty data frame with the same structure
    vals <- data.frame(pixel_ID = integer(0),
                       lc_class = integer(0),
                       patch_ID = integer(0),
                       x = numeric(0),
                       y = numeric(0))
  }
  
  return(vals)
}

# Filter for woodland patches that intersect buffers
woods_buffers <- st_filter(wood_polys, buffered_woods_3km, .predicate = st_intersects)

# Convert to spatial object
POLY <- as(woods_buffers, "Spatial")
veg <- map

# Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = length(POLY),
  clear = FALSE,
  width = 60
)

# Process each polygon and collect results
woodpix <- ldply(1:length(POLY), function(i) {  # for each polygon
  pb$tick()  # Update the progress bar
  out <- woodpixels(POLY, veg = veg, i = i)  # compute metrics
  return(out)
})

write.csv(woodpix, here("woodland_type_pixelIDs_3km.csv"))
```

### Calculate connectivity for every woodland inside the 3km buffers

```{r}
#### SET CONNECTIVITY PARAMETERS ####
# connectivity parametetrs - % of dispersers reaching a set distance #
percentage_dispersers <- 0.05 ### 5% - 95% of individuals will go to woodlands within dispersal_distance
dispersal_distance <- 400 
dispersal_contribution <-
  -((log(1 / percentage_dispersers)) / dispersal_distance)

# Buffer distance represents the cut off - this stops the script measuring every pairwise combination
dispersal_cutoff <- 0.999 ### 99.9% cut off
#buffer_cutoff <-
  #round(log(1 / (1 - dispersal_cutoff)) / (log(1 / percentage_dispersers) /
                                             #dispersal_distance), digits = 2)
buffer_cutoff = 1000 #1km

# test & plot of dispersal contribution for sequence of distance up to buffer_cutoff
test <-
  data.frame(distance = (seq(0, buffer_cutoff, by = 0.5))) %>%
  mutate(contribution = (exp(dispersal_contribution * distance) * 100))

ggplot(data = test, aes(x = distance, y = contribution, group = 1)) +
  geom_line() +
  geom_point() +
  geom_vline(xintercept = buffer_cutoff, color = "red")

#See how buffer size looks
focal_patch_buffer <- st_buffer(woods[1,], buffer_cutoff)
plot(st_geometry(focal_patch_buffer))
plot(st_geometry(woods[1,]),add=TRUE)

```

```{r}

#Connectivity function ####

connectivity_results <- list()

  # Use the 5km buffer as the source patches area
  source_woods <- st_filter(wood_polys, buffered_woods_5km, .predicate = st_intersects)

  # Crop the woodlands in the original 3km square to get the focal patches
  focal_woods <- st_filter(wood_polys, buffered_woods_3km, .predicate = st_intersects)
  n_distinct(focal_woods$patch_ID)

  # Calculate connectivity for each focal woodland
  Connectivity_table <- NULL
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(focal_woods),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)

  # Loop through each focal woodland to calculate connectivity
  for (j in 1:nrow(focal_woods)) {
    
    # Focal patch in the 10km square
    focal_patch <- focal_woods[j, ]
    
    #update progress bar
    pb$tick
    
    # Buffer 1000m around the focal patch
    focal_patch_buffer <- st_buffer(focal_patch, buffer_cutoff)
    
    # Find source patches within this buffer
    source_patches <- st_filter(source_woods, focal_patch_buffer, .predicate = st_intersects)
    
    # Calculate distances between the focal and source patches
    patch_dist <- as.vector(st_distance(focal_patch, source_patches))
    
    # Create a data frame for connectivity information
    Conn_table_site <- data.frame(
      focal_patch = focal_patch$patch_ID,
      focal_patch_area = as.numeric(focal_patch$Shape_Area),
      source_patch = source_patches$patch_ID,
      source_patch_area = as.numeric(source_patches$Shape_Area),
      distance = patch_dist,
      incoming_connect = ifelse(
        focal_patch$patch_ID == source_patches$patch_ID,
        NA,
        source_patches$Shape_Area * exp(dispersal_contribution * patch_dist)
      )
    )
    
    # Append to 'Connectivity_table'
    if (is.null(Connectivity_table)) {
      Connectivity_table <- Conn_table_site
    } else {
      Connectivity_table <- rbind(Connectivity_table, Conn_table_site)
    }
  
  n_distinct(Connectivity_table$focal_patch)

  # Store results for this woodland
  connectivity_results[[paste("focal_wood", focal_woods$patch_ID[j])]] <- Connectivity_table
}
```

```{r}

#Remove rows where focal_patch = source_patch
Connectivity_table_filt <- Connectivity_table %>%
  filter(focal_patch != source_patch)

# Identify missing unique patch IDs
# These patches have no source patches so will have been filtered out
# We need to put them back in so we can give them an incoming_connect value of 0
missing_patch_ids <- setdiff(
     unique(Connectivity_table$focal_patch),
     unique(Connectivity_table_filt$focal_patch)
 )
 
 print(missing_patch_ids)  # This will show which patch IDs are missing

missing_patches <- Connectivity_table %>% filter(focal_patch %in% missing_patch_ids) %>%
  # Replace NAs in 'incoming_connect' with 0
  mutate(incoming_connect = coalesce(incoming_connect, 0))  # Replace NA with 0

#Make rows for patches with no connectivity so we don't lose any patches
Connectivity_table_filt <- rbind(Connectivity_table_filt, missing_patches)

#Ensure no duplicate rows
Connectivity_table_filt <- Connectivity_table_filt %>% distinct()

#Now no patches should be missing
n_distinct(Connectivity_table_filt$focal_patch)

#sum incoming connectivity by focal patch
incoming_connectivity_sum <- Connectivity_table_filt %>%
  dplyr::select(focal_patch, incoming_connect) %>%
  group_by(focal_patch) %>%
  dplyr::summarise(total_connect = sum(incoming_connect),
            n = n()) %>%
  mutate(n = if_else(total_connect == 0, 0, n)) %>%  # Set 'n' to 0 when 'incoming_connect' is 0
  rename(patch_ID = focal_patch)

#save incoming connectivity table

write.csv(incoming_connectivity_sum, here("incoming_connectivity_sum_3km.csv"))
```

### Ready the spatial data for the Bayesian Belief Network

#### Calculate Nutritional Landscape Composition (at 1km scale)

```{r}
#CALCULATE NUTRITIONAL LANDSCAPE QUALITY ####

respoly<-read.csv(here("landscapemetrics_1000_3km.csv"))
lcm <- respoly
#lcm <- lcm%>%rename(class = lc_class)
#Load landscape nutritional scores for raster categories
#Edge forage quality scores sum of woodland score + adjacent land cover score
#Water scores retained to prevent loss of woodland edge pixels
#Water edge pixels receive one point higher than regular woodland
#Still an edge so will be highly nutritious
edge.val.types <- read.csv(file.path(here("Expert_derived_quality_ranks_edge_types_with_water.csv")))
edge.val.types <- edge.val.types%>%dplyr::select(-"X")

#Stitch the quality scores into the lcm dataset

lc.edge.types <- lcm %>% left_join(edge.val.types, by = "class") %>% mutate(Forage_Q = ifelse(is.na(Forage_Q), 0, Forage_Q))%>%
  rename(scale = buffer_size)  %>%
 dplyr::select(-"X") %>% filter(!is.na(Land.cover))

#Calculate Nutritional Landscape Index (NLI) for separate edges and lc ####
#Weight edges by their type: https://www.sciencedirect.com/science/article/pii/S0304380009003871 #fig1  

# Calculate the weighted score for each pixel type
lc.edge.types <- lc.edge.types %>%
  mutate(weighted_score = perc_cov * Forage_Q)

# Sum the weighted scores
weighted_sum_by_section_buffer <- lc.edge.types %>%
  group_by(patch_ID) %>%
  summarise(total_weighted_score = sum(weighted_score),
            total_perc_cov = sum(perc_cov))

# Calculate the weighted average score for each woodland polygon
weighted_average_score_buffer <- weighted_sum_by_section_buffer %>%
  mutate(weighted_avg_score = total_weighted_score / total_perc_cov)

weighted_average_score_buffer <- weighted_average_score_buffer %>%
  group_by(patch_ID) %>%
  summarise(
    nli_edge_types_1000 = sum(weighted_avg_score, na.rm = TRUE)
  )

# Join the calculated values back to the original dataset:
#woods_buffers <- st_filter(wood_polys, buffered_woods_3km, .predicate = st_intersects)
#lc.edge.types.NLI <- left_join(woods_buffers, #weighted_average_score_buffer, by = c("patch_ID"))

```

#### Assign woodland pixel quality values

```{r}

woodpix <- read.csv(here("woodland_type_pixelIDs_3km.csv"))
woods <- woodpix
woods <- woods%>%rename(class = lc_class)
unique(woods$class)

#Load landscape nutritional scores for raster categories
#Edge forage quality scores sum of woodland score + adjacent land cover score
#Water scores added in to prevent loss of woodland edge pixels
#Water edge pixels receive one point higher than regular woodland
#Still an edge so will be highly nutritious
edge.val.types <- read.csv(file.path(here("Expert_derived_quality_ranks_edge_types_with_water.csv")))
edge.val.types <- edge.val.types%>%dplyr::select(-"X")

#Check edge.val.types$class and woods$class match in format
unique(edge.val.types$class)
class(edge.val.types$class)
edge.val.types <- edge.val.types%>%arrange(class)
unique(edge.val.types$class)

class(woods$class)
woods <- woods%>%arrange(class)
unique(woods$class)
pcm <-woods%>%arrange(class)
unique(pcm$class)

# Round to 3 decimal places
woods$class <- round(woods$class, 3)
edge.val.types$class <- round(edge.val.types$class, 3)

#Stitch the quality scores into the lcm dataset

wood.edge.types <- woods %>% left_join(edge.val.types, by = c("class")) #%>%
                                          #filter(!is.na(Land.cover))

#Replace "NA" in Forage_Q column with 0
wood.edge.types <- wood.edge.types%>% 
 mutate(Forage_Q = ifelse(is.na(Forage_Q), 0, Forage_Q))%>%
  mutate(scale = "patch")%>%
   dplyr::select(-"X")


```

#### left_join the data sets together

```{r}

#Forage quality of pixels and Nutritional Landscape Index for patch
df <- left_join(weighted_average_score_buffer, wood.edge.types, by = "patch_ID")

#Connectivity ####
incoming_connectivity_sum<-read.csv(here("data/derived-data/Tool-extract-function-test/incoming_connectivity_sum_3km.csv"))
incoming_connectivity_sum<-incoming_connectivity_sum %>%
     dplyr::select(-c("X","n"))
df <- left_join(df, incoming_connectivity_sum, by=c("patch_ID")) 

#MAX DAMS ####
dams_scores <- dams_extraction%>%rename(max.dams.1000=max_dams)%>%
  dplyr::select(-c("patch_ID"))%>%
  st_drop_geometry()

df <- left_join(df, dams_scores, by=c("pixel_ID","x","y"))

# Plot raster using the color palette with custom labels
ggplot(df) +
  geom_raster(aes(x = x, y = y, fill = max.dams.1000)) +
  theme_minimal() +
  labs(title = "Predicted Damage Raster",
       fill = "Damage Level") +
  guides(fill = guide_legend(override.aes = list(alpha = 1))) # Ensure all levels are shown in the legend

#df <- left_join(df, dams_scores, by=c("patch_ID")) 

#Road density
#roads<-road_extraction%>%dplyr::select(-c("patch_ID"))
#df <- left_join(df, roads, by=c("pixel_ID","x","y"))
#df <- df %>% rename(all_roads_250=roads)

#Nearest road
nearest_road_data <- nearest_road_extraction%>%dplyr::select(-c("patch_ID"))
df <- left_join(df, nearest_road_data, by=c("pixel_ID","x","y"))

#Nearest urban and suburban ####
nearest_suburban_extraction<- nearest_suburban_extraction%>%dplyr::select(-c("patch_ID"))
nearest_urban_extraction<- nearest_urban_extraction%>%dplyr::select(-c("patch_ID"))

df <- left_join(df, nearest_suburban_extraction, by=c("pixel_ID","x","y"))

df <- left_join(df, nearest_urban_extraction, by=c("pixel_ID","x","y"))

# Plot raster using the color palette with custom labels
ggplot(df) +
  geom_raster(aes(x = x, y = y, fill = nearest_road)) +
  theme_minimal() +
  guides(fill = guide_legend(override.aes = list(alpha = 1))) # Ensure all levels are shown in the legend

#Make column for nearest urban and suburban using minimum value
df$urban_proximity <- pmin(df$nearest_suburban, df$nearest_urban)
#Drop individual columns
df <- df%>%dplyr::select(-c("nearest_suburban","nearest_urban"))

#Linear features####
lf_extraction <- lf_extraction%>%dplyr::select(-c("patch_ID"))
df <- left_join(df, lf_extraction,  by=c("pixel_ID","x","y"))

#Remove the NAs

df <- na.omit(df)

saveRDS(df, here("bbn_test_df_5_patches_pixel_level_3km.rds"))
```

#### Convert numerical values to categorical (LOW, MED, HIGH)

```{r}

#Replace values in numerical columns with LOW, MED, HIGH
#This is based on logic from google spreadsheet "BN thresholds" 

df <- readRDS(here("bbn_test_df_5_patches_pixel_level_3km.rds"))

#Look at distribution of values across national datasets ####
#We need to decide at what point values change between categories- need to do this using whole datasets rather than localised cropped datasets.

#Roads
#hist(all_roads$all_roads_250_mosaic_all_tiles_corrected)
#df$all_roads_250 <- ifelse(is.na(df$all_roads_250), NA,
#                           ifelse(df$all_roads_250 <= 0, "LOW",
#                           ifelse(df$all_roads_250 > 0 & df$all_roads_250 <= 100, "MED",
#                           ifelse(df$all_roads_250 > 100, "HIGH", NA))))
#unique(df$all_roads_250)

#Nearest roads
df$nearest_road <- ifelse(is.na(df$nearest_road), NA,
                           ifelse(df$nearest_road <= 100, "LOW",
                           ifelse(df$nearest_road > 100 & df$nearest_road <= 500, "MED",
                           ifelse(df$nearest_road > 500, "HIGH", NA))))

#Linear features
#hist(lf$lf_1000_mosaic_all_tiles_corrected)
# Replace numeric values with categorical labels
df$lf <- ifelse(df$lf<= 2500, "LOW",
            ifelse(df$lf> 2500 & df$lf <= 5000, "MED",
            ifelse(df$lf> 5000, "HIGH", NA)))
unique(df$lf)


#nearest urban/suburban
#Proximity, therefore "HIGH" means lower value (lower distance)
#These thresholds are based on deer biology rather than the spead of data
#hist(nearest_urban$nearest_urban_raster_2015_all_tiles_corrected)
#hist(nearest_suburb$nearest_suburban_raster_2015_all_tiles_corrected)
df$urban_proximity <- ifelse(df$urban_proximity <= 500, "HIGH",
                          ifelse(df$urban_proximity >500 & df$urban_proximity <= 1000, "MED",
                           ifelse(df$urban_proximity >1000,"LOW",NA)))  
unique(df$urban_proximity)

#MAX DAMS
#hist(dams$max.dams.1000)
df$max.dams.1000 <- ifelse(df$max.dams.1000 <=15,"LOW",
                      ifelse(df$max.dams.1000>15 & df$max.dams.1000 <= 20, "MED",
                        ifelse(df$max.dams.1000 >20, "HIGH",NA)))

#nutritional landscape composition score - do we need to change these thresholds????
hist(edge.val.types$Forage_Q)
hist(df$nli_edge_types_1000)
df$nli_edge_types_1000 <- ifelse(df$nli_edge_types_1000 <= 4, "LOW",
                          ifelse(df$nli_edge_types_1000 >5 & df$nli_edge_types_1000 <=8, "MED",
                           ifelse(df$nli_edge_types_1000 > 8, "HIGH",NA)))

#nutritional patch composition score

hist(edge.val.types$Forage_Q)
hist(df$Forage_Q)
df$Forage_Q <- ifelse(df$Forage_Q <= 5, "LOW",
                          ifelse(df$Forage_Q >5 & df$Forage_Q <=10, "MED",
                           ifelse(df$Forage_Q>10, "HIGH",NA)))

#connectivity
#hist(df$total_connect)
#quantile(df$total_connect)

#log transform connectivity values, they are too big for easy interpretation

df$total_connect_log <- log(df$total_connect)

#hist(df$total_connect_log)
quantile(df$total_connect_log)

df$total_connect_log <- ifelse(df$total_connect_log <=10, "LOW",
                          ifelse(df$total_connect_log >10 & df$total_connect_log  <=15, "MED",
                           ifelse(df$total_connect_log  > 15, "HIGH",NA)))

#remove original connectivity column

df <- df%>%dplyr::select(-c("total_connect"))

#Export

df_categorical <- df

#Remove NAs from pixel-level measured variables
#NEED TO FIX THIS?? Why am I getting NAs?

df_categorical <- na.omit(df_categorical)

saveRDS(df_categorical, file = file.path(here("tool_test_extracted_data_categorical_pixel_level_3km.rds")))

```

### Create the Bayesian Belief Network

#### Initialize Conditional Probability Tables

```{r}
#CPTS for measured nodes ####
urban_prox_cpt<-matrix(c(0.333333,0.333333,0.333333), ncol=3, dimnames = list(NULL, c("LOW", "MED","HIGH")))

nearest_road_cpt<-matrix(c(0.333333,0.333333,0.333333), ncol=3, dimnames = list(NULL, c("LOW", "MED","HIGH")))

connectivity_cpt<-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

landscape_nutritional_score_cpt<-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

pixel_nutritional_score_cpt <- matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

dams_cpt <- landscape_nutritional_score_cpt<-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

lf_density_cpt <-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))
```

```{r}
#CPT for Disturbance Index ####

disturb_index_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3, 3, 3),  # Shape of the array (3x3x3)
  dimnames = list(
    disturb_index = c("LOW", "MED", "HIGH"),
    nearest_road = c("LOW", "MED", "HIGH"),
    urban_prox = c("LOW","MED","HIGH"),
    connectivity = c("LOW", "MED", "HIGH")
  )
)
#order = disturb_index, road_density, urban_prox,connectivity

#table1: urban_prox = LOW, connectivity = LOW
disturb_index_cpt["LOW","LOW","LOW","LOW"] <- 1
disturb_index_cpt["MED","MED","LOW","LOW"] <- 1
disturb_index_cpt["HIGH","HIGH","LOW","LOW"] <- 1
#table2:  urban_prox = MED, connectivity = LOW
disturb_index_cpt["MED","LOW","MED","LOW"] <- 1
disturb_index_cpt["MED","MED","MED","LOW"] <- 1
disturb_index_cpt["HIGH","HIGH","MED","LOW"] <- 1
#table3: urban_prox = HIGH, connectivity = LOW
disturb_index_cpt["HIGH","LOW","HIGH","LOW"] <- 1
disturb_index_cpt["HIGH","MED","HIGH","LOW"] <- 1
disturb_index_cpt["HIGH","HIGH","HIGH","LOW"] <- 1
#table4: urban_prox = LOW, connectivity = MED
disturb_index_cpt["LOW","LOW","LOW","MED"] <- 1
disturb_index_cpt["MED","MED","LOW","MED"] <- 1
disturb_index_cpt["MED","HIGH","LOW","MED"] <- 1
#table5: urban_prox = MED, connectivity = MED
disturb_index_cpt["LOW","LOW","MED","MED"] <- 1
disturb_index_cpt["MED","MED","MED","MED"] <- 1
disturb_index_cpt["HIGH","HIGH","MED","MED"] <- 1
#table6: urban_prox = HIGH, connectivity = MED
disturb_index_cpt["HIGH","LOW","HIGH","MED"] <- 1
disturb_index_cpt["HIGH","MED","HIGH","MED"] <- 1
disturb_index_cpt["HIGH","HIGH","HIGH","MED"] <- 1
#table7: urban_prox = LOW, connectivity = HIGH
disturb_index_cpt["LOW","LOW","LOW","HIGH"] <- 1
disturb_index_cpt["LOW","MED","LOW","HIGH"] <- 1
disturb_index_cpt["MED","HIGH","LOW","HIGH"] <- 1
#table8: urban_prox = MED, connectivity = HIGH
disturb_index_cpt["MED","LOW","MED","HIGH"] <- 1
disturb_index_cpt["MED","MED","MED","HIGH"] <- 1
disturb_index_cpt["HIGH","HIGH","MED","HIGH"] <- 1
#table9: urban_prox = HIGH, connectivity = HIGH
disturb_index_cpt["MED","LOW","HIGH","HIGH"] <- 1
disturb_index_cpt["MED","MED","HIGH","HIGH"] <- 1
disturb_index_cpt["HIGH","HIGH","HIGH","HIGH"] <- 1

disturb_index_cpt

```

```{r}
#CPT for Nutritional Pixel Index ####

NPI_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3),  # Shape of the array (3x3)
  dimnames = list(
    NPI = c("LOW", "MED", "HIGH"),
    pixel_nutritional_score = c("LOW", "MED", "HIGH")
  )
)

NPI_cpt["LOW","LOW"] <- 1
NPI_cpt["MED","MED"] <- 1
NPI_cpt["HIGH","HIGH"] <- 1

NPI_cpt
```

```{r}
#CPT for Nutritional Landscape Index ####


NLI_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3, 3),  # Shape of the array (3x3)
  dimnames = list(
    NLI = c("LOW", "MED", "HIGH"),
    landscape_nutritional_score = c("LOW", "MED", "HIGH"),
    lf_density = c("LOW","MED","HIGH")
  )
)
#order = NLI, landscape_nutritional_score, lf_density
NLI_cpt["LOW","LOW","LOW"] <- 1
NLI_cpt["MED","MED","LOW"] <- 1
NLI_cpt["HIGH","HIGH","LOW"] <- 1
NLI_cpt["LOW","LOW","MED"] <- 1
NLI_cpt["HIGH","MED","MED"] <- 1
NLI_cpt["HIGH","HIGH","MED"] <- 1
NLI_cpt["MED","LOW","HIGH"] <- 1
NLI_cpt["HIGH","MED","HIGH"] <- 1
NLI_cpt["HIGH","HIGH","HIGH"] <- 1

NLI_cpt
```

```{r}
#CPT for Thermoregulation Index ####

thermoreg_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3),  # Shape of the array
  dimnames = list(
    thermoreg_index = c("LOW", "MED", "HIGH"),
    DAMS_MAX = c("LOW", "MED", "HIGH")
  )
)
#order = DAMS_MAX, thermoreg_index
thermoreg_cpt["LOW","LOW"] <- 1
thermoreg_cpt["MED","MED"] <- 1
thermoreg_cpt["HIGH","HIGH"] <- 1

thermoreg_cpt
```

```{r}
#Final CPT: deer damage risk

damage_risk_cpt <- array(
  0,  # Default probability for each cell
  dim = c(5, 3, 3, 3, 3),  # Shape of the array
  dimnames = list(
    damage_risk = c("LOW","LOW-MED","MED","MED-HIGH","HIGH"),
    NPI = c("LOW", "MED", "HIGH"),
    NLI = c("LOW", "MED", "HIGH"),
    thermoreg_index = c("LOW", "MED", "HIGH"),
    disturb_index = c("LOW", "MED", "HIGH")
    
  )
)
#order = damage_risk, patch_quality,landscape_nutrition,thermoreg_index,disturbance

#table1 landscape_nutrition = LOW, thermoreg_index = LOW, disturbance_index = LOW
damage_risk_cpt["LOW-MED","LOW","LOW","LOW","LOW"] <- 1 
damage_risk_cpt["LOW-MED","MED","LOW","LOW","LOW"] <- 1
damage_risk_cpt["MED","HIGH","LOW","LOW","LOW"] <- 1
#table2 landscape_nutrition = MED, thermoreg_index = LOW, disturbance_index = LOW
damage_risk_cpt["LOW-MED","LOW","MED","LOW","LOW"] <- 1
damage_risk_cpt["LOW-MED","MED","MED","LOW","LOW"] <- 1
damage_risk_cpt["MED","HIGH","MED","LOW","LOW"] <- 1
#table3 landscape_nutrition = HIGH, thermoreg_index = LOW, disturbance_index = LOW
damage_risk_cpt["LOW-MED","LOW","HIGH","LOW","LOW"] <- 1
damage_risk_cpt["MED","MED","HIGH","LOW","LOW"] <- 1
damage_risk_cpt["MED","HIGH","HIGH","LOW","LOW"] <- 1
#table4 landscape_nutrition = LOW, thermoreg_index = MED, disturbance_index = LOW
damage_risk_cpt["MED","LOW","LOW","MED","LOW"] <- 1
damage_risk_cpt["MED","MED","LOW","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","LOW","MED","LOW"] <- 1
#table5 landscape_nutrition = MED, thermoreg_index = MED, disturbance_index = LOW
damage_risk_cpt["MED","LOW","MED","MED","LOW"] <- 1
damage_risk_cpt["MED","MED","MED","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","MED","MED","LOW"] <- 1
#table6 landscape_nutrition = HIGH, thermoreg_index = MED, disturbance_index = LOW
damage_risk_cpt["MED","LOW","HIGH","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","MED","HIGH","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","HIGH","MED","LOW"] <- 1
#table7  landscape_nutrition = LOW, thermoreg_index = HIGH, disturbance_index = LOW
damage_risk_cpt["MED-HIGH","LOW","LOW","HIGH","LOW"] <- 1
damage_risk_cpt["MED-HIGH","MED","LOW","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","HIGH","LOW","HIGH","LOW"] <- 1
#table8 landscape_nutrition = MED, thermoreg_index = HIGH, disturbance_index = LOW
damage_risk_cpt["MED-HIGH","LOW","MED","HIGH","LOW"] <- 1
damage_risk_cpt["MED-HIGH","MED","MED","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","HIGH","MED","HIGH","LOW"] <- 1
#table9 landscape_nutrition = HIGH, thermoreg_index = HIGH, disturbance_index = LOW
damage_risk_cpt["MED-HIGH","LOW","HIGH","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","MED","HIGH","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","HIGH","HIGH","HIGH","LOW"] <- 1
#table10 landscape_nutrition = LOW, thermoreg_index = LOW, disturbance_index = MED
damage_risk_cpt["LOW","LOW","LOW","LOW","MED"] <- 1
damage_risk_cpt["LOW","MED","LOW","LOW","MED"] <- 1
damage_risk_cpt["LOW-MED","HIGH","LOW","LOW","MED"] <- 1
#table11 landscape_nutrition = MED, thermoreg_index = LOW, disturbance_index = MED
damage_risk_cpt["LOW-MED","LOW","MED","LOW","MED"] <- 1
damage_risk_cpt["LOW-MED","MED","MED","LOW","MED"] <- 1
damage_risk_cpt["MED","HIGH","MED","LOW","MED"] <- 1
#table12 landscape_nutrition = HIGH, thermoreg_index = LOW, disturbance_index = MED
damage_risk_cpt["LOW-MED","LOW","HIGH","LOW","MED"] <- 1
damage_risk_cpt["MED","MED","HIGH","LOW","MED"] <- 1
damage_risk_cpt["MED","HIGH","HIGH","LOW","MED"] <- 1
#table13 landscape_nutrition = LOW, thermoreg_index = MED, disturbance_index = MED
damage_risk_cpt["LOW-MED","LOW","LOW","MED","MED"] <- 1
damage_risk_cpt["LOW-MED","MED","LOW","MED","MED"] <- 1
damage_risk_cpt["MED","HIGH","LOW","MED","MED"] <- 1
#table14 landscape_nutrition = MED, thermoreg_index = MED, disturbance_index = MED
damage_risk_cpt["MED","LOW","MED","MED","MED"] <- 1
damage_risk_cpt["MED","MED","MED","MED","MED"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","MED","MED","MED"] <- 1
#table15  landscape_nutrition = HIGH, thermoreg_index = MED, disturbance_index = MED
damage_risk_cpt["MED","LOW","HIGH","MED","MED"] <- 1
damage_risk_cpt["MED-HIGH","MED","HIGH","MED","MED"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","HIGH","MED","MED"] <- 1
#table16 landscape_nutrition = LOW, thermoreg_index = HIGH, disturbance_index = MED
damage_risk_cpt["MED","LOW","LOW","HIGH","MED"] <- 1
damage_risk_cpt["MED","MED","LOW","HIGH","MED"] <- 1
damage_risk_cpt["MED","HIGH","LOW","HIGH","MED"] <- 1
#table17 landscape_nutrition = MED, thermoreg_index = HIGH, disturbance_index = MED
damage_risk_cpt["MED-HIGH","LOW","MED","HIGH","MED"] <- 1
damage_risk_cpt["MED-HIGH","MED","MED","HIGH","MED"] <- 1
damage_risk_cpt["HIGH","HIGH","MED","HIGH","MED"] <- 1
#table18 landscape_nutrition = HIGH, thermoreg_index = HIGH, disturbance_index = MED
damage_risk_cpt["MED-HIGH","LOW","HIGH","HIGH","MED"] <- 1
damage_risk_cpt["HIGH","MED","HIGH","HIGH","MED"] <- 1
damage_risk_cpt["HIGH","HIGH","HIGH","HIGH","MED"] <- 1
#table19 landscape_nutrition = LOW, thermoreg_index = LOW, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","LOW","LOW","HIGH"] <- 1
damage_risk_cpt["LOW","MED","LOW","LOW","HIGH"] <- 1
damage_risk_cpt["LOW","HIGH","LOW","LOW","HIGH"] <- 1
#table20 landscape_nutrition = MED, thermoreg_index = LOW, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","MED","LOW","HIGH"] <- 1
damage_risk_cpt["LOW","MED","MED","LOW","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","MED","LOW","HIGH"] <- 1
#table21 landscape_nutrition = HIGH, thermoreg_index = LOW, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","HIGH","LOW","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","HIGH","LOW","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","HIGH","LOW","HIGH"] <- 1
#table22 landscape_nutrition = LOW, thermoreg_index = MED, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","LOW","MED","HIGH"] <- 1
damage_risk_cpt["LOW","MED","LOW","MED","HIGH"] <- 1
damage_risk_cpt["LOW","HIGH","LOW","MED","HIGH"] <- 1
#table23 landscape_nutrition = MED, thermoreg_index = MED, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","MED","MED","HIGH"] <- 1
damage_risk_cpt["LOW","MED","MED","MED","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","MED","MED","HIGH"] <- 1
#table24 landscape_nutrition = HIGH, thermoreg_index = MED, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","HIGH","MED","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","HIGH","MED","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","HIGH","MED","HIGH"] <- 1
#table25  landscape_nutrition = LOW, thermoreg_index = HIGH, disturbance_index = HIGH
damage_risk_cpt["LOW-MED","LOW","LOW","HIGH","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","LOW","HIGH","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","LOW","HIGH","HIGH"] <- 1
#table26 landscape_nutrition = MED, thermoreg_index = HIGH, disturbance_index = HIGH
damage_risk_cpt["LOW-MED","LOW","MED","HIGH","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","MED","HIGH","HIGH"] <- 1
damage_risk_cpt["MED","HIGH","MED","HIGH","HIGH"] <- 1
#table27 landscape_nutrition = HIGH, thermoreg_index = HIGH, disturbance_index = HIGH
damage_risk_cpt["LOW-MED","LOW","HIGH","HIGH","HIGH"] <- 1
damage_risk_cpt["MED","MED","HIGH","HIGH","HIGH"] <- 1
damage_risk_cpt["MED","HIGH","HIGH","HIGH","HIGH"] <- 1

damage_risk_cpt
```

#### Create BBN structure

```{r}
# Step 1: Explicitly define the nodes in the network
nodes <- c(
  "disturb_index",
  "urban_prox",
  "nearest_road",
  "connectivity",
  "pixel_nutritional_score",
  "landscape_nutritional_score",
  "NLI",
  "lf_density",
  "thermoreg_index",
  "DAMS_MAX",
  "damage_risk",
  "NPI"
)

#empty graph
e = empty.graph(nodes)

arc.set = matrix(c("urban_prox", "disturb_index",
                     "nearest_road", "disturb_index", 
                     "connectivity", "disturb_index",
                     "lf_density","NLI",
                     "landscape_nutritional_score","NLI",
                     "pixel_nutritional_score","NPI",
                     "DAMS_MAX","thermoreg_index",
                     "disturb_index","damage_risk",
                     "NLI","damage_risk",
                     "NPI","damage_risk",
                     "thermoreg_index","damage_risk"),
             ncol = 2, byrow = TRUE,
             dimnames = list(NULL, c("from", "to")))

arcs(e) <- arc.set

model_string <- modelstring(e) 

net<-model2network(model_string) 
# You dont need to repeat the names of the nodes with parents (child nodes) 

# Custom fitting network (matching up the nodes to their CPTs)
dfit = custom.fit(net, dist = list(nearest_road=nearest_road_cpt, 
                                   urban_prox=urban_prox_cpt,
                                   connectivity=connectivity_cpt,
                                   lf_density=lf_density_cpt,
                                   DAMS_MAX=dams_cpt,
                      pixel_nutritional_score=pixel_nutritional_score_cpt,
                      landscape_nutritional_score=landscape_nutritional_score_cpt, 
                                   disturb_index=disturb_index_cpt, 
                                   NPI=NPI_cpt,
                                   NLI=NLI_cpt,
                                   thermoreg_index=thermoreg_cpt,
                                   damage_risk=damage_risk_cpt))

```

#### Plot BN structure

```{r plot the network}
#| echo: true
#| warning: false
#| message: false

graphviz.plot(net)

```

#### Check Parameters

```{r look at the dfit}
#| echo: true
#| warning: false
#| message: false
dfit
```

#### Set evidence

```{r set evidence}
#| echo: true
#| warning: false
#| message: false
cpquery(dfit, c(lf_density == "LOW"|DAMS_MAX == "HIGH"|landscape_nutritional_score == "HIGH"|connectivity=="LOW"|urban_prox=="LOW"|nearest_road == "HIGH"|DAMS_MAX=="HIGH"), (damage_risk=="HIGH"))

```

### Make predictions from BBN

```{r}
predict_dat <- df_categorical %>%st_drop_geometry() %>% #drop geometry
   mutate(across(where(is.character), toupper)) %>% #convert characters to upper case
   dplyr::select(-c("patch_ID","class","Land.cover","scale","pixel_ID","x","y"))
#Ensure variable names match those in BBN
predict_dat<-predict_dat%>%rename(
                              nearest_road=nearest_road,
                              lf_density=lf,
                              urban_prox=urban_proximity,
                              landscape_nutritional_score=nli_edge_types_1000,
                              pixel_nutritional_score=Forage_Q,
                              DAMS_MAX=max.dams.1000,
                              connectivity=total_connect_log)

#Make empty columns for the latent (unobserved) variables

predict_dat$NPI <- NA
predict_dat$NLI <- NA
predict_dat$disturb_index <- NA
predict_dat$thermoreg_index <- NA
predict_dat$damage_risk <- NA

#Ensure all columns are factors, not characters
predict_dat <- predict_dat %>% mutate_all(as.factor)

#Ensure factor levels are in correct order
levels(predict_dat$nearest_road)<-c("LOW", "MED", "HIGH")
levels(predict_dat$landscape_nutritional_score)<-c("LOW", "MED", "HIGH")
levels(predict_dat$pixel_nutritional_score)<-c("LOW", "MED", "HIGH")
levels(predict_dat$urban_prox)<-c("LOW", "MED", "HIGH")
levels(predict_dat$lf_density)<-c("LOW", "MED", "HIGH")
levels(predict_dat$DAMS_MAX)<-c("LOW", "MED", "HIGH")
levels(predict_dat$connectivity)<-c("LOW", "MED", "HIGH")
levels(predict_dat$NPI)<-c("LOW", "MED", "HIGH")
levels(predict_dat$NLI)<-c("LOW", "MED", "HIGH")
levels(predict_dat$disturb_index)<-c("LOW", "MED", "HIGH")
levels(predict_dat$thermoreg_index)<-c("LOW", "MED", "HIGH")
levels(predict_dat$damage_risk)<-c("LOW","LOW-MED","MED","MED-HIGH","HIGH")

#Ensure predict_data is a data.frame
predict_dat <- as.data.frame(predict_dat)

#Predict values for latent variables
pred_nli = predict(dfit,node="NLI",data=predict_dat, method = "bayes-lw")
pred_npi = predict(dfit,node="NPI",data=predict_dat, method = "bayes-lw")
pred_thermoreg = predict(dfit,node="thermoreg_index",data=predict_dat, method = "bayes-lw")
pred_disturb = predict(dfit,node="disturb_index",data=predict_dat, method = "bayes-lw")

#fill the columns
predict_dat$NLI <- pred_nli
predict_dat$NPI <- pred_npi
predict_dat$thermoreg_index <- pred_thermoreg
predict_dat$disturb_index <- pred_disturb
  
#Predict damage
pred_damage = predict(dfit,node = "damage_risk",data = predict_dat,method = "bayes-lw")

predict_dat$damage_risk <- pred_damage

#Add damage risk back into spatial dataset

df_categorical$pred_damage <- pred_damage

#save
saveRDS(df_categorical, file = file.path(here("predicted_data_categorical_pixel_level_3km.rds")))
```

### Draw maps

```{r}
#| echo: true
#| warning: false
#| message: false


df_categorical <- readRDS(file = file.path(here("predicted_data_categorical_pixel_level_3km.rds")))

# Make sf object
df_sf <- st_as_sf(df_categorical, coords = c("x", "y"), crs = st_crs(bng))

# Define the value mapping for pred_damage
damage_values <- c("LOW" = 1, "LOW-MED" = 2, "MED" = 3, "MED-HIGH" = 4, "HIGH" = 5)
df_sf$value <- damage_values[df_sf$pred_damage]

# Rasterize
raster_data <- stars::st_rasterize(df_sf %>% dplyr::select(value, geometry))

# Get last 5 colours from YIOrRd color palette
color_palette <- c("LOW" = "yellow", "LOW-MED" = "#FED976", "MED" = "#FD8D3C", "MED-HIGH" = "#FC4E2A", "HIGH" = "#E31A1C")

# Initialize a list to store individual plots
plots <- list()

# Loop through each area in 'buffered_woods_3km'
for (i in 1:nrow(buffered_woods_5km)) {
  # Extract the i-th area
  area <- buffered_woods_5km[i, ]
  
  cropped_data <- st_crop(raster_data, area)
  
  # Convert the cropped data to a data frame for ggplot2
  cropped_df <- as.data.frame(cropped_data, xy = TRUE)
  
  # Create a factor for the damage categories
  cropped_df$value <- factor(cropped_df$value, levels = 1:5, labels = names(damage_values))
  
  # Replace NAs with 0
  cropped_df <- cropped_df %>%
    filter(!is.na(value))
  
  # Plot using ggplot2
  plot_title <- buffered_woods_5km$Region[i]
  
  plot <- ggplot(cropped_df) +
    geom_raster(aes(x = x, y = y, fill = value)) +
    scale_fill_manual(
      values = color_palette, 
      name = "Deer Impact Risk",
      breaks = names(damage_values),  # Ensure all levels are shown in the legend
      labels = names(damage_values)
    ) +
    theme_minimal() +
    coord_equal() +
    labs(
      title = plot_title,
      x = "Longitude",
      y = "Latitude"
    )
  
  # Store the plot in the list
  plots[[i]] <- plot
}

# Save each plot to a separate PDF file with 300 dpi resolution
output_dir <- here("output/figures")
dpi <- 300  # Resolution in dots per inch
width <- 6  # Width of the PDF in inches
height <- 6  # Height of the PDF in inches

for (i in 1:length(plots)) {
  plot <- plots[[i]]  # Get the plot from the list
  filename <- file.path(output_dir, paste0("25m_res_map_3km", i, ".pdf"))
  
  # Save the plot as a 300 dpi PDF
  ggsave(
    filename = filename, 
    plot = plot, 
    device = "pdf",  # Specify the file format
    width = width, 
    height = height, 
    dpi = dpi  # Set the resolution
  )
}

```

```{r}

#Adding the predicted damage data to an interactive satellite image ####
library(leaflet)

# Load the data
df_categorical <- readRDS(file = file.path(here("predicted_data_categorical_pixel_level_3km.rds")))

# Make sf object
df_sf <- st_as_sf(df_categorical, coords = c("x", "y"), crs = st_crs(bng))
df_sf <- st_transform(df_sf, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# Define the value mapping for pred_damage
damage_values <- c("LOW" = 1, "LOW-MED" = 2, "MED" = 3, "MED-HIGH" = 4, "HIGH" = 5)
df_sf$value <- damage_values[df_sf$pred_damage]

# Set up the colors
val <- 1:5
pal <- c("yellow", "#FED976", "#FD8D3C", "#FC4E2A", "#E31A1C")

#Convert to raster
r <- df_sf %>% dplyr::select(geometry, value) %>% stars::st_rasterize()

# Convert the raster to a data frame for ggplot
r_df <- as.data.frame(r, xy = TRUE)

# Create a factor for the value with all levels
r_df$value <- factor(r_df$value, levels = val, labels = names(damage_values))

# Add missing levels to the data
missing_levels <- data.frame(x = NA, y = NA, value = factor(val, labels = names(damage_values)))
r_df <- rbind(r_df, missing_levels)

# Plot raster using the color palette with custom labels
ggplot(r_df) +
  geom_raster(aes(x = x, y = y, fill = value)) +
  scale_fill_manual(values = pal, 
                    breaks = names(damage_values),
                    labels = names(damage_values),
                    na.value = "transparent") +
  theme_minimal() +
  labs(title = "Predicted Damage Raster",
       fill = "Damage Level") +
  guides(fill = guide_legend(override.aes = list(alpha = 1))) # Ensure all levels are shown in the legend

# Define the color palette function for leaflet
color_pal <- colorNumeric(palette = pal, domain = val, na.color = "transparent")

# Convert the stars object to a RasterLayer
raster_layer <- as(r, "Raster")

#Round all values to an integer
rounded_raster <- round(raster_layer)
#Project for leaflet map
rounded_raster <- projectRasterForLeaflet(rounded_raster, method="ngb")

# Create a color palette for the raster
color_fact <- colorFactor(
  palette = c("yellow", "#FED976", "#FD8D3C", "#FC4E2A", "#E31A1C"),
  domain = damage_values,
na.color = "transparent"  # Set the color for NA values to transparent

)

# Create a reversed mapping for labels
damage_labels <- names(damage_values)
names(damage_labels) <- damage_values

# Create the Leaflet map
leaflet() %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addRasterImage(rounded_raster, colors = color_fact, opacity = 0.9, project = FALSE) %>%
  addLegend(
    pal = color_fact,
    values = damage_values,  # Use the correct domain of values
    title = "Deer impact risk",
    labFormat = labelFormat(
      transform = function(x) {
        # Use the label mapping to get the labels
        damage_labels[as.character(x)]
      }
    )
  )

```

#END \####
