This script is testing changes to the BBN model for large deer impact risk following the final expert meeting summarising findings phase 2 model validation interviews (06/05/2025)

Changes made:

-   Disturbance (distance to nearest urban/suburban area removed)

-   Add in "Low" category to woodland "yummy-ness" or "palatability" for mature conifer woods

    --\> Try taking edges out of model and see what happens

```{r}
library(sf)
library(tidyverse)
library(raster)
library(igraph)
```

#tiles \<- ew10k %\>% filter(tile_name %in% c("TL87","TL88","TL97","TL98")) #northampton

#tiles \<- ew10k %\>% filter(tile_name %in% c("NY21","NY22","NY12","NY11")) #lake district

#tiles \<- ew10k %\>% filter(tile_name %in% c("SE71","SE72","SE61","SE62")) #thorne moors

#tiles \<- ew10k %\>% filter(tile_name %in% c("SX16","SX17","SX27","SX26")) #cornwall

tiles \<- ew10k %\>% filter(tile_name %in% c("SU20","SU30","SZ29","SZ39")) #new forest

```{r}
#Read in NFI 2023 dataset

nfi <- st_read("C:/Users/ik929086/Documents/iDeer/data/raw-data/nfi-2023-raw/National_Forest_Inventory_GB_2023.shp")

unique(nfi$IFT_IOA)

#England and Wales 10k squares
ew10k <- sf::st_read("C:/Users/ik929086/Documents/iDeer/data/derived-data/10k_tiles_EW.shp")
bng <- 27700

#Get Chilterns tile as an example
#SU79 <- ew10k %>% filter(tile_name %in% c("SU79"))

#tiles <- ew10k %>% filter(tile_name %in% c("TL87","TL88","TL97","TL98")) #northampton
#tiles <- ew10k %>% filter(tile_name %in% c("NY21","NY22","NY12","NY11")) #lake district
#tiles <- ew10k %>% filter(tile_name %in% c("SE71","SE72","SE61","SE62")) #thorne moors
#tiles <- ew10k %>% filter(tile_name %in% c("SX16","SX17","SX27","SX26")) #cornwall
tiles <- ew10k %>% filter(tile_name %in% c("SU20","SU30","SZ29","SZ39")) #new forest

plot(tiles$geometry)
```

```{r}
#Explore dataset

woods_only <- nfi %>% filter(CATEGORY %in% c("Woodland"))
unique(woods_only$IFT_IOA)

non_woods <- nfi %>% filter(CATEGORY %in% c("Non woodland"))
unique(non_woods$IFT_IOA)
#Grassland, Agricultural land, Other vegetation and bare area would be areas for foraging
#Other things not usable habitat

#Mixed structure habitats = open areas, or areas with good potential for foraging resources
open_woodland <- nfi %>% filter(IFT_IOA %in% c("Assumed woodland","Failed","Windblow","Ground prep","Felled","Low density", "Young trees","Shrub"))

#Closed habitats = mature woodland
closed_woodland <- nfi %>% filter(IFT_IOA %in% c("Conifer","Broadleaved","Mixed mainly broadleaved","Mixed mainly conifer","Coppice with standards","Coppice"))

#subset for testing

closed_woodland <- st_filter(closed_woodland, tiles)
open_woodland <- st_filter(open_woodland, tiles)

plot(closed_woodland$geometry, col = "red")
plot(open_woodland$geometry, col = "blue",add=TRUE)
```

```{r}
#Make an NFI raster of the two habitat types

# Create the IFT_ID column based on the IFT_IOA values
#1 = open woodland types
#2 = mature woodland types
#3 = non woodland habitats with potential for foraging

# nfi <- nfi %>%
#   mutate(group_ID = case_when(
#     IFT_IOA %in% c("Assumed woodland","Failed","Windblow","Ground prep","Felled","Low density", "Young trees","Shrub") ~ 1,
#     IFT_IOA %in% c("Conifer","Broadleaved","Mixed mainly broadleaved","Mixed mainly conifer","Coppice with standards","Coppice") ~ 2,
#     #IFT_IOA %in% c("Agricultural land","Grassland","Other vegetation") ~ 3, #don't need this line- boundaries function automatically obtains edges of classes (1 and 2) with NAs
#     TRUE ~ NA_real_  # This handles other values (if needed, or you can leave it out)
#   ))
# 
# #Make raster of NFI
 template_rast <- raster("C:/Users/ik929086/Documents/iDeer/data/raw-data/lcm-2023/gblcm2023_25m.tif")
 crs(template_rast) <- bng
# 
# nfi_rast <- fasterize::fasterize(nfi, template_rast, field = "group_ID")
# 
# # Get boundaries between 1 and 2
#  boundaries_nfi <- boundaries(nfi_rast, type = "inner", classes = TRUE, #make edges between classes as well as NA
#                               directions = 8, #search for edges in 8 directions
#                               asNA = TRUE) #non-edges returned as NA
#  plot(boundaries_nfi)
# crs(boundaries_nfi) <- bng
# 
# terra::writeRaster(boundaries_nfi, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/boundaries_open_closed_woods_NFI_2023_bng.tif",overwrite=TRUE)
 
boundaries_nfi <- raster("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/boundaries_open_closed_woods_NFI_2023_bng.tif")

#plot(boundaries_nfi)

# boundaries_nfi_sample <- raster::crop(boundaries_nfi, tiles)
# 
# # Plot the other geometries
# plot(closed_woodland$geometry, col = "forestgreen")
# plot(open_woodland$geometry, col = "magenta", add = TRUE)
# 
# #plot boundaries - internal boundaries and edge boundaries
# plot(boundaries_nfi_sample)
```

```{r}
#Open woodland types will have a high risk of impact from small deer
#Closed woodland types will have a lower risk of impact

#Add a column woodland_vulnerability to woods_only with following conditions
#Where open habitats have highest
#3 = "Recent woodland"
#2 = mature broadleaved woodland
#1 = mature conifer woodland

# woods_only <- woods_only %>%
#   mutate(woodland_food_value = case_when(
#     IFT_IOA %in% c("Assumed woodland","Failed","Windblow","Ground prep","Felled","Low density", "Young trees","Shrub") ~ 3,
#     IFT_IOA %in% c("Broadleaved", "Mixed mainly broadleaved","Coppice with standards","Coppice") ~ 2,
#     IFT_IOA %in% c("Conifer","Mixed mainly conifer") ~ 1,
#     TRUE ~ NA_real_  # This handles other values (if needed, or you can leave it out)
#   ))
# 
# woodland_food_value_raster <- fasterize::fasterize(woods_only, template_rast,
#                                                    field = "woodland_food_value")
# crs(woodland_food_value_raster) <- bng
# 
# terra::writeRaster(woodland_food_value_raster, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_woodland_food_value_index_large_deer.tif",
#               overwrite=TRUE)

#plot(woodland_food_value_raster)

woodland_food_value_raster <- raster("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_woodland_food_value_index_large_deer.tif")
#woodland_food_value_raster <- crop(woodland_food_value_raster, tiles)
#plot(woodland_food_value_raster)
```

```{r}
# #PERCENTAGE COVER WOODLAND EDGE
# #Edges now include both external woodland edges and interior woodland edges produced in boundaries_nfi above
# 
# #create circular buffers around every pixel of 1000 metres
# circle.buff = raster::focalWeight (boundaries_nfi, d=1000, type="circle",fillNA=T )#Create buffer
# circle.buff[circle.buff > 0] <- 1   # replacing weights by 1
# Focal1000_EDGES= raster::focal(boundaries_nfi, w=circle.buff, fun=sum, na.rm=T,pad=TRUE, padValue=NA)
# 
# crs(Focal1000_EDGES) <- bng
# 
# writeRaster(Focal1000_EDGES, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_focal_woodland_edge_area_1000m.tif",
#              overwrite=TRUE)
# 
# #Divide number of pixels by the total number of 25m2 pixels in a 1000m buffer*100
# Focal1000_EDGES_PERC <- (Focal1000_EDGES/5027)*100
# plot(Focal1000_EDGES_PERC)
# plot(boundaries_nfi_sample,add=TRUE)
# 
# writeRaster(Focal1000_EDGES_PERC, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_focal_woodland_edge_area_1000m_percentage_cover.tif",
#              overwrite=TRUE)
# 
# Focal1000_EDGES_PERC <- raster("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_focal_woodland_edge_area_1000m_percentage_cover.tif")
# #Focal1000_EDGES_PERC <- crop(Focal1000_EDGES_PERC, tiles)
# #plot(Focal1000_EDGES_PERC)
```

```{r}
#PERCENTAGE COVER PERENNIAL 

#ON INSPECTION, CEH LAND COVER MAP DOES A GOOD JOB OF CATCHING PERENNIAL VEGETATION INSIDE WOODS
#DON'T NEED TO COMBINE WITH PERENNIAL LAND COVER FROM NON WOODS IN NFI DATASET

# lcm <- raster("./data/raw-data/lcm-2023/gblcm2023_25m.tif")
# crs(lcm) <- bng
# #make peren map only
# reclass_peren <- data.frame(is = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,
#                                     18,19,20,21), becomes = c(0,0,0,1,1,1,1,1,1,1,1,0,0,0,0,0,0,0,1,0,0))
# peren_raster_GB <- reclassify(lcm, reclass_peren)
# crs(peren_raster_GB) <- bng
# writeRaster(peren_raster_GB, here("./output/GB_datasets_2023/peren_raster_GB_2023.tif"),overwrite=TRUE)

#peren_raster_GB <- raster("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/peren_raster_GB_2023.tif")

#get perennial cover inside NFI woodlands
#In NFI documentation, Agricultural land could be pasture or arable
#Took decision to include in perennial
# unique(non_woods$IFT_IOA)
# perennial_NFI <- nfi %>% filter(IFT_IOA %in% c("Grassland","Other vegetation","Agricultural land"))
# perennial_NFI_raster <- fasterize::fasterize(perennial_NFI, template_rast)
# writeRaster(perennial_NFI_raster, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/NFI_peren_raster_GB_2023.tif")
# perennial_NFI_raster_test <- crop(perennial_NFI_raster, tiles)
# peren_raster_test <- crop(peren_raster_GB, tiles)
# #Merge the perennial GB map with perennial NFI vegetation
# nfi_lcm_perennial_raster <- mosaic(perennial_NFI_raster_test, peren_raster_test, fun = max)
# plot(perennial_NFI_raster_test)
# plot(peren_raster_test)
# plot(nfi_lcm_perennial_raster)

# #Focal extraction of perennial cover
# circle.buff = raster::focalWeight(peren_raster_GB, d=1000, type="circle",fillNA=T )#Create buffer
# circle.buff[circle.buff > 0] <- 1   # replacing weights by 1
# Focal1000_peren= raster::focal(x=peren_raster_GB, w=circle.buff, fun=sum, na.rm=T, pad=TRUE, padValue=NA)
#                        
# #Divide number of pixels by the total number of 25m2 pixels in a buffer*100 = percentage peren
# Focal1000_PEREN_PERC <- (Focal1000_peren/5027)*100
# 
# plot(Focal1000_PEREN_PERC)

Focal1000_PEREN_PERC <- raster("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/percentage_cover_perennial_1000m_2023.tif")
#Focal1000_PEREN_PERC <- crop(Focal1000_PEREN_PERC, tiles)
#plot(Focal1000_PEREN_PERC)
```

```{r}
#PERCENTAGE COVER ARABLE
#GB land cover map
# lcm <- raster("./data/raw-data/lcm-2023/gblcm2023_25m.tif")
# crs(lcm) <- bng
# #make arable map only
# reclass_arable <- data.frame(is = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,
#                               18,19,20,21), becomes = c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0))
# arable_raster_GB <- reclassify(lcm, reclass_arable)
# crs(arable_raster_GB) <- bng
# writeRaster(arable_raster_GB, here("output/GB_datasets_2023/2023/arable_raster_GB_2023.tif"),overwrite=TRUE)

# arable_raster_GB <- raster("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/arable_raster_GB_2023.tif")
# 
# #arable_raster_test <- crop(arable_raster_GB, tiles)
# #plot(arable_raster_test)
# 
# circle.buff = raster::focalWeight(arable_raster_GB, d=1000, type="circle",fillNA=T )#Create buffer
# circle.buff[circle.buff > 0] <- 1   # replacing weights by 1
# Focal1000_arable= raster::focal(x=arable_raster_GB, w=circle.buff, fun=sum, na.rm=T, pad=TRUE, padValue=NA)
# #Divide number of pixels by the total number of 25m2 pixels in a buffer*100 = percentage arable
# Focal1000_ARABLE_PERC <- (Focal1000_arable/5027)*100
# plot(Focal1000_ARABLE_PERC)
#writeRaster(Focal1000_ARABLE_PERC, "./output/GB_datasets_2023/large_deer/percentage_cover_arable_1000_2023.tif")

Focal1000_ARABLE_PERC <- raster("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/percentage_cover_arable_1000_2023.tif")
```

```{r}
#EXPOSURE OF NON-WOODLAND AND OPEN WOODLAND TYPES WITHIN 1000m

#Get woodlands only
#woods_only_test <- st_filter(woods_only, tiles)

# #Get merged polygons for thermoreg map ####
# 
# # Step 1: Get touching relationships
# touching <- st_touches(woods_only_test)
# 
# # Step 2: Build graph and identify groups
# g <- graph_from_adj_list(touching)
# groups <- components(g)$membership
# 
# # Step 3: Add group ID to the data
# woods_only_test$group_ID <- groups
# 
# # Step 4: Dissolve (union) polygons by group
# merged_woods <- woods_only_test %>%
#   group_by(group_ID) %>%
#   summarise(geometry = st_union(geometry), .groups = "drop")
# 
# plot(merged_woods$geometry)

merged_woods <- read_rds("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/small_deer/NFI_2023_all_woods_UK_touching_merged.rds")

# #DAMS
# #Interpolated from 50m resolution to 25m resolution
# #To match the resolution of other raster layers
# #Interpolation method = bilinear
dams <- raster("C:/Users/ik929086/Documents/iDeer/data/raw-data/DAMS/dams_25m_bng.tif")
crs(dams) <- bng
# Crop dams to match the extent of lcm
dams_crop <- crop(dams, extent(template_rast))
#Reclass lcm so all habitats set to 1 (including woods)
#Provides a blank slate that we can then burn the NFI woods into as zeros
reclass <- data.frame(is = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,
                             18,19,20,21), becomes = c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1))
binary_raster_GB <- reclassify(template_rast, reclass)

#open_raster_GB <- raster("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/woods_only_raster_GB_2023.tif")
#crs(open_raster_GB) <- bng

#Reclass NFI so all shelter habitats set to 0
thermoreg_habitats <- nfi <- nfi %>%
  mutate(thermoreg_ID = case_when(
    IFT_IOA %in% c("Bare area", "Assumed woodland", "Failed",
                   "Windblow", "Grassland",
                   "Ground prep", "Agricultural land",
                   "Felled", "Low density", "Young trees","Urban","Windfarm","Quarry","River","Open water","Road", "Other vegetation") ~ 1,
    IFT_IOA %in% c("Conifer", "Broadleaved", "Mixed mainly broadleaved",
                   "Mixed mainly conifer","Shrub","Coppice","Coppice with standards","Power line") ~ 0,
    TRUE ~ NA_real_  # This handles other values (if needed, or you can leave it out)
  ))


#Make binary open habitat raster ####
nfi_rast_open <- fasterize::fasterize(thermoreg_habitats, template_rast,
                                      field = "thermoreg_ID")

#binary_raster_test <- crop(binary_raster_GB, tiles)
#nfi_rast_open_test <- crop(nfi_rast_open, tiles)
#Mosaic the two rasters, giving priority to the NFI
nfi_lcm_open_mosaic <- mosaic(nfi_rast_open, binary_raster_GB,
                              fun = min #zeros will remain 0
                              )
crs(nfi_lcm_open_mosaic) <- bng

writeRaster(nfi_lcm_open_mosaic, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/woods_only_raster_GB_NFI_LCM_mosaic_2023.tif",overwrite=TRUE)
#plot(nfi_rast_open_test)
#plot(open_raster_test)
#plot(nfi_lcm_open_mosaic_test)
#Multiply DAMs and open land cover raster together - woodlands will be set to 0 DAMS
DAMS_open_raster <- dams_crop*nfi_lcm_open_mosaic
writeRaster(DAMS_open_raster, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/DAMS_non_wood_raster_GB_NFI_LCM_mosaic_2023.tif",overwrite=TRUE)
plot(DAMS_open_raster)

#DAMS_open_raster_test <- crop(DAMS_open_raster, tiles)

circle.buff = raster::focalWeight(DAMS_open_raster, d=1000, type="circle",fillNA=T )#Create buffer
circle.buff[circle.buff > 0] <- 1   # replacing weights by 1
Focal1000_DAMS= raster::focal(x=DAMS_open_raster, w=circle.buff, fun=sum, na.rm=T, pad=TRUE, padValue=NA#,
                         #filename = here("output/GB_datasets_2023/large-deer/sum_dams_1000_non_wood_2023.tif"), 
                         #overwrite=T)#do focal window analysis with weights matrix
)
#plot(Focal1000_DAMS)
print("MAX DAMS 1km raster saved")
writeRaster(Focal1000_DAMS, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/summed_dams_1000_non_wood_NFI_LCM_mosaic_2023.tif")

Focal1000_DAMS <- raster("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/summed_dams_1000_non_wood_NFI_LCM_mosaic_2023.tif")

#Aggregate to patch, where the score for each patch is the normalised maximum summed DAMS score for any pixel in patch
# Assuming you have a patch identifier, e.g., 'patch_id' in your woodland polygons
patch_agg <- raster::extract(Focal1000_DAMS, merged_woods, fun=max, na.rm=TRUE, df=TRUE)

patch_agg <- patch_agg %>% rename(group_ID = ID,
                                  max_dams_1000_non_wood_2023 = summed_dams_1000_non_wood_NFI_LCM_mosaic_2023)

#left_join the max summed DAMS score to GB_polys_merged

polys_merged_max_summed_DAMS <- left_join(patch_agg, merged_woods, by = c("group_ID"))

polys_merged_max_summed_DAMS <- st_as_sf(polys_merged_max_summed_DAMS)

#saveRDS(polys_merged_max_summed_DAMS, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_max_summed_DAMS_extracted_to_polys_1000m.rds")

polys_merged_max_summed_DAMS <- read_rds("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_max_summed_DAMS_extracted_to_polys_1000m.rds")

#Rasterize

polys_merged_max_summed_DAMS_raster <- fasterize::fasterize(sf = polys_merged_max_summed_DAMS,
raster = template_rast,
field = "max_dams_1000_non_wood_2023")

summed_dams_min <- polys_merged_max_summed_DAMS_raster@data@min
summed_dams_max <- polys_merged_max_summed_DAMS_raster@data@max

#Normalise summed DAMS scores from 0-1
polys_merged_max_summed_DAMS_raster_normalised <- (polys_merged_max_summed_DAMS_raster - summed_dams_min) / 
  (summed_dams_max - summed_dams_min)

plot(polys_merged_max_summed_DAMS_raster_normalised)

#writeRaster(polys_merged_max_summed_DAMS_raster_normalised, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_CEH_LCM_2023_summed_max_dams_non_wood_1000m_normalised.tif")

polys_merged_max_summed_DAMS_raster_normalised <- raster("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_CEH_LCM_2023_summed_max_dams_non_wood_1000m_normalised.tif")

#polys_merged_max_summed_DAMS_raster_normalised <- crop(polys_merged_max_summed_DAMS_raster_normalised, tiles)
#plot(polys_merged_max_summed_DAMS_raster_normalised)
```

```{r}
#TEST - TRY EXTRACTING DAMS WITHOUT CLASSIFYING WOODS AS 0 EXPOSURE

#merged_woods <- read_rds("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/small_deer/NFI_2023_all_woods_UK_touching_merged.rds")
#st_crs(merged_woods) <- bng


merged_woods_crop <- st_filter(merged_woods, tiles)
merged_woods_crop$poly_id <- 1:nrow(merged_woods_crop)

template_rast_crop <- crop(template_rast, tiles)

# #DAMS
# #Interpolated from 50m resolution to 25m resolution
# #To match the resolution of other raster layers
# #Interpolation method = bilinear
#dams <- raster("C:/Users/ik929086/Documents/iDeer/data/raw-data/DAMS/dams_25m_bng.tif")
#crs(dams) <- bng

#crop dams to tiles

dams_crop <- crop(dams, tiles)

#DAMS_open_raster_test <- crop(DAMS_open_raster, tiles)

circle.buff = raster::focalWeight(dams_crop, d=1000, type="circle",fillNA=T )#Create buffer
circle.buff[circle.buff > 0] <- 1   # replacing weights by 1
Focal1000_DAMS= raster::focal(x=dams_crop, w=circle.buff, fun=sum, na.rm=T, pad=TRUE, padValue=NA#,
                         #filename = here("output/GB_datasets_2023/large-deer/sum_dams_1000_non_wood_2023.tif"), 
                         #overwrite=T)#do focal window analysis with weights matrix
)

#Aggregate to patch, where the score for each patch is the normalised maximum summed DAMS score for any pixel in patch
# Assuming you have a patch identifier, e.g., 'patch_id' in your woodland polygons
patch_agg <- raster::extract(Focal1000_DAMS, merged_woods_crop, fun=max, na.rm=TRUE, df=TRUE)

patch_agg$poly_id <- 1:nrow(patch_agg)

patch_agg <- patch_agg %>% rename(#group_ID = poly_id,
                                  max_dams_1000_non_wood_2023 = layer)

#left_join the max summed DAMS score to GB_polys_merged

polys_merged_max_summed_DAMS <- left_join(merged_woods_crop, patch_agg, by=c("poly_id"="poly_id"))

polys_merged_max_summed_DAMS <- st_as_sf(polys_merged_max_summed_DAMS)

#Rasterize

polys_merged_max_summed_DAMS_raster <- fasterize::fasterize(sf = polys_merged_max_summed_DAMS,
raster = template_rast_crop,
field = "max_dams_1000_non_wood_2023")

plot(polys_merged_max_summed_DAMS_raster)

#To normnalise, use min and max from full dataset
#This is from the layer where woods = 0 dams but fine for now
#> summed_dams_min
#[1] 2184.247
#> summed_dams_max
#[1] 117529.5

summed_dams_min <- 2184.247
summed_dams_max <- 117529.5

#Normalise summed DAMS scores from 0-1
polys_merged_max_summed_DAMS_raster_normalised <- (polys_merged_max_summed_DAMS_raster - summed_dams_min) / 
  (summed_dams_max - summed_dams_min)

plot(polys_merged_max_summed_DAMS_raster_normalised)
```

```{r}
#Only these woodland types will provide good shelter for thermoregulation:
#"Conifer", "Broadleaved", "Mixed mainly broadleaved",
#"Mixed mainly conifer","Shrub","Coppice","Coppice with standards"
#Classify NFI woods by their shelter quality

unique(woods_only$IFT_IOA)
#  [1] "Broadleaved"              "Assumed woodland"         "Conifer"                  "Felled"                  
#  [5] "Failed"                   "Ground prep"              "Low density"              "Mixed mainly broadleaved"
#  [9] "Mixed mainly conifer"     "Young trees"              "Coppice"                  "Coppice with standards"  
# [13] "Shrub"                    "Windblow"  

woods_only <- woods_only %>%
  mutate(woodland_shelter_value = case_when(
    IFT_IOA %in% c("Assumed woodland","Failed","Windblow","Ground prep","Felled","Low density", "Young trees") ~ 1,
    IFT_IOA %in% c("Coppice with standards","Coppice","Shrub") ~ 2,
    IFT_IOA %in% c("Broadleaved", "Mixed mainly broadleaved","Conifer","Mixed mainly conifer") ~ 3,
    TRUE ~ NA_real_  # This handles other values (if needed, or you can leave it out)
  ))

woodland_shelter_value_raster <- fasterize::fasterize(woods_only, template_rast,
                                                   field = "woodland_shelter_value")

crs(woodland_shelter_value_raster) <- bng

writeRaster(woodland_shelter_value_raster, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_woodland_shelter_value_index_large_deer.tif",overwrite=TRUE)

woodland_shelter_value_raster <- raster("C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_woodland_shelter_value_index_large_deer.tif")

#woodland_shelter_value_raster <- crop(woodland_shelter_value_raster, tiles)
#plot(woodland_shelter_value_raster)
```

```{r}
#function to extract pixel values from raster layers
source("C:/Users/ik929086/Documents/iDeer/code/functions/extract_raster_pixels_func_10km_chunks.R")

```

```{r}
#Conditional probability tables ####

large_attraction_cpt_df <- read.csv("C:/Users/ik929086/Documents/iDeer/output/CPTs/Large_Deer_Attraction_Index_CPT_no_edges.csv")
#Thermoregulation Index
large_thermoreg_cpt_df <- read.csv("C:/Users/ik929086/Documents/iDeer/output/CPTs/Large_Deer_Thermoreg_Index_CPT_post_interviews.csv")
#Connectivity Index
#large_connectivity_cpt_df <- read.csv("C:/Users/ik929086/Documents/iDeer/data/CPTs/Largedeer_Connectivity_Index_Woodland_cover_1000m_Amy_EDITED.csv")

#Large deer impact risk
large_risk_cpt_df <- read.csv("C:/Users/ik929086/Documents/iDeer/output/CPTs/Large_Deer_Damage_Risk_CPT_post_interviews.csv")

#Disturbance Index
large_urban_dist_cpt_df <- read.csv("C:/Users/ik929086/Documents/iDeer/data/CPTs/Largedeer_Urban_Distance_Index_Urban_proximity_Amy_EDITED.csv")

#woodland food value index

large_wood_food_value_cpt_df <- read.csv("C:/Users/ik929086/Documents/iDeer/data/CPTs/Largedeer_Woodland_food_value_Index.csv")
```

```{r}
#Create a dataframe containing all extracted raster values within user's landscape

#template_rast_tiles <- crop(template_rast, tiles)

#tiles <- ew10k %>% filter(tile_name %in% c("SU79","SU78","SU88","SU89"))

#EXTRACT WOOD EDGE / ARABLE AREA

#edgepix <- extract_raster(map= Focal1000_EDGES_PERC,
#                          lcm = template_rast, #lcm
#                          tiles = tiles) #ew10k

#saveRDS(edgepix, "C:/Users/ik929086/Documents/iDeer/output/edge_pix_1000m_df_EW.RDS")
#read
#edgepix<- readRDS("C:/Users/ik929086/Documents/iDeer/output/edge_pix_1000m_df_EW.RDS")


#EXTRACT WOODLAND AREA WITHIN 1km
#woodpix <- extract_raster(map = Focal1000_WOODS_PERC,
#                          lcm = lcm,
#                          tiles = ew10k)
#saveRDS(woodpix, "C:/Users/ik929086/Documents/iDeer/output/woodpix_1000m_df_EW.RDS")
#read
#woodpix <- readRDS("C:/Users/ik929086/Documents/iDeer/output/woodpix_1000m_df_EW.RDS")

#EXTRACT WOODLAND FOOD QUALITY

foodpix <- extract_raster(map = woodland_food_value_raster,
                          lcm = template_rast,
                          tiles = tiles)

#EXTRACT WOODLAND SHELTER QUALITY

shelterpix <- extract_raster(map = woodland_shelter_value_raster,
                             lcm = template_rast,
                             tiles = tiles)

#EXTRACT summed DAMS
damspix <- extract_raster(map = polys_merged_max_summed_DAMS_raster_normalised,
                          lcm = template_rast,
                          tiles = tiles)

#saveRDS(damspix, "C:/Users/ik929086/Documents/iDeer/output/dams_1000m_df_EW.RDS")
#read
#damspix <- readRDS("C:/Users/ik929086/Documents/iDeer/output/dams_1000m_df_EW.RDS")

# #EXTRACT nearest urban
# 
# urbanpix <- extract_raster(map = NEAREST_URB_SUBURB,
#                           lcm = template_rast,
#                          tiles = tiles)

#saveRDS(urbanpix, "C:/Users/ik929086/Documents/iDeer/output/urbanpix_df_EW.RDS")
#read
#urbanpix <- readRDS("C:/Users/ik929086/Documents/iDeer/output/urbanpix_df_EW.RDS")

#EXTRACT PERENNIAL COVER

perenpix <- extract_raster(map = Focal1000_PEREN_PERC,
                         lcm = template_rast,
                         tiles = tiles)

#saveRDS(perenpix, "C:/Users/ik929086/Documents/iDeer/output/perenpix_1000m_df_EW.RDS")
#read
#perenpix <- readRDS("C:/Users/ik929086/Documents/iDeer/output/perenpix_1000m_df_EW.RDS")

#EXTRACT ARABLE AREA WITHIN 1000M

arablepix <- extract_raster(map = Focal1000_ARABLE_PERC,
                           lcm = template_rast,
                         tiles = tiles)
#saveRDS(arablepix, "C:/Users/ik929086/Documents/iDeer/output/arablepix_1000m_df_EW.RDS")
#read
#arablepix <- readRDS("C:/Users/ik929086/Documents/iDeer/output/arablepix_1000m_df_EW.RDS")

#left_join the datasets together #---------------

df <- left_join(foodpix, damspix, by = c("x","y")) %>%
  left_join(shelterpix, by = c("x","y")) %>%
  left_join(perenpix, by = c("x","y")) %>%
  #left_join(urbanpix, by = c("x","y")) %>%
  left_join(arablepix, by = c("x","y"))

df <- df %>% rename(Focal1000_MAX_DAMS = polys_merged_max_summed_DAMS_raster_normalised)
               
#make any NAs zeros apart DAMS, woodland cover and urban distance
#these contain correct pixel IDs
df.no.nas <- df %>%
  mutate(across(c(Focal1000_ARABLE_PERC,
                  Focal1000_PEREN_PERC),
                ~ tidyr::replace_na(., 0)))

#Remove the NAs

df.no.nas <- na.omit(df.no.nas)
df.no.nas.full <- df.no.nas

#save

#saveRDS(df.no.nas, "C:/Users/ik929086/Documents/iDeer/output/large_deer_full_df_no_nas.RDS")

# #read
# df.no.nas <- readRDS("C:/Users/ik929086/Documents/iDeer/output/large_deer_full_df_no_nas.RDS")
# 
#  #FOR TESTING: TAKE SUBSET
#  df.no.nas.full <- df.no.nas
#    df.no.nas <- df.no.nas.full %>%
#      slice(1:100000) %>%
#      st_as_sf(coords = c("x","y"))
   
#beepr::beep()

```

```{r}
# #calculate the optimum breakpoints using Jenks natural breaks optimization.
# #to convert variables into categorical
# 
# #DON'T NEED TO DO THIS IN R
# #CAN LOOK AT FULL RASTERS IN ARCGIS PRO, MUCH QUICKER
# 
# #k = number of breaks = 3
# #need to introduce subsets - dataset too large
# 
# #Nearest urban/suburban
# #Predetermined using literature
# #Low <=100  metres
# #Medium >=100 - 250 metres
# #High > 250 metres
# hist(df.no.nas.full$NEAREST_URB_SUBURB)
# abline(v = 100, col = "red")
# abline(v = 250, col = "red")
# 
# #Summed DAMS
# #Need to bring thresholds up so only the super high DAMS scores score as high
# dams_tert <- quantile(df.no.nas.full$Focal1000_MAX_DAMS, c(0:3/3))
# hist(df.no.nas.full$Focal1000_MAX_DAMS)
# abline(v = 0.60, col = "red")
# abline(v = 0.80, col = "red")
# 
# #Connectivity
# #connect_tert <- quantile(df.no.nas.full$Focal1000_WOODS_PERC, c(0:3/3))
# #hist(df.no.nas.full$Focal1000_WOODS_PERC)
# #abline(v = 20, col = "red")
# #abline(v = 40, col = "red")
# 
# #Edge density
# edge_Tert = quantile(df.no.nas.full$Focal1000_EDGE_AREA_PERC, c(0:3/3))
# hist(df.no.nas.full$Focal1000_EDGES_PERC)
# abline(v =16, col = "red")
# abline(v = 30, col = "red")
# 
# #Perennial cover
# peren_tert <- quantile(df.no.nas.full$Focal1000_PEREN_PERC, c(0:3/3))
# hist(df.no.nas.full$Focal1000_PEREN_PERC)
# abline(v = 25, col = "red")
# abline(v = 50, col = "red")
# 
# #Arable cover
# arable_tert <- quantile(df.no.nas.full$Focal1000_ARABLE_PERC, c(0:3/3))
# hist(df.no.nas.full$Focal1000_ARABLE_PERC)
# abline(v = 0, col = "red")
# abline(v = 15, col = "red")
# 
# #Woodland food value
# #Medium = 1
# #High = 2
```

```{r}
#Reclassify continuous data into pre-specified Low,Medium,High categories using breaks from JenksBreaks

df_cat <- df.no.nas.full 

df_cat <- df_cat %>% rename(Woodland_food_value = woodland_food_value_raster,
                            Woodland_shelter_value = woodland_shelter_value_raster)

# #Edges within 1000m
# #LOW/MED/HIGH
# df_cat$Focal1000_EDGES_PERC<- ifelse(df_cat$Focal1000_EDGES_PERC <= 5, "Low",
#           ifelse(df_cat$Focal1000_EDGES_PERC > 5 & df_cat$Focal1000_EDGES_PERC <=30, "Medium",
#           ifelse(df_cat$Focal1000_EDGES_PERC>30, "High", NA)))#)

 
# #Percentage woodland cover within 1000m 
# #LOW/MED/HIGH
#  df_cat$Focal1000_WOODS_PERC<- ifelse(
#   df_cat$Focal1000_WOODS_PERC<= 20, "Low",
#   ifelse(df_cat$Focal1000_WOODS_PERC> 20 & df_cat$Focal1000_WOODS_PERC<= 40, "Medium",
#     ifelse(df_cat$Focal1000_WOODS_PERC> 40, "High", NA)))#)


# Woodland forage quality
df_cat$Woodland_food_value <-
         ifelse(df_cat$Woodland_food_value==1, "Low",
            ifelse(df_cat$Woodland_food_value==2, "Medium",
            ifelse(df_cat$Woodland_food_value==3 , "High", NA)))

# Woodland shelter quality
df_cat$Woodland_shelter_value <-
            ifelse(df_cat$Woodland_shelter_value==1, "Low",
            ifelse(df_cat$Woodland_shelter_value==2, "Medium",
            ifelse(df_cat$Woodland_shelter_value==3 , "High", NA)))


#Arable within 1000m
#LOW/MED/HIGH
#Manually adjusted
df_cat$Focal1000_ARABLE_PERC <- ifelse(df_cat$Focal1000_ARABLE_PERC <=25, "Low",
           ifelse(df_cat$Focal1000_ARABLE_PERC > 25 & df_cat$Focal1000_ARABLE_PERC <=50, "Medium",
           ifelse(df_cat$Focal1000_ARABLE_PERC >50, "High", NA)))

#Perennial cover within 1000m
df_cat$Focal1000_PEREN_PERC <- ifelse(df_cat$Focal1000_PEREN_PERC <=25, "Low",
          ifelse(df_cat$Focal1000_PEREN_PERC > 25 & df_cat$Focal1000_PEREN_PERC <=50, "Medium",
          ifelse(df_cat$Focal1000_PEREN_PERC >50, "High", NA)))#)

#Max summed DAMS within 1000m
df_cat$Focal1000_MAX_DAMS <- ifelse(df_cat$Focal1000_MAX_DAMS <= 0.60, "Low",
                      ifelse(df_cat$Focal1000_MAX_DAMS > 0.60  & df_cat$Focal1000_MAX_DAMS <=0.80, "Medium",
                      ifelse(df_cat$Focal1000_MAX_DAMS > 0.80, "High", NA)))#)

# #Nearest urban/suburban area
# df_cat$NEAREST_URB_SUBURB <- ifelse(df_cat$NEAREST_URB_SUBURB  <= 100, "High",
#                       ifelse(df_cat$NEAREST_URB_SUBURB> 100 & df_cat$NEAREST_URB_SUBURB  <=250, "Medium",
#                       ifelse(df_cat$NEAREST_URB_SUBURB  > 250, "Low", NA)))

#ensure factor levels in correct order 
df_cat$Focal1000_MAX_DAMS <- factor(df_cat$Focal1000_MAX_DAMS, levels = c("Low","Medium","High"))

df_cat$Woodland_food_value<- factor(df_cat$Woodland_food_value, levels= c("Low","Medium","High"))

df_cat$Focal1000_ARABLE_PERC <- factor(df_cat$Focal1000_ARABLE_PERC, levels = c("Low","Medium","High"))

#df_cat$Focal1000_EDGES_PERC<- factor(df_cat$Focal1000_EDGES_PERC, levels = c("Low","Medium","High"))

#df_cat$NEAREST_URB_SUBURB <- factor(df_cat$NEAREST_URB_SUBURB, levels = c("Low","Medium","High"))

df_cat$Focal1000_PEREN_PERC <- factor(df_cat$Focal1000_PEREN_PERC, levels = c("Low","Medium","High"))

df_cat$Woodland_shelter_value <- factor(df_cat$Woodland_shelter_value, levels = c("Low","Medium","High"))

#unique(df_cat$NEAREST_URB_SUBURB)
unique(df_cat$Focal1000_MAX_DAMS)
unique(df_cat$Focal1000_PEREN_PERC)
unique(df_cat$Focal1000_ARABLE_PERC)
#unique(df_cat$Focal1000_EDGES_PERC)
unique(df_cat$Woodland_food_value)
unique(df_cat$Woodland_shelter_value)
```

```{r}
#CPTS for measured nodes

#nearest urban/suburban
#urban_cpt <-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("Low","Medium","High")))
#perennial cover within 1000m
peren_cpt <- matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("Low", "Medium","High")))
#arable area within 1000m
arable_cpt <- matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("Low", "Medium","High")))
#Max summed dams within 1000m
dams_cpt<-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("Low", "Medium","High")))
# #woodland connectivity within 1000m
# wood_connect_cpt <- matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("Low", "Medium","High")))
#woodland value
wood_food_value_cpt <- matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("Low", "Medium","High")))
wood_shelter_value_cpt <- matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("Low", "Medium","High")))
#woodland edge area within 1000m
#edge_cpt <-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("Low", "Medium","High")))
```

```{r}
#CPT for Attraction index ---------------#

attraction_cpt_df <- large_attraction_cpt_df %>%
  #dplyr::select(-c("X","Expertise","Confidence","Total"))
  dplyr::select(-c("X","total_score"))
head(attraction_cpt_df)

# Define the dimensions
attraction_index_levels <- c("Low", "Medium", "High")
peren_area_levels <- unique(attraction_cpt_df$Perennial_cover_1000m)
arable_area_levels <- unique(attraction_cpt_df$Arable_cover_1000m)
#edge_levels <- unique(attraction_cpt_df$Woodland_edge_cover_1000m)

# Initialize the array with dimensions and names
attraction_index_cpt <- array(
  NA,  # Placeholder value
  dim = c(length(attraction_index_levels),
          length(peren_area_levels),
          length(arable_area_levels)),
  dimnames = list(
    attraction_index = attraction_index_levels,
    Focal1000_PEREN_PERC = peren_area_levels,
    #Focal1000_EDGES_PERC = edge_levels,
    Focal1000_ARABLE_PERC = arable_area_levels
  )
)

# Populate the array
for (i in 1:nrow(attraction_cpt_df)) {
  peren_val <- attraction_cpt_df$Perennial_cover_1000m[i]
  #edge_val <- attraction_cpt_df$Woodland_edge_cover_1000m[i]
  arable_val <- attraction_cpt_df$Arable_cover_1000m[i]

  attraction_index_cpt["Low", as.character(peren_val), as.character(arable_val)] <- attraction_cpt_df$Low[i]
  attraction_index_cpt["Medium", as.character(peren_val),as.character(arable_val)] <- attraction_cpt_df$Medium[i]
  attraction_index_cpt["High", as.character(peren_val),as.character(arable_val)] <- attraction_cpt_df$High[i]

}

#attraction_index_cpt <- attraction_index_cpt / 100
attraction_index_cpt
```

```{r}
#CPT for thermoregulation index --------------------------------####

thermoreg_cpt_df<- large_thermoreg_cpt_df %>%
  dplyr::select(-c("X"))

# Define the dimensions
thermoreg_index_levels <- c("Low", "Medium", "High")
max_dams_1000m_levels <- unique(thermoreg_cpt_df$MAX_DAMS_1000m)
woodland_shelter_quality_levels <- unique(thermoreg_cpt_df$woodland_shelter_value)

# Initialize the array
thermoreg_index_cpt <- array(
  NA,
  dim = c(length(thermoreg_index_levels), length(max_dams_1000m_levels), length(woodland_shelter_quality_levels)),
  dimnames = list(
    thermoreg_index = thermoreg_index_levels,
    MAX_DAMS_1000m = max_dams_1000m_levels,
    Woodland_shelter_value = woodland_shelter_quality_levels
  )
)

# Populate the array with values from the DataFrame
for (i in 1:nrow(thermoreg_cpt_df)) {
  dams_val <- thermoreg_cpt_df$MAX_DAMS_1000m[i]
  shelter_val <- thermoreg_cpt_df$woodland_shelter_value[i]

  thermoreg_index_cpt["Low", as.character(dams_val), as.character(shelter_val)] <- thermoreg_cpt_df$Low[i]
  thermoreg_index_cpt["Medium", as.character(dams_val),as.character(shelter_val)] <- thermoreg_cpt_df$Medium[i]
  thermoreg_index_cpt["High", as.character(dams_val),as.character(shelter_val)] <- thermoreg_cpt_df$High[i]
}

thermoreg_index_cpt
```

```{r}
#CPT for woodland food value index ------------------------------------------------####
wood_food_value_cpt_df <- large_wood_food_value_cpt_df %>%
dplyr::select(-c("X","Total","Expertise","Confidence"))  

head(wood_food_value_cpt_df)

# Define the dimensions
wood_value_index_levels <- c("Low","Medium","High")
wood_food_value_levels <- unique(wood_food_value_cpt_df$Woodland_food_value)

# Initialize the array with dimensions and names
wood_value_index_cpt <- array(
  NA,  # Placeholder value
  dim = c(length(wood_value_index_levels), 
          length(wood_food_value_levels)),
  dimnames = list(
    wood_value_index = wood_value_index_levels,
    Woodland_food_value =wood_food_value_levels
  )
)

# Populate the array with values from the DataFrame
for (i in 1:nrow(wood_food_value_cpt_df)) {
  # Get the indices for each dimension
  wood_val_idx <- which(dimnames(wood_value_index_cpt)$Woodland_food_value == wood_food_value_cpt_df$Woodland_food_value[i])
  
  # Assign values for each forage pressure index level
    wood_value_index_cpt["Low", wood_val_idx] <- wood_food_value_cpt_df$Low[i]
  wood_value_index_cpt["Medium", wood_val_idx] <- wood_food_value_cpt_df$Medium[i]
  wood_value_index_cpt["High",wood_val_idx] <- wood_food_value_cpt_df$High[i]
}
#wood_value_index_cpt <- wood_value_index_cpt / 100
wood_value_index_cpt 
```

```{r}
#Final CPT: deer damage risk -----------------------------------####

risk_cpt_df<- large_risk_cpt_df %>%
  dplyr::select(-c("X","total_score")) %>%
  rename(attraction_index=Attraction_Index,
         wood_value_index=Woodland_Food_Value_Index,
         thermoreg_index=Thermoregulation_Index)#,
         #urban_dist_index= Urban_Distance_Index)

# Replace '..' with '_' in column names
#colnames(risk_cpt_df) <- gsub("\\.\\.", "_", colnames(risk_cpt_df))
  
head(risk_cpt_df)

# Define the dimensions
large_risk_index_levels <- c("Low", "Low_Medium" ,"Medium", "Medium_High" , "High")
wood_value_index_levels <- unique(risk_cpt_df$wood_value_index)
attraction_index_levels <- unique(risk_cpt_df$attraction_index)
thermoreg_index_levels <- unique(risk_cpt_df$thermoreg_index)
#urban_dist_levels <- unique(risk_cpt_df$urban_dist_index)

# Initialize the array with dimensions and names
large_risk_index_cpt <- array(
  NA,  # Placeholder value
  dim = c(length(large_risk_index_levels), 
          length(wood_value_index_levels), 
          length(attraction_index_levels), 
          length(thermoreg_index_levels)),
  dimnames = list(
    large_risk_index = large_risk_index_levels,
    wood_value_index = wood_value_index_levels,
    attraction_index = attraction_index_levels,
    thermoreg_index = thermoreg_index_levels
  )
)

# Populate the array with values from the DataFrame
for (i in 1:nrow(risk_cpt_df)) {
  wood_food_val <- risk_cpt_df$wood_value_index[i]
  attract_val <- risk_cpt_df$attraction_index[i]
  thermoreg_val <- risk_cpt_df$thermoreg_index[i]
  #urban_dist_val <- risk_cpt_df$urban_dist_index[i]

  large_risk_index_cpt["Low", as.character(wood_food_val), as.character(attract_val), as.character(thermoreg_val)] <- risk_cpt_df$Low[i]
  large_risk_index_cpt["Low_Medium", as.character(wood_food_val), as.character(attract_val), as.character(thermoreg_val)] <- risk_cpt_df$Low_Medium[i]
  large_risk_index_cpt["Medium", as.character(wood_food_val), as.character(attract_val), as.character(thermoreg_val)] <- risk_cpt_df$Medium[i]
  large_risk_index_cpt["Medium_High", as.character(wood_food_val), as.character(attract_val), as.character(thermoreg_val)] <- risk_cpt_df$Medium_High[i]
  large_risk_index_cpt["High", as.character(wood_food_val), as.character(attract_val), as.character(thermoreg_val)] <- risk_cpt_df$High[i]
}
large_risk_index_cpt 
```

```{r}
#Create BBN structure#------------------------------------

# Step 1: Explicitly define the nodes in the network
nodes <- c(
  "Woodland_food_value",
  "Woodland_shelter_value",
  #"dist_to_urban",
  "Focal1000_PEREN_PERC",
  "Focal1000_ARABLE_PERC",
  "MAX_DAMS_1000m",
  #"Focal1000_EDGES_PERC",
  "attraction_index",
  "thermoreg_index",
  "wood_value_index",
  #"urban_dist_index",
  "large_risk_index"
)

#empty graph
library(bnlearn)
e = empty.graph(nodes)

arc.set = matrix(c("Woodland_food_value", "wood_value_index",
                   "Focal1000_PEREN_PERC", "attraction_index",
                   "Focal1000_ARABLE_PERC", "attraction_index",
                   "MAX_DAMS_1000m","thermoreg_index",
                   "Woodland_shelter_value","thermoreg_index",
                   "thermoreg_index","large_risk_index",
                   "wood_value_index","large_risk_index",
                   "attraction_index","large_risk_index"),
                 ncol = 2, byrow = TRUE,
                 dimnames = list(NULL, c("from", "to")))

arcs(e) <- arc.set

model_string <- modelstring(e) 

net<-model2network(model_string) 
# You dont need to repeat the names of the nodes with parents (child nodes) 

# Custom fitting network (matching up the nodes to their CPTs)
dfit = custom.fit(net, dist = list(Woodland_food_value = wood_food_value_cpt, 
                                   Focal1000_PEREN_PERC = peren_cpt,
                                   MAX_DAMS_1000m = dams_cpt,
                                   Focal1000_ARABLE_PERC = arable_cpt,
                                   Woodland_shelter_value = wood_shelter_value_cpt,
                                   thermoreg_index = thermoreg_index_cpt,
                                   attraction_index =attraction_index_cpt,
                                   large_risk_index=large_risk_index_cpt,
                                   wood_value_index=wood_value_index_cpt))

#Plot BN structure ####
#graphviz.plot(net)

#Check parameters ####
dfit
```

```{r}
predict_dat <- df_cat %>%st_drop_geometry() #%>% #drop geometry
  #mutate(across(where(is.character), toupper)) %>% #convert characters to upper case
  #dplyr::select(-c("x","y"))

#Ensure variable names match those in BBN
predict_dat<-predict_dat%>%rename(
  Focal1000_ARABLE_PERC=Focal1000_ARABLE_PERC,
  Focal1000_PEREN_PERC=Focal1000_PEREN_PERC,
  MAX_DAMS_1000m=Focal1000_MAX_DAMS,
  Woodland_food_value=Woodland_food_value)#,
  #Focal1000_EDGES_PERC = Focal1000_EDGES_PERC)

#Make empty columns for the latent (unobserved) variables

predict_dat$attraction_index <- NA
predict_dat$wood_value_index <- NA
predict_dat$thermoreg_index <- NA
predict_dat$large_risk <- NA

head(predict_dat)

#Ensure all columns are factors, not characters
predict_dat <- predict_dat %>% mutate_all(as.factor)

#Ensure factor levels are in correct order
levels(predict_dat$Woodland_food_value)<-c("Low", "Medium", "High")
levels(predict_dat$Focal1000_PEREN_PERC)<-c("Low", "Medium", "High")
levels(predict_dat$Focal1000_ARABLE_PERC)<-c("Low", "Medium", "High")
levels(predict_dat$MAX_DAMS_1000m)<-c("Low", "Medium", "High")
#levels(predict_dat$Focal1000_EDGES_PERC)<-c("Low", "Medium", "High")
levels(predict_dat$Woodland_shelter_value) <- c("Low", "Medium", "High")

predict_dat$attraction_index <- factor(NA, levels = c("Low", "Medium", "High"))
predict_dat$wood_value_index <- factor(NA, levels = c("Low","Medium", "High"))
predict_dat$thermoreg_index <- factor(NA, levels = c("Low", "Medium", "High"))
predict_dat$large_risk <- factor(NA, levels = c("Low", "Low_Medium", "Medium", "Medium_High", "High"))


all_nodes <- names(dfit)

#PREDICTIONS TAKE A LONG TIME, ABOUT 3 MINUTES PER LINE
#Predict values for latent variables
pred_thermoreg = predict(object=dfit,node="thermoreg_index",data=predict_dat, method = "parents")
pred_fpi = predict(dfit,node="attraction_index",data=predict_dat, method = "parents")
pred_wood_value = predict(dfit,node="wood_value_index",data=predict_dat, method = "parents")

#fill the columns
predict_dat$thermoreg_index <- pred_thermoreg
predict_dat$attraction_index <- pred_fpi
predict_dat$wood_value_index <- pred_wood_value

#Predict damage
pred_damage = predict(dfit,node = "large_risk_index",data = predict_dat,method = "parents")
predict_dat$large_risk <- pred_damage

# # scatter plot of forage pressure index and alt forage quality
# predict_dat_sample <- predict_dat[sample(nrow(predict_dat), 10000), ]
# ggplot(predict_dat_sample, aes(x = alt_forage_qual_1000m, y = attraction_index)) +
#   geom_jitter(width = 0.2, height = 0.2, size = 3, color = "blue") +
#   labs(x = "Alt forage Q", y = "Index") +
#   theme_minimal()

#Add damage risk back into spatial dataset

df_cat$pred_damage <- pred_damage
df_cat$pred_thermoreg <- pred_thermoreg
df_cat$pred_attraction_index <- pred_fpi
df_cat$pred_wood_value_index <- pred_wood_value

beepr::beep()
```

```{r}
#Make risk raster #------------------------

#par(mfrow = c(1, 1))

 # Make sf object
 df_sf <- st_as_sf(df_cat, coords = c("x", "y"), crs = st_crs(bng))
 
 #save sf object
 #saveRDS(df_sf, "C:/Users/ik929086/Documents/iDeer/output/large_deer_full_df_cat_point_geometry.RDS")
 
#df_sf <- readRDS("C:/Users/ik929086/Documents/iDeer/output/large_deer_full_df_cat_point_geometry.RDS")
 
#Make large deer predict_dat rasters to check layers
# Define the color palette

#Forage pressure index
forage_vals <- c("Low"=1,"Medium"=2,"High"=3)
df_sf$forage_vals <- forage_vals[df_sf$pred_attraction_index]
df_sf_selected <- df_sf%>%dplyr::select(c(geometry, forage_vals))
#forage_r <- stars::st_rasterize(sf = df_sf_selected, template = stars_template)
forage_r <- stars::st_rasterize(sf = df_sf_selected)
#stars::write_stars(forage_r, "C:/Users/ik929086/Documents/iDeer/output/large_deer_attraction_index_rast_test.tif",overwrite=TRUE)
forage_r[forage_r == 0] <- NA #make zero's NAs

unique_vals <- sort(na.omit(unique(as.vector(forage_r[[1]]))))

# Define full palette
forage_palette <- c("1" = "green", "2"= "orange", "3" = "red")

# Subset palette to only existing levels in the raster
active_palette <- forage_palette[names(forage_palette) %in% unique_vals]

# Set breaks: one more than number of colours
breaks_vals <- c(min(unique_vals) - 0.5, unique_vals + 0.5)

# Plot with explicit breaks and colours
plot(forage_r, col = active_palette, breaks = breaks_vals)

#save

#stars::write_stars(forage_r, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_attraction_index_example_borrowdale.tif")

#plot(forage_r)
# 
# #Connectivity index
# connect_vals <- c("Low"=1,"Medium"=2,"High"=3)
# df_sf$connect_vals <- connect_vals[df_sf$pred_connect]
# df_sf_selected <- df_sf%>%dplyr::select(c(geometry, connect_vals))
# #forage_r <- stars::st_rasterize(sf = df_sf_selected, template = stars_template)
# connect_r <- stars::st_rasterize(sf = df_sf_selected)
# #stars::write_stars(connect_r, "C:/Users/ik929086/Documents/iDeer/output/large_deer_connectivity_index_rast_test.tif",overwrite=TRUE)
# plot(connect_r, col = index_palette)

wood_value_palette <- c("1" = "green", "2"= "orange", "3" = "red")
#Wood value index
wood_value_vals <- c("Low"=1,"Medium"=2,"High"=3)
df_sf$wood_value_vals <- wood_value_vals[df_sf$pred_wood_value_index]
df_sf_selected <- df_sf%>%dplyr::select(geometry, wood_value_vals)
#forage_r <- stars::st_rasterize(sf = df_sf_selected, template = stars_template)
wood_value_r <- stars::st_rasterize(sf = df_sf_selected)
wood_value_r[wood_value_r == 0] <- NA #make zero's NAs
plot(wood_value_r, col = wood_value_palette)
unique_vals <- unique(as.vector(wood_value_r[[1]]))


#Thermoreg index

thermoreg_vals <- c("Low"=1,"Medium"=2,"High"=3)
 df_sf$thermoreg_vals <- thermoreg_vals[df_sf$pred_thermoreg]
 df_sf_selected <- df_sf%>%dplyr::select(geometry, thermoreg_vals)
 #forage_r <- stars::st_rasterize(sf = df_sf_selected, template = stars_template)
 thermoreg_r <- stars::st_rasterize(sf = df_sf_selected)
thermoreg_r[thermoreg_r == 0] <- NA #make zero's NAs
 # Define full palette
full_palette <- c("1" = "green", "2" = "orange", "3" = "red")

#Get unique values
unique_vals <- sort(na.omit(unique(as.vector(thermoreg_r[[1]]))))

# Subset palette to only existing levels in the raster
active_palette <- full_palette[names(full_palette) %in% unique_vals]

 #stars::write_stars(connect_r, "C:/Users/ik929086/Documents/iDeer/output/large_deer_connectivity_index_rast_test.tif",overwrite=TRUE)
plot(thermoreg_r, col = active_palette)
#stars::write_stars(thermoreg_r, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_thermoreg_index_example_borrowdale.tif")
# 
#  #Disturbance index
# disturbance_vals <- c("Low"=1,"Medium"=2,"High"=3)
# df_sf$disturbance_vals <- disturbance_vals[df_sf$pred_disturb]
# df_sf_selected <- df_sf%>%dplyr::select(geometry, disturbance_vals)
# disturbance_r <- stars::st_rasterize(sf = df_sf_selected)
# disturbance_r[disturbance_r == 0] <- NA #make zero's NAs
# 
#  plot(disturbance_r, col = full_palette)
# 
 # Define the value mapping for pred_damage
damage_values <- c("Low" = 1, "Low_Medium" = 2, "Medium" = 3, "Medium_High" = 4, "High" = 5)
df_sf$value <- damage_values[df_sf$pred_damage]
damage_palette <- c("1" = "green",
                    "2" = "yellow",
                    "3" = "orange",
                    "4" = "red" ,
                    "5" = "firebrick4")

 #Convert to raster
 # Select the geometry and value column
 df_sf_selected <- df_sf %>% dplyr::select(geometry, value)
 #make a stars template
 # Convert the RasterLayer to a stars object
# stars_template <- stars::st_as_stars(lcm)
 # Use the template raster for extent, resolution, and CRS
 r <- stars::st_rasterize(sf = df_sf_selected)
 #r <- stars::st_rasterize(sf = df_sf_selected)
 r[r == 0] <- NA #make zero's NAs
  # Check unique values
unique_vals <- unique(as.vector(r[[1]]))
#print(unique_vals)

# Define full palette
risk_palette <- c("1" = "green", "2" = "yellow", "3" = "orange", "4" = "red", "5" = "brown")

# Subset palette to only existing levels in the raster
active_palette <- risk_palette[names(risk_palette) %in% unique_vals]

plot(r, col = active_palette)
#stars::write_stars(r, "C:/Users/ik929086/Documents/iDeer/output/GB_datasets_2023/large_deer/NFI_large_deer_risk_raster_example_borrowdale.tif",overwrite=TRUE)
#beepr::beep()

#r <- raster("C:/Users/ik929086/Documents/iDeer/output/large_deer_risk_raster.tif")
#r[is.na(r)] <- 0
```

```{r}
#Extract to polygons

r_rast <- as(r, "Raster")

woods_only <- nfi %>% filter(CATEGORY %in% c("Woodland"))
unique(woods_only$IFT_IOA)
site_woods_unmerged <- st_filter(woods_only, tiles)
site_woods_unmerged <- site_woods_unmerged %>% rename(patch_ID = OBJECTID)

#function to extract risk maps to polygons
source("C:/Users/ik929086/Documents/iDeer-tool/code/functions/extract_raster_pixels_to_wood_polys_func_site_level.R")

large_deer_risk_poly_vals <- extract_raster_to_wood_polygons(risk_map = r_rast,
                                                             site_woods_unmerged = site_woods_unmerged
                                                             )
# Get polygon IDs
poly_ids <- unique(large_deer_risk_poly_vals$patch_ID)

# # Group by patch_ID and calculate the counts for each raster_layer category
deer_risk_landcover <- large_deer_risk_poly_vals %>%
  filter(!is.na(r_rast)) %>% # Remove NA values
  group_by(patch_ID) %>%
  summarise(
    risk_count_1 = sum(r_rast == 1, na.rm = TRUE),
    risk_count_2 = sum(r_rast == 2, na.rm = TRUE),
    risk_count_3 = sum(r_rast == 3, na.rm = TRUE),
    risk_count_4 = sum(r_rast == 4, na.rm = TRUE),
    risk_count_5 = sum(r_rast == 5, na.rm = TRUE)
  )

# # Combine results into a matrix, where each column corresponds to a polygon
deer_risk_landcover_matrix <- do.call(rbind, deer_risk_landcover)
#make colnames patch_IDs
colnames(deer_risk_landcover_matrix) <- deer_risk_landcover_matrix[1,]

#Add polygon IDs to the matrix
deer_risk_landcover_df <- data.frame(t(deer_risk_landcover_matrix))

#Convert the matrix to a long format
deer_risk_landcover_long <- reshape2::melt(
  deer_risk_landcover_df,
  id.vars = "patch_ID",  # Specify the ID variable(s)
  variable.name = "risk_category",  # Name for the risk category column
  value.name = "value"  # Name for the value column
)

mean_risk <- deer_risk_landcover_long %>%
  mutate(
    risk_category = as.numeric(gsub("[^0-9]", "", risk_category))  # Remove non-numeric characters and convert to numeric
  ) %>%
  group_by(patch_ID) %>%
  filter(value != 0) %>%  # Exclude rows with zero values
  summarise(
    Weighted_Mean = sum(risk_category * value) / sum(value),
    #Weighted SD calculation
    sd = sqrt(sum(value * (risk_category - Weighted_Mean)^2) / sum(value)),
    # Weighted SE calculation
    se = sd / sqrt(sum(value)),
    .groups = 'drop'
  )

#Append the mean risk categories to the polygon dataset
polygon_with_mean_risk <- site_woods_unmerged %>%
  left_join(mean_risk, by = "patch_ID")

#Plot the risk
ggplot() +
  geom_sf(data = polygon_with_mean_risk, mapping = aes(fill = Weighted_Mean), color = "black", linewidth = 0.1) +
  scale_fill_gradientn(
  colours = c("green", "yellow", "orange", "red", "brown"),
  limits = c(1,5),
  breaks = 1:5,
  name = "Damage Risk"
) +
  theme_minimal()

#save map

st_write(polygon_with_mean_risk, "C:/Users/ik929086/Documents/iDeer/output/Testing/large_deer_risk_map_test_new_forest.shp",append = TRUE)

```
