---
title: "Chloe's extraction of data"
output: html_notebook
---

#set up, read libraries etc

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
.libPaths("C:/R-Packages")#set library path
root.dir = "C:/LocalData/NFI/iDeer" #set home directory
#setwd(root.dir) # or setwd("C:/LocalData/NFI/#####")

#Directories - create outputs and scratch folders if they don't already exist in the root directory
#ifelse(!dir.exists(file.path(root.dir, "Scratch")), dir.create(file.path(root.dir, "Scratch")), FALSE)
#ifelse(!dir.exists(file.path(root.dir, "Outputs")), dir.create(file.path(root.dir, "Outputs")), FALSE)

```

```{r}
library(pacman)
#Install or load required packages
pacman::p_load(tidyverse, dplyr, plyr, sp,  sf, terra)
```



```{r}
#Set coordinate system
crs = 27700 #coordinate system EPSG code (British National Grid = epsg:27700) 
crs2 = "+init=epsg:27700"#coordinate system (British National Grid = "+init=epsg:27700")
```

Read in the NFI sections and select the columns we want. Also filter by year, as we are only working with the first wave, pre-12016.
```{r}

#Read in the sections
#sections = sf::st_read(paste0("Inputs/NFI_sections.rds"), crs = crs)

sections = readRDS("C:/LocalData/NFI/iDeer/Data/Inputs/NFI_sections.rds")
st_crs(sections, crs)


sections <- sections%>%
  unite(SQUID_SECTION, SQUID,SECTION_LETTER.x, sep="_", remove = F)%>%#make unique id
 # mutate(YEAR= substring(ASSESSMENT,1,4))%>%
  mutate(SECTION_AREA= SHAPE_Area)%>%
 # filter(YEAR<2016)%>% #don't have year info, but shoudl all be first cycle anyway
  dplyr::select(SQUID_SECTION, SQUID,SECTION_AREA)#removed year


sections$nparts <- unlist(lapply(st_geometry(sf::st_as_sf(sections)), length))
hist(sections$nparts)

sections <- sections %>% filter(nparts<2) # only do sections that are spatially contiguous (nparts=1), else the damage assessment might be a bit wierd?
sections$nparts <- NULL # remove column





#rEAD IN AND Join local data
NFILocal = read.csv("C:/LocalData/NFI/iDeer/Data/Inputs/All_NFILocal_GIS_SectionData.csv")
Damage = read.csv("C:/LocalData/NFI/iDeer/Data/Inputs/Browsing_Stripping.csv")

Damage <- Damage%>%
  unite(SQUID_SECTION, SQUID,SECTION_LETTER, sep="_", remove = F)

NFILocalDeer = merge(NFILocal,Damage, by = "SQUID_SECTION", all.x = TRUE)


#Add local survey data to shapefile
Sections_LocalData <- merge(sections,NFILocalDeer, by = "SQUID_SECTION", all.x = FALSE)


st_write(sections , "C:/LocalData/NFI/iDeer/Data/Inputs/iDeerSectionsInput.shp", driver = "ESRI Shapefile")  # create to a shapefile 


```

#Reduce the number of sections

Remove Scotland sections, as we are only focusing on Eng and Wales
And to save computing time, remove sections with weird IFT categories
Perhaps also a minumum size of section
For the buffer project, Will Peden outputted a load of .csv files
We can use those to subset sections

```{r}

#remove scotland sections
#scot = sf::st_read("C:/Users/ij926050/Dropbox/BEAST_R/Outputs/SCOT_SECTs.shp", crs = crs)
#sections <- sections %>% filter(!SQUID%in%scot$SQUID)
Sections_LocalData = Sections_LocalData %>%
  filter(!SCOTLAND=='t')

plot(st_geometry(Sections_LocalData))#check distribution

```


```{r}

ift_cats <- read.csv ("C:/LocalData/NFI/iDeer/Data/Inputs/IFT_IOA_class.csv")

ift_cats <- ift_cats%>%
  unite(SQUID_SECTION, SQUID,SECTION_LETTER, sep="_", remove = F)

ift_cats <- ift_cats %>% filter (Conifer ==1 | Broadleaved ==1 | Mixed_Predom_Broadleaf ==1 | Mixed_Predom_Conifer == 1)


Sections_LocalData <- Sections_LocalData%>% filter(SQUID_SECTION %in% ift_cats$SQUID_SECTION)

#a minimum size? best not remove!

hist(Sections_LocalData$SECTION_AREA.x)

plot(st_geometry(Sections_LocalData))
```
DAMS extraction
```{r}
#DAMs extraction
#These are focal statistics rasters for three buffer sizes, 250, 500 and 1000

dams_file_path <- "C:/LocalData/NFI/iDeer/Data/Inputs/DAMS"

rasterlist <- list.files(path = dams_file_path, pattern='.tif',
                       all.files=TRUE, full.names=TRUE)


rasterNames=sub('.*/', '', rasterlist)#remove file pathways from list
rasterNames=sub('*.tiff', '', rasterNames)#remove extension from list

all_rasters <- terra::rast(rasterlist)#Create SpatRaster from list


names(all_rasters) = rasterNames#rename the rasters in the raster stack by the raster file names

library(terra)


DAMs_ectract <- terra::extract(all_rasters, vect(Sections_LocalData), fun = mean, , na.rm=TRUE)  # returns a DF
DAMS_extract<- cbind(Sections_LocalData, DAMs_ectract)
write.csv(st_drop_geometry(DAMS_extract), "C:/LocalData/NFI/iDeer/Data/Outputs/DAMS_NFI_extract.csv") # save (maybe change the filepath?)
```





landscape metrics (donut)
```{r}
lc <- raster::raster("C:/LocalData/NFI/iDeer/Data/Inputs/edge/edge_core_raster_all_tiles_EW.tif")

plot(lc)

POLY=as(Sections_LocalData, "Spatial")
veg = lc
#ext<-raster::extract(lc, POLY, method='simple')
#class(ext)  # a list
## [1] "list"
#length(ext) # three elements, a vector of land use values for each region
## [1] 3

polyFragStats <- function(POLY, veg) {
  # Create a buffer around the polygon
  buffer_poly <- buffer(POLY, width = 250, dissolve = TRUE)
  
  # Cut out the original polygon from the buffer
  buffer_area <- erase(buffer_poly, POLY)
  
  # Crop land cover map by the buffer area
  fcrop <- crop(veg, buffer_area)
  
  # Mask the cropped land cover map
  fmask <- mask(fcrop, buffer_area)
  
  # Extract class statistics
  res <- raster::extract(fmask, buffer_area, method = "simple")
  
  # Convert the results to a data frame and calculate statistics
  res <- as.data.frame(table(res))
  res$SQUID_SECTION <- POLY$SQUID_SECTION
  
  res <- res %>%
    mutate(total_pix = as.numeric(sum(Freq)),
           class = res) %>%
    dplyr::select(SQUID_SECTION, class, Freq, total_pix) %>%
    mutate(perc_cov = Freq * 100 / total_pix, buff_area = area(buffer_area))
  
  return(res)
}


respoly <- ldply(1:length(POLY), function(i) {  # for each point
  out <- polyFragStats(POLY[i,],  veg=veg) # compute fragstats metrics 
  out$buffer_size <- "b150"
  return(out)
})  

respoly

write.csv(respoly, "C:/LocalData/NFI/iDeer/Data/Outputs/landscapemetrics_250.csv")
```




```{r}

#ext<-raster::extract(lc, POLY, method='simple')
#class(ext)  # a list
## [1] "list"
#length(ext) # three elements, a vector of land use values for each region
## [1] 3

polyFragStats <- function(POLY, veg) {
  # Create a buffer around the polygon
  buffer_poly <- buffer(POLY, width = 500, dissolve = TRUE)
  
  # Cut out the original polygon from the buffer
  buffer_area <- erase(buffer_poly, POLY)
  
  # Crop land cover map by the buffer area
  fcrop <- crop(veg, buffer_area)
  
  # Mask the cropped land cover map
  fmask <- mask(fcrop, buffer_area)
  
  # Extract class statistics
  res <- raster::extract(fmask, buffer_area, method = "simple")
  
  # Convert the results to a data frame and calculate statistics
  res <- as.data.frame(table(res))
  res$SQUID_SECTION <- POLY$SQUID_SECTION
  
  res <- res %>%
    mutate(total_pix = as.numeric(sum(Freq)),
           class = res) %>%
    dplyr::select(SQUID_SECTION, class, Freq, total_pix) %>%
    mutate(perc_cov = Freq * 100 / total_pix, buff_area = area(buffer_area))
  
  return(res)
}


respoly <- ldply(1:length(POLY), function(i) {  # for each point
  out <- polyFragStats(POLY[i,],  veg=veg) # compute fragstats metrics 
  out$buffer_size <- "b500"
  return(out)
})  

respoly500 <- respoly
write.csv(respoly500, "C:/LocalData/NFI/iDeer/Data/Outputs/landscapemetrics_500.csv")
```





```{r}
polyFragStats <- function(POLY, veg) {
  # Create a buffer around the polygon
  buffer_poly <- buffer(POLY, width = 1000, dissolve = TRUE)
  
  # Cut out the original polygon from the buffer
  buffer_area <- erase(buffer_poly, POLY)
  
  # Crop land cover map by the buffer area
  fcrop <- crop(veg, buffer_area)
  
  # Mask the cropped land cover map
  fmask <- mask(fcrop, buffer_area)
  
  # Extract class statistics
  res <- raster::extract(fmask, buffer_area, method = "simple")
  
  # Convert the results to a data frame and calculate statistics
  res <- as.data.frame(table(res))
  res$SQUID_SECTION <- POLY$SQUID_SECTION
  
  res <- res %>%
    mutate(total_pix = as.numeric(sum(Freq)),
           class = res) %>%
    dplyr::select(SQUID_SECTION, class, Freq, total_pix) %>%
    mutate(perc_cov = Freq * 100 / total_pix, buff_area = area(buffer_area))
  
  return(res)
}


respoly <- ldply(1:length(POLY), function(i) {  # for each point
  out <- polyFragStats(POLY[i,],  veg=veg) # compute fragstats metrics 
  out$buffer_size <- "b1000"
  return(out)
})  

respoly

write.csv(respoly, "C:/LocalData/NFI/iDeer/Data/Outputs/landscapemetrics_1000.csv")
```







Nearest urban/suburban features extraction

```{r}
#These are nearest_dist rasters

urban_file_path <- "C:/LocalData/NFI/iDeer/Data/Inputs/Urban"

rasterlist <- list.files(path = urban_file_path, pattern='.tif',
                       all.files=TRUE, full.names=TRUE)


rasterNames=sub('.*/', '', rasterlist)#remove file pathways from list
rasterNames=sub('*.tiff', '', rasterNames)#remove extension from list

all_rasters <- terra::rast(rasterlist)#Create SpatRaster from list


names(all_rasters) = rasterNames#rename the rasters in the raster stack by the raster file names

urban_ectract <- terra::extract(all_rasters, vect(Sections_LocalData), fun = mean, na.rm=TRUE)  # returns a DF 
urban_extract<- cbind(Sections_LocalData, urban_ectract)
write.csv(st_drop_geometry(urban_extract), "C:/LocalData/NFI/iDeer/Data/Outputs/urban_NFI_extract.csv") # save (maybe change the filepath?)

```


#Roads

```{r}
#These are nearest_dist rasters

road_file_path <- "C:/LocalData/NFI/iDeer/Data/Inputs/roads"

rasterlist <- list.files(path = road_file_path, pattern='.tif',
                       all.files=TRUE, full.names=TRUE)

#rasterlist[8] <-NA
#rasterlist=na.omit(rasterlist)
#rasterlist[8]=NULL


rasterNames=sub('.*/', '', rasterlist)#remove file pathways from list
rasterNames=sub('*.tiff', '', rasterNames)#remove extension from list
rasterNames = substr(rasterNames,1,nchar(rasterNames)-11) # remove "mosaic"


all_rasters <- terra::rast(rasterlist)#Create SpatRaster from list


names(all_rasters) = rasterNames#rename the rasters in the raster stack by the raster file names


rd_extract <- terra::extract(all_rasters, vect(Sections_LocalData), fun = mean, na.rm=TRUE)  # returns a DF
rd_extract<- cbind(Sections_LocalData, rd_extract)
write.csv(st_drop_geometry(rd_extract), "C:/LocalData/NFI/iDeer/Data/Outputs/rd_NFI_extract.csv") # save (maybe change the filepath?)

hist(rd_extract$Brd_250_mosaic_all_tiles_co)
hist(rd_extract$Brd_1000_mosaic_all_tiles_co)
hist(rd_extract$motorway_1000_mosaic_all_tiles_co)


plot(all_rasters[1])

```









```{r}
#These woody linear features

wood_file_path <- "C:/LocalData/NFI/iDeer/Data/Inputs/LinearFeatures"

rasterlist <- list.files(path = wood_file_path, pattern='.tif',
                       all.files=TRUE, full.names=TRUE)

#rasterlist[8] <-NA
#rasterlist=na.omit(rasterlist)
#rasterlist[8]=NULL


rasterNames=sub('.*/', '', rasterlist)#remove file pathways from list
rasterNames=sub('*.tiff', '', rasterNames)#remove extension from list
rasterNames = substr(rasterNames,1,nchar(rasterNames)-11) # remove "mosaic"


all_rasters <- terra::rast(rasterlist)#Create SpatRaster from list


names(all_rasters) = rasterNames#rename the rasters in the raster stack by the raster file names


wood_extract <- terra::extract(all_rasters, vect(Sections_LocalData), fun = mean,  na.rm=TRUE)  # returns a DF
wood_extract<- cbind(Sections_LocalData, wood_extract)
write.csv(st_drop_geometry(wood_extract), "C:/LocalData/NFI/iDeer/Data/Outputs/woody_NFI_extract.csv") # save (maybe change the filepath?)
```


