---
title: "iDeer Tool Function"
format: html
editor: visual

#User selects woodland
#Metrics are calculated for the focal woodland and all woodlands in surrounding 10km landscape
#Cut all rasters to 10km, buffer by 2km for spatial processes (12km total)
#Extract the metrics (roads, urban, connectivity etc)
#Run the BBN
#Loop this process for the 5 different landscapes
#Do this for the 5 DIA regions
#Write as a function
---

```{r}
library(plyr)
library(ggplot2)
library(dplyr)
library(sf)
library(raster)
library(terra)
library(here)
library(progress)
library(bnlearn)
library(RColorBrewer)
library(rasterVis)
library(ggmap)

here()

bng <- 27700

```

### Load spatial data derived for England and Wales

```{r}

#Load NFI/LCM2015 master map
map <- raster(here("data/derived-data/2015-spatial-layers/edge_core_raster_all_tiles_EW.tif"))
crs(map) <- bng

#Load main roads at 250m scale
#Summed density of A roads, B roads and motorways
#all_roads <- raster(here("data/derived-data/2015-spatial-layers/all_roads_250_mosaic_all_tiles_corrected.tif"))
#crs(all_roads) <- bng

#Load nearest road
#Nearest A road, B road or motorway
nearest_road <- raster(here("data/derived-data/2015-spatial-layers/nearest_road_raster_2015_all_tiles_corrected.tif"))
crs(nearest_road) <- bng

#Load nearest urban
nearest_urban <- raster(here("data/derived-data/2015-spatial-layers/nearest_urban_raster_2015_all_tiles_corrected.tif"))
crs(nearest_urban) <- bng
nearest_suburb <- raster(here("data/derived-data/2015-spatial-layers/nearest_suburban_raster_2015_all_tiles_corrected.tif"))
crs(nearest_suburb) <- bng

#Load linear features at 1km scale
lf <- raster(here("data/derived-data/2015-spatial-layers/linear_features/lf_1000_mosaic_all_tiles_corrected.tif"))
crs(lf) <- bng

#Load MAX DAMS calculated at 1km scale
dams <- raster(here("data/derived-data/2015-spatial-layers/Max_DAMS/focal.dams.1000.max.non_wood_EW.tif"))

#Read in woodland polygon dataset
#Combined NFI 2015 and LCM 2015 woodlands
#These polygons will be used to calculate the connectivity, they are unified
#All touching polygons have become unified
hab_patches_all <- st_read(here("data/derived-data/nfi_lcm_overlaid_shapefile_woods_export.shp"))
st_crs(hab_patches_all) <- bng

head(hab_patches_all)

hab_patches_all <- hab_patches_all %>%
  rename(patch_ID = Id) %>%
  rename(patch_area = Shape_Area)

#Non-unified lcm polygons
#This layer is made from nfi_lcm_2015_overlaid.tif
#nfi_lcm_mosaic <- raster(here("data/derived-data/nfi_lcm_2015_overlaid.tif"))
#Made in ArcGIS
#Raster to polygon of all land cover classes

lcm_polys <- st_read(here("data/derived-data/nfi_lcm_2015_overlaid.shp"))
#Select for woodland polys only
wood_polys <- lcm_polys %>% filter(gridcode %in% c("1","2","22","23")) %>%
  rename(patch_ID = Id,
         class = gridcode)


```

### Choose woodlands

NEED TO CHANGE THIS TO BE ABLE TO DRAW A POLYGON INSIDE WOODLANDS SO THAT USER CAN SELECT THEIR LAND PARCEL

```{r}

#Choose five woodlands within the DIA priority areas to extract data for
#This is simulating the user selecting a woodland of interest

#Suffolk
#suffolk_wood <- hab_patches_all %>% subset(patch_ID == 260977)
#Cumbria
#cumbria_wood <- hab_patches_all %>% subset(patch_ID == 142765)
#South Wales 
#sw_wood <- hab_patches_all %>% subset(patch_ID == 340972)
#North Wales
#nw_wood <- hab_patches_all %>% subset(patch_ID == 188539)
#Oxford 
#ox_wood <- hab_patches_all %>% subset(patch_ID == 370647)

#select all woodlands together
#woods <- wood_polys %>%
#  filter(patch_ID %in% c("213709", "377707", "455598", "188539", "858585"))
#woods$Region <- c("North West","Cumbria","North Wales","Suffolk","Oxford")
#woods <- st_as_sf(woods)

#Just do one wood for now

woods <- wood_polys %>% filter(patch_ID %in% c("455598"))
woods$Region <- c("Suffolk")
woods <- st_as_sf(woods)



```

### Buffer woodlands for spatial data extraction

```{r}
#Buffer all woods by 5km
buffered_woods_5km <- st_as_sf(st_buffer(woods, dist = 5000))  # dist = 5km in meters
#Buffer all woods by 7km (5km landscape + 2km edge buffer)
buffered_woods_7km <- st_as_sf(st_buffer(woods, dist = 7000))  # dist = 7km in meters

#ensure crs of buffers is BNG
st_crs(buffered_woods_7km) <- bng
st_crs(buffered_woods_5km) <- bng

region_colors <- c("Cumbria" = "red","North Wales" = "blue",
                   "Suffolk"="magenta","North West"="yellow","Oxford"="orange")
buffered_woods_5km$color <- region_colors[buffered_woods_5km$Region]

# Plot the first geometry with no color
plot(st_geometry(buffered_woods_7km), col = NA, main = "Buffered Woods")

# Plot the second geometry with colors based on regions
plot(st_geometry(buffered_woods_5km), col = buffered_woods_5km$color, add = TRUE)

#Check points, woodlands buffers line up
plot(st_geometry(buffered_woods_7km), col = NA)
plot(st_geometry(buffered_woods_5km), col = buffered_woods_5km$color, add = TRUE)
plot(st_geometry(woods), col = "black", add = TRUE)
#plot(st_geometry(wood_points), add = TRUE, col = "red")

#Filter for woodland patches that intersect buffers
woods_buffers <- st_filter(wood_polys, buffered_woods_5km, .predicate = st_intersects)

#Inspect land cover layer

landscape_5km <- crop(map, buffered_woods_5km)
plot(landscape_5km)

```

### Extract raster variables for all woodlands in 5km buffer

Produce dataframes with pixel ID's from land cover map. This will allow calculation of pixel-level deer damage risk. Need to ensure there is a "patch_ID" column so that the connectivity values can be assigned to pixels within each patch (calculated at patch-level).

```{r}
#NEAREST URBAN EXTRACTION ####

source(here("code/functions/extract_urban_raster_function.R"))

nearest_urban_extraction <- extract_urban_rasters(fivekm_buffer=buffered_woods_5km, 
raster_layer=nearest_urban,
land_cover_raster=map)

write.csv(nearest_urban_extraction, here("data/derived-data/Tool-extract-function-test/nearest_urban_extraction_pixelIDs_5km.csv"))

#Plot
ggplot(nearest_urban_extraction) +
  geom_tile(aes(x = x, y = y, fill = nearest_urban)) +
  scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Distance to Nearest Urban area",
       fill = "Distance (m)")  # Adjust the legend title to be more descriptive
```

```{r}

source(here("code/functions/extract_suburban_raster_function.R"))

nearest_suburban_extraction <- extract_suburban_rasters(fivekm_buffer=buffered_woods_5km, 
raster_layer=nearest_suburb,
land_cover_raster=map)

write.csv(nearest_suburban_extraction, here("data/derived-data/Tool-extract-function-test/nearest_suburban_extraction_pixelIDs_5km.csv"))

#Plot
ggplot(nearest_suburban_extraction) +
  geom_tile(aes(x = x, y = y, fill = nearest_suburban)) +
  scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Distance to Nearest Suburban area",
       fill = "Distance (m)")  # Adjust the legend title to be more descriptive
```

```{r}
#MAX DAMS EXTRACTION AT PIXEL LEVEL ####

#Call function to extract values from pixels in dams raster
#The extracted output contains the following columns:
#max_dams = maximum dams value within 1km of that pixel
#pixel_ID = obtained from full land cover map
#x = x coordinate of pixel
#y = y coordinate of pixel
source(here("code/functions/extract_dams_raster_function.R"))

dams_extraction <- extract_dams_raster(fivekm_buffer=buffered_woods_5km, 
raster_layer=dams,
land_cover_raster=map)

write.csv(dams_extraction, here("data/derived-data/Tool-extract-function-test/dams_extraction_pixelIDs_1000_5km.csv"))

#Plot
ggplot(dams_extraction) +
  geom_tile(aes(x = x, y = y, fill = max_dams)) +
  scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Max Dams within 1km",
       fill = "Max DAMS")  
```

```{r}

#LINEAR FEATURE DENSITY WITHIN 1KM EXTRACTION AT PIXEL LEVEL ####

#Call function to extract values from pixels in linear feature raster
#The extracted output contains the following columns:
#lf = total length of linear features within 1km of that pixel
#pixel_ID = obtained from full land cover map
#x = x coordinate of pixel
#y = y coordinate of pixel
source(here("code/functions/extract_linear_feature_density_raster_function.R"))

lf_extraction <- extract_lf_raster(buffered_woods_5km=buffered_woods_5km, 
raster_layer=lf,
land_cover_raster=map)

write.csv(lf_extraction, here("data/derived-data/Tool-extract-function-test/linear_feature_extraction_pixelIDs_1000_5km.csv"))

#Plot
ggplot(lf_extraction) +
  geom_tile(aes(x = x, y = y, fill = lf)) +
  scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Linear feature density",
       fill = "Linear feature length within 1km")  # Adjust the legend title to be more descriptive
```

```{r}
#NEAREST ROAD EXTRACTION AT PIXEL LEVEL ####

#Call function to extract values from pixels in nearest road raster
#This raster layer includes A roads, B roads and motorways only (not minor roads)
#The extracted output contains the following columns:
#nearest_road = distance to the main road nearest to that pixel.
#pixel_ID = obtained from full land cover map
#x = x coordinate of pixel
#y = y coordinate of pixel
source(here("code/functions/extract_nearest_road_raster_function.R"))

nearest_road_extraction <- nearest_road_rasters(buffered_woods_5km=buffered_woods_5km, 
raster_layer=nearest_road,
land_cover_raster=map)

write.csv(nearest_road_extraction, here("data/derived-data/Tool-extract-function-test/nearest_road_extraction_pixelIDs_5km.csv"))

#Plot
ggplot(nearest_road_extraction) +
  geom_tile(aes(x = x, y = y, fill = nearest_road)) +
  scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Nearest road",
       fill = "Distance (m)")  
```

#### Land cover extraction at 1km around woodland pixels

```{r}
#LAND COVER EXTRACTION AT 1KM AROUND WOODLAND PIXELS ####

#Reclassify land cover map using forage quality scores

edge.val.types <- read.csv(file.path(here("data/raw-data/Expert_Qnaire/Expert_derived_quality_ranks_edge_types_with_water.csv")))

#make is-becomes matrix
reclass_vals <- edge.val.types%>%dplyr::select(Forage_Q,class) %>%
  rename(becomes = Forage_Q,
         is = class) %>%
  relocate(is) %>%
  filter(!is.na(is))

#Crop map to buffer_7km
map_7km <- crop(map, buffered_woods_7km)

#Reclassify
map_reclass <- raster::reclassify(map_7km, reclass_vals)

circle.buff = raster::focalWeight (map_reclass, d=1000, type="circle",fillNA=T )#Create buffer
circle.buff[circle.buff > 0] <- 1   # replacing weights by 1
Focal1000= raster::focal(map_reclass, w=circle.buff, fun=mean, na.rm=T, here("output/intermediate-outputs/focal.forage.quality.mean.tif"), overwrite=T)#do focal

plot(Focal1000)


```

```{r}

#Extract the mean forage quality values ####

# Use woodland patches that intersect 5km buffer
plot(woods_buffers$geometry)

# Crop land cover map by the 5km buffer
fcrop <- crop(Focal1000, buffered_woods_5km)
plot(fcrop)

#Mask the land cover map with the woodland polygons
fmask <- mask(fcrop, woods_buffers)
plot(fmask)

# Get extent of 5km buffer
km5_extent <- as(extent(st_bbox(buffered_woods_5km)), "Extent")
  
  # Get the cell numbers within the extent
  pixel_ID <- cellsFromExtent(map, km5_extent)
  
  # Get the coordinates for these cells
  cell_coords <- as.data.frame(xyFromCell(map, pixel_ID))
  
  # Extract pixel values
  pixel_values <- extract(fmask, cell_coords)
  
    # Create a data frame with the extracted values and cell indices
    lc_quality_df <- data.frame(pixel_ID = pixel_ID,
                       lc_forage_q = pixel_values,
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    lc_quality_df <- lc_quality_df[!is.na(lc_quality_df$lc_forage_q), ]

write.csv(lc_quality_df, here("data/derived-data/Tool-extract-function-test/landscapemetrics_1000_pixelIDs.csv"))

#Plot
ggplot(lc_quality_df) +
  geom_tile(aes(x = x, y = y, fill = lc_forage_q)) +
  scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Mean forage quality within 1km",
       fill = "Mean forage quality")
```

#### Woodland pixel extraction

```{r}
#WOODLAND PIXEL EXTRACTION ####
#Do this at the pixel level

# Use woodland patches that intersect 5km buffer
plot(woods_buffers$geometry)

# Crop land cover map by the 5km buffer
fcrop <- crop(map, buffered_woods_5km)
plot(fcrop)

#Mask the land cover map with the woodland polygons
fmask <- mask(fcrop, woods_buffers)
plot(fmask)

# Get extent of 5km buffer
km5_extent <- as(extent(st_bbox(buffered_woods_5km)), "Extent")
  
  # Get the cell numbers within the extent
  pixel_ID <- cellsFromExtent(map, km5_extent)
  
  # Get the coordinates for these cells
  cell_coords <- as.data.frame(xyFromCell(map, pixel_ID))
  
  # Extract pixel values
  pixel_values <- extract(fmask, cell_coords)
  
    # Create a data frame with the extracted values and cell indices
    woodpix <- data.frame(pixel_ID = pixel_ID,
                       class = pixel_values,
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    woodpix <- woodpix[!is.na(woodpix$class), ]
 
write.csv(woodpix, here("data/derived-data/Tool-extract-function-test/woodland_type_pixelIDs_5km.csv"))

#Plot
ggplot(woodpix) +
  geom_tile(aes(x = x, y = y, fill = class)) +
  scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Land cover",
       fill = "Pixel class")  
```

### Assign woodland pixel quality values

```{r}
woodpix <- read.csv(here("data/derived-data/Tool-extract-function-test/woodland_type_pixelIDs_5km.csv"))
woodpix <- woodpix%>%dplyr::select(-c("X"))

unique(woodpix$class)

#Load landscape nutritional scores for raster categories
#Edge forage quality scores sum of woodland score + adjacent land cover score
#Water scores added in to prevent loss of woodland edge pixels
#Water edge pixels receive one point higher than regular woodland
#Still an edge so will be highly nutritious
edge.val.types <- read.csv(file.path(here("data/raw-data/Expert_Qnaire/Expert_derived_quality_ranks_edge_types_with_water.csv")))
edge.val.types <- edge.val.types%>%dplyr::select(-"X")

#Check edge.val.types$class and woods$class match in format
unique(edge.val.types$class)
class(edge.val.types$class)
edge.val.types <- edge.val.types%>%arrange(class)
unique(edge.val.types$class)

class(woodpix$class)
woods <- woodpix%>%arrange(class)
unique(woodpix$class)
pcm <-woodpix%>%arrange(class)
unique(pcm$class)

# Round to 3 decimal places
woodpix$class <- round(woodpix$class, 3)
edge.val.types$class <- round(edge.val.types$class, 3)

#Stitch the quality scores into the lcm dataset

wood.edge.types <- woodpix %>% left_join(edge.val.types, by = c("class")) #%>%
                                          #filter(!is.na(Land.cover))

#Replace "NA" in Forage_Q column with 0
wood.edge.types <- wood.edge.types%>% 
 mutate(Forage_Q = ifelse(is.na(Forage_Q), 0, Forage_Q))%>%
   dplyr::select(-c("class","Land.cover"))%>%
  rename(woodland_q = Forage_Q)

#Plot
ggplot(wood.edge.types) +
  geom_tile(aes(x = x, y = y, fill = woodland_q)) +
  scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Woodland forage quality",
       fill = "Pixel quality")  
```

### Calculate connectivity for every woodland inside the 5km buffers

```{r}
#### SET CONNECTIVITY PARAMETERS ####
# connectivity parametetrs - % of dispersers reaching a set distance #
percentage_dispersers <- 0.05 ### 5% - 95% of individuals will go to woodlands
dispersal_distance <- 400 #Gets curve up to 1000m on x axis
dispersal_contribution <-
  -((log(1 / percentage_dispersers)) / dispersal_distance)

# Buffer distance represents the cut off - this stops the script measuring every pairwise combination
dispersal_cutoff <- 0.999 ### 99.9% cut off
#buffer_cutoff <-
  #round(log(1 / (1 - dispersal_cutoff)) / (log(1 / percentage_dispersers) /
                                             #dispersal_distance), digits = 2)
buffer_cutoff = 1000 #1km

# test & plot of dispersal contribution for sequence of distance up to buffer_cutoff
test <-
  data.frame(distance = (seq(0, buffer_cutoff, by = 0.5))) %>%
  mutate(contribution = (exp(dispersal_contribution * distance) * 100))

ggplot(data = test, aes(x = distance, y = contribution, group = 1)) +
  geom_line() +
  geom_point() +
  geom_vline(xintercept = buffer_cutoff, color = "red")

#See how buffer size looks
#focal_patch_buffer <- st_buffer(woods[1,], buffer_cutoff)
#plot(st_geometry(focal_patch_buffer))
#plot(st_geometry(woods[1,]),add=TRUE)

```

```{r}

#Connectivity for loop ####

connectivity_results <- list()

  # Use the 7km buffer as the source patches area
  source_woods <- st_filter(hab_patches_all, buffered_woods_7km, .predicate = st_intersects)

  # Filter the woodlands in the original 5km square to get the focal patches
  focal_woods <- st_filter(hab_patches_all, buffered_woods_5km, .predicate = st_intersects)
  n_distinct(focal_woods$patch_ID)

  # Calculate connectivity for each focal woodland
  Connectivity_table <- NULL
  
  # Loop through each focal woodland to calculate connectivity
  for (j in 1:nrow(focal_woods)) {
    
    # Focal patch in the 10km square
    focal_patch <- focal_woods[j, ]
  
    # Buffer 1000m around the focal patch
    focal_patch_buffer <- st_buffer(focal_patch, buffer_cutoff)
    
    # Find source patches within this buffer
    source_patches <- st_filter(source_woods, focal_patch_buffer, .predicate = st_intersects)
    
    # Calculate distances between the focal and source patches
    patch_dist <- as.vector(st_distance(focal_patch, source_patches))
    
    # Create a data frame for connectivity information
    Conn_table_site <- data.frame(
      focal_patch = focal_patch$patch_ID,
      focal_patch_area = as.numeric(focal_patch$patch_area),
      source_patch = source_patches$patch_ID,
      source_patch_area = as.numeric(source_patches$patch_area),
      distance = patch_dist,
      incoming_connect = ifelse(
        focal_patch$patch_ID == source_patches$patch_ID,
        NA,
        source_patches$patch_area * exp(dispersal_contribution * patch_dist)
      )
    )
    
    # Append to 'Connectivity_table'
    if (is.null(Connectivity_table)) {
      Connectivity_table <- Conn_table_site
    } else {
      Connectivity_table <- rbind(Connectivity_table, Conn_table_site)
    }
  
  n_distinct(Connectivity_table$focal_patch)

  # Store results for this woodland
  connectivity_results[[paste("focal_wood", focal_woods$patch_ID[j])]] <- Connectivity_table
}
```

```{r}

#Remove rows where focal_patch = source_patch
Connectivity_table_filt <- Connectivity_table %>%
  filter(focal_patch != source_patch)

# Identify missing unique patch IDs
# These patches have no source patches so will have been filtered out
# We need to put them back in so we can give them an incoming_connect value of 0
missing_patch_ids <- setdiff(
     unique(Connectivity_table$focal_patch),
     unique(Connectivity_table_filt$focal_patch)
 )
 
 print(missing_patch_ids)  # This will show which patch IDs are missing

missing_patches <- Connectivity_table %>% filter(focal_patch %in% missing_patch_ids) %>%
  # Replace NAs in 'incoming_connect' with 0
  mutate(incoming_connect = coalesce(incoming_connect, 0))  # Replace NA with 0

#Make rows for patches with no connectivity so we don't lose any patches
Connectivity_table_filt <- rbind(Connectivity_table_filt, missing_patches)

#Ensure no duplicate rows
Connectivity_table_filt <- Connectivity_table_filt %>% distinct()

#Now no patches should be missing
n_distinct(Connectivity_table_filt$focal_patch)

#sum incoming connectivity by focal patch
incoming_connectivity_sum <- Connectivity_table_filt %>%
  dplyr::select(focal_patch, incoming_connect) %>%
  group_by(focal_patch) %>%
  dplyr::summarise(total_connect = sum(incoming_connect),
            n = n()) %>%
  mutate(n = if_else(total_connect == 0, 0, n)) %>%  # Set 'n' to 0 when 'incoming_connect' is 0
  rename(patch_ID = focal_patch)


```

### Get pixel-level connectivity from patch-scale data

```{r}

#subset hab_patches_all that have a connectivity value in incoming_connectivity_sum
hab_patches_connect <- hab_patches_all %>%
  filter(patch_ID %in% incoming_connectivity_sum$patch_ID)
#left_join the connectivity dataset to the geometry
incoming_connect_vals <- incoming_connectivity_sum %>%
  dplyr::select(-c("n"))

hab_patches_connect <- left_join(hab_patches_connect, incoming_connect_vals, by="patch_ID")
hab_patches_connect<-st_as_sf(hab_patches_connect)

#plot
ggplot(data = hab_patches_connect) +
  geom_sf(aes(fill = total_connect)) +
  scale_fill_viridis_c() +  # Optional: for a nice color scale
  theme_minimal() 

#fasterize, use land cover map as template
connect_raster <- fasterize::fasterize(hab_patches_connect, raster=Focal1000,field="total_connect")
plot(connect_raster)
crs(connect_raster) <- bng

#crop raster to buffer_5km

connect_raster_5km <- crop(connect_raster, buffered_woods_5km)
plot(connect_raster_5km)

# List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
  buffer_geometry <- buffered_woods_5km$geometry

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
 # Get the cell numbers within the extent
pixel_ID <- cellsFromExtent(map, buffer_extent)

# Get the coordinates for these cells
cell_coords <- as.data.frame(xyFromCell(map, pixel_ID))

#Extract pixel values
pixel_values <- extract(connect_raster_5km, cell_coords)

    # Create a data frame with the extracted values and cell indices
    connect.vals <- data.frame(total_connect = pixel_values, 
                       pixel_ID = pixel_ID,
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    connect.vals <- connect.vals[!is.na(connect.vals$total_connect), ]
    
    # Append to the results list
    extraction_results <- connect.vals
  
  # Combine all results into a single dataframe
  connect_df <- bind_rows(extraction_results)
  
  #Plot
ggplot(connect_df) +
  geom_tile(aes(x = x, y = y, fill = total_connect)) +
  scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Incoming connectivity",
       fill = "Total connectivity")  # Adjust the legend title to be more descriptive

```

#### left_join the data sets together

```{r}

#Check pixel_ID cols match

# Extract pixel_ID columns from both dataframes
#pixel_IDs_lc_df <- lc_quality_df$pixel_ID
#pixel_IDs_woodpix <- wood.edge.types$pixel_ID
# To find specific differences, you can use set operations
#not_shared <- setdiff(pixel_IDs_woodpix, pixel_IDs_lc_df)

#lc_quality_df <-lc_quality_df%>%dplyr::select(-c("patch_ID"))
#wood.edge.types<-wood.edge.types%>%dplyr::select(-c("patch_ID"))

#Forage quality of pixels and Nutritional Landscape Index for patch
df <- left_join(lc_quality_df, wood.edge.types, by = c("pixel_ID","x","y"))

#Connectivity ####
#incoming_connectivity_sum<-read.csv(here("data/derived-data/Tool-extract-function-test/incoming_connectivity_sum_5km.csv"))
#incoming_connectivity_sum<-incoming_connectivity_sum %>%
     #dplyr::select(-c("X","n"))
#connect_df <- connect_df %>% dplyr::select(-c("patch_ID"))
df <- left_join(df, connect_df, by=c("pixel_ID","x","y"))

#MAX DAMS ####
dams_scores <- dams_extraction%>%rename(max.dams.1000=max_dams)%>%
  st_drop_geometry()

df <- left_join(df, dams_scores, by=c("pixel_ID","x","y"))

#Nearest road
df <- left_join(df, nearest_road_extraction, by=c("pixel_ID","x","y"))

#Nearest urban and suburban ####

df <- left_join(df, nearest_suburban_extraction, by=c("pixel_ID","x","y"))

df <- left_join(df, nearest_urban_extraction, by=c("pixel_ID","x","y"))

#Make column for nearest urban and suburban using minimum value
df$urban_proximity <- pmin(df$nearest_suburban, df$nearest_urban)
#Drop individual columns
df <- df%>%dplyr::select(-c("nearest_suburban","nearest_urban"))

#Linear features####
df <- left_join(df, lf_extraction,  by=c("pixel_ID","x","y"))

#Remove the NAs

df <- na.omit(df)


```

```{r}

#Plot the data ####

#Nearest_road
ggplot(df) +
  geom_tile(aes(x = x, y = y, fill = nearest_road)) +
  scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Distance to Nearest Road",
       fill = "Distance (m)")  # Adjust the legend title to be more descriptive

#Urban proximity
ggplot(df) +
  geom_tile(aes(x = x, y = y, fill = urban_proximity)) +
   scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Distance to Nearest Urban Area",
       fill = "Distance (m)")  # Adjust the legend title to be more descriptive

#Linear features

ggplot(df) +
  geom_tile(aes(x = x, y = y, fill = lf)) +
  scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Linear feature density",
       fill = "LF length within 1km")  # Adjust the legend title to be more descriptive

#Nutritional landscape composition
ggplot(df) +
  geom_tile(aes(x = x, y = y, fill = lc_forage_q)) +
  scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Nutritional landscape composition",
       fill = "Weighted mean lc within 1km")  # Adjust the legend title to be more descriptive

#Connectivity
ggplot(df) +
  geom_tile(aes(x = x, y = y, fill = total_connect)) +
   scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Connectivity",
       fill = "Total incoming connectivity")  # Adjust the legend title to be more descriptive

#Woodland Forage quality
ggplot(df) +
  geom_tile(aes(x = x, y = y, fill = woodland_q)) +
  scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Pixel Forage quality",
       fill = "Forage Quality score")  # Adjust the legend title to be more descriptive

#Max DAMS within 1km
ggplot(df) +
  geom_tile(aes(x = x, y = y, fill = max.dams.1000)) +
   scale_fill_viridis_c(option = "plasma") +  # Apply a continuous viridis color scale
  theme_minimal() +
  guides(fill = guide_colorbar(barwidth = 1, barheight = 10)) +  # Customize the colorbar appearance
  labs(title = "Max DAMS within 1km",
       fill = "Max DAMS")  # Adjust the legend title to be more descriptive

saveRDS(df, here("data/derived-data/Tool-extract-function-test/bbn_test_df_5_patches_pixel_level_5km.rds"))
```

#### Convert numerical values to categorical (LOW, MED, HIGH)

```{r}

#Replace values in numerical columns with LOW, MED, HIGH
#This is based on logic from google spreadsheet "BN thresholds" 

df <- readRDS(here("data/derived-data/Tool-extract-function-test/bbn_test_df_5_patches_pixel_level_5km.rds"))

#Look at distribution of values across national datasets ####
#We need to decide at what point values change between categories- need to do this using whole datasets rather than localised cropped datasets.

#Roads
#hist(all_roads$all_roads_250_mosaic_all_tiles_corrected)
#df$all_roads_250 <- ifelse(is.na(df$all_roads_250), NA,
#                           ifelse(df$all_roads_250 <= 0, "LOW",
#                           ifelse(df$all_roads_250 > 0 & df$all_roads_250 <= 100, "MED",
#                           ifelse(df$all_roads_250 > 100, "HIGH", NA))))
#unique(df$all_roads_250)

#Nearest roads
#Proximity, therefore lower distance = higher category
df$nearest_road <- ifelse(is.na(df$nearest_road), NA,
                           ifelse(df$nearest_road <= 100, "HIGH",
                           ifelse(df$nearest_road > 100 & df$nearest_road <= 500, "MED",
                           ifelse(df$nearest_road > 500, "LOW", NA))))

#Linear features
#hist(lf$lf_1000_mosaic_all_tiles_corrected)
# Replace numeric values with categorical labels
df$lf <- ifelse(df$lf<= 2500, "LOW",
            ifelse(df$lf> 2500 & df$lf <= 5000, "MED",
            ifelse(df$lf> 5000, "HIGH", NA)))
unique(df$lf)


#nearest urban/suburban
#Proximity, therefore "HIGH" means lower value (lower distance)
#These thresholds are based on deer biology rather than the spead of data
#hist(nearest_urban$nearest_urban_raster_2015_all_tiles_corrected)
#hist(nearest_suburb$nearest_suburban_raster_2015_all_tiles_corrected)
df$urban_proximity <- ifelse(df$urban_proximity <= 500, "HIGH",
                          ifelse(df$urban_proximity >500 & df$urban_proximity <= 1000, "MED",
                           ifelse(df$urban_proximity >1000,"LOW",NA)))  
unique(df$urban_proximity)

#MAX DAMS

df$max.dams.1000 <- ifelse(df$max.dams.1000 <=15,"LOW",
                      ifelse(df$max.dams.1000>15 & df$max.dams.1000 <= 20, "MED",
                        ifelse(df$max.dams.1000 >20, "HIGH",NA)))

#nutritional landscape composition score - do we need to change these thresholds????
hist(edge.val.types$Forage_Q)
hist(df$lc_forage_q)
df$lc_forage_q <- ifelse(df$lc_forage_q <= 4, "LOW",
                          ifelse(df$lc_forage_q >4 & df$lc_forage_q <=8, "MED",
                           ifelse(df$lc_forage_q > 8, "HIGH",NA)))

#nutritional patch composition score

hist(edge.val.types$Forage_Q)
hist(df$woodland_q)
df$woodland_q <- ifelse(df$woodland_q<= 4, "LOW",
                          ifelse(df$woodland_q >4 & df$woodland_q <=8, "MED",
                           ifelse(df$woodland_q>8, "HIGH",NA)))

#connectivity
#hist(df$total_connect)
#quantile(df$total_connect)

#log transform connectivity values, they are too big for easy interpretation

df$total_connect_log <- log(df$total_connect)

#hist(df$total_connect_log)
quantile(df$total_connect_log)

df$total_connect_log <- ifelse(df$total_connect_log <=10, "LOW",
                          ifelse(df$total_connect_log >10 & df$total_connect_log  <=15, "MED",
                           ifelse(df$total_connect_log  > 15, "HIGH",NA)))

#remove original connectivity column

df <- df%>%dplyr::select(-c("total_connect"))

#Export

df_categorical <- df

df_categorical <- na.omit(df_categorical)

saveRDS(df_categorical, file = file.path(here("data/derived-data/Tool-extract-function-test/tool_test_extracted_data_categorical_pixel_level_5km.rds")))

```

### Create the Bayesian Belief Network

#### Initialize Conditional Probability Tables

```{r}
#CPTS for measured nodes ####
urban_prox_cpt<-matrix(c(0.333333,0.333333,0.333333), ncol=3, dimnames = list(NULL, c("LOW", "MED","HIGH")))

nearest_road_cpt<-matrix(c(0.333333,0.333333,0.333333), ncol=3, dimnames = list(NULL, c("LOW", "MED","HIGH")))

connectivity_cpt<-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

landscape_nutritional_score_cpt<-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

pixel_nutritional_score_cpt <- matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

dams_cpt <- landscape_nutritional_score_cpt<-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

lf_density_cpt <-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))
```

```{r}
#CPT for Disturbance Index ####

disturb_index_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3, 3, 3),  # Shape of the array (3x3x3)
  dimnames = list(
    disturb_index = c("LOW", "MED", "HIGH"),
    nearest_road = c("LOW", "MED", "HIGH"),
    urban_prox = c("LOW","MED","HIGH"),
    connectivity = c("LOW", "MED", "HIGH")
  )
)
#order = disturb_index, nearest_road, urban_prox,connectivity

#table1: urban_prox = LOW, connectivity = LOW
disturb_index_cpt["LOW","LOW","LOW","LOW"] <- 1
disturb_index_cpt["MED","MED","LOW","LOW"] <- 1
disturb_index_cpt["HIGH","HIGH","LOW","LOW"] <- 1
#table2:  urban_prox = MED, connectivity = LOW
disturb_index_cpt["MED","LOW","MED","LOW"] <- 1
disturb_index_cpt["MED","MED","MED","LOW"] <- 1
disturb_index_cpt["HIGH","HIGH","MED","LOW"] <- 1
#table3: urban_prox = HIGH, connectivity = LOW
disturb_index_cpt["HIGH","LOW","HIGH","LOW"] <- 1
disturb_index_cpt["HIGH","MED","HIGH","LOW"] <- 1
disturb_index_cpt["HIGH","HIGH","HIGH","LOW"] <- 1
#table4: urban_prox = LOW, connectivity = MED
disturb_index_cpt["LOW","LOW","LOW","MED"] <- 1
disturb_index_cpt["MED","MED","LOW","MED"] <- 1
disturb_index_cpt["MED","HIGH","LOW","MED"] <- 1
#table5: urban_prox = MED, connectivity = MED
disturb_index_cpt["MED","LOW","MED","MED"] <- 1
disturb_index_cpt["MED","MED","MED","MED"] <- 1
disturb_index_cpt["HIGH","HIGH","MED","MED"] <- 1
#table6: urban_prox = HIGH, connectivity = MED
disturb_index_cpt["HIGH","LOW","HIGH","MED"] <- 1
disturb_index_cpt["HIGH","MED","HIGH","MED"] <- 1
disturb_index_cpt["HIGH","HIGH","HIGH","MED"] <- 1
#table7: urban_prox = LOW, connectivity = HIGH
disturb_index_cpt["LOW","LOW","LOW","HIGH"] <- 1
disturb_index_cpt["LOW","MED","LOW","HIGH"] <- 1
disturb_index_cpt["MED","HIGH","LOW","HIGH"] <- 1
#table8: urban_prox = MED, connectivity = HIGH
disturb_index_cpt["MED","LOW","MED","HIGH"] <- 1
disturb_index_cpt["MED","MED","MED","HIGH"] <- 1
disturb_index_cpt["HIGH","HIGH","MED","HIGH"] <- 1
#table9: urban_prox = HIGH, connectivity = HIGH
disturb_index_cpt["MED","LOW","HIGH","HIGH"] <- 1
disturb_index_cpt["MED","MED","HIGH","HIGH"] <- 1
disturb_index_cpt["HIGH","HIGH","HIGH","HIGH"] <- 1

disturb_index_cpt

```

```{r}
#CPT for Nutritional Landscape Index ####


NLI_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3, 3),  # Shape of the array (3x3)
  dimnames = list(
    NLI = c("LOW", "MED", "HIGH"),
    landscape_nutritional_score = c("LOW", "MED", "HIGH"),
    lf_density = c("LOW","MED","HIGH")
  )
)
#order = NLI, landscape_nutritional_score, lf_density
NLI_cpt["LOW","LOW","LOW"] <- 1
NLI_cpt["MED","MED","LOW"] <- 1
NLI_cpt["HIGH","HIGH","LOW"] <- 1
NLI_cpt["LOW","LOW","MED"] <- 1
NLI_cpt["HIGH","MED","MED"] <- 1
NLI_cpt["HIGH","HIGH","MED"] <- 1
NLI_cpt["MED","LOW","HIGH"] <- 1
NLI_cpt["HIGH","MED","HIGH"] <- 1
NLI_cpt["HIGH","HIGH","HIGH"] <- 1

NLI_cpt
```

```{r}
#CPT for Thermoregulation Index ####

thermoreg_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3),  # Shape of the array
  dimnames = list(
    thermoreg_index = c("LOW", "MED", "HIGH"),
    DAMS_MAX = c("LOW", "MED", "HIGH")
  )
)
#order = DAMS_MAX, thermoreg_index
thermoreg_cpt["LOW","LOW"] <- 1
thermoreg_cpt["MED","MED"] <- 1
thermoreg_cpt["HIGH","HIGH"] <- 1

thermoreg_cpt
```

```{r}
#Final CPT: deer damage risk

damage_risk_cpt <- array(
  0,  # Default probability for each cell
  dim = c(5, 3, 3, 3, 3),  # Shape of the array
  dimnames = list(
    damage_risk = c("LOW","LOW-MED","MED","MED-HIGH","HIGH"),
    woodland_q = c("LOW", "MED", "HIGH"),
    NLI = c("LOW", "MED", "HIGH"),
    thermoreg_index = c("LOW", "MED", "HIGH"),
    disturb_index = c("LOW", "MED", "HIGH")
    
  )
)
#order = damage_risk, patch_quality,landscape_nutrition,thermoreg_index,disturbance

#table1 landscape_nutrition = LOW, thermoreg_index = LOW, disturbance_index = LOW
damage_risk_cpt["LOW-MED","LOW","LOW","LOW","LOW"] <- 1 
damage_risk_cpt["LOW-MED","MED","LOW","LOW","LOW"] <- 1
damage_risk_cpt["MED","HIGH","LOW","LOW","LOW"] <- 1
#table2 landscape_nutrition = MED, thermoreg_index = LOW, disturbance_index = LOW
damage_risk_cpt["LOW-MED","LOW","MED","LOW","LOW"] <- 1
damage_risk_cpt["LOW-MED","MED","MED","LOW","LOW"] <- 1
damage_risk_cpt["MED","HIGH","MED","LOW","LOW"] <- 1
#table3 landscape_nutrition = HIGH, thermoreg_index = LOW, disturbance_index = LOW
damage_risk_cpt["LOW-MED","LOW","HIGH","LOW","LOW"] <- 1
damage_risk_cpt["MED","MED","HIGH","LOW","LOW"] <- 1
damage_risk_cpt["MED","HIGH","HIGH","LOW","LOW"] <- 1
#table4 landscape_nutrition = LOW, thermoreg_index = MED, disturbance_index = LOW
damage_risk_cpt["MED","LOW","LOW","MED","LOW"] <- 1
damage_risk_cpt["MED","MED","LOW","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","LOW","MED","LOW"] <- 1
#table5 landscape_nutrition = MED, thermoreg_index = MED, disturbance_index = LOW
damage_risk_cpt["MED","LOW","MED","MED","LOW"] <- 1
damage_risk_cpt["MED","MED","MED","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","MED","MED","LOW"] <- 1
#table6 landscape_nutrition = HIGH, thermoreg_index = MED, disturbance_index = LOW
damage_risk_cpt["MED","LOW","HIGH","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","MED","HIGH","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","HIGH","MED","LOW"] <- 1
#table7  landscape_nutrition = LOW, thermoreg_index = HIGH, disturbance_index = LOW
damage_risk_cpt["MED-HIGH","LOW","LOW","HIGH","LOW"] <- 1
damage_risk_cpt["MED-HIGH","MED","LOW","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","HIGH","LOW","HIGH","LOW"] <- 1
#table8 landscape_nutrition = MED, thermoreg_index = HIGH, disturbance_index = LOW
damage_risk_cpt["MED-HIGH","LOW","MED","HIGH","LOW"] <- 1
damage_risk_cpt["MED-HIGH","MED","MED","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","HIGH","MED","HIGH","LOW"] <- 1
#table9 landscape_nutrition = HIGH, thermoreg_index = HIGH, disturbance_index = LOW
damage_risk_cpt["MED-HIGH","LOW","HIGH","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","MED","HIGH","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","HIGH","HIGH","HIGH","LOW"] <- 1
#table10 landscape_nutrition = LOW, thermoreg_index = LOW, disturbance_index = MED
damage_risk_cpt["LOW","LOW","LOW","LOW","MED"] <- 1
damage_risk_cpt["LOW","MED","LOW","LOW","MED"] <- 1
damage_risk_cpt["LOW-MED","HIGH","LOW","LOW","MED"] <- 1
#table11 landscape_nutrition = MED, thermoreg_index = LOW, disturbance_index = MED
damage_risk_cpt["LOW-MED","LOW","MED","LOW","MED"] <- 1
damage_risk_cpt["LOW-MED","MED","MED","LOW","MED"] <- 1
damage_risk_cpt["MED","HIGH","MED","LOW","MED"] <- 1
#table12 landscape_nutrition = HIGH, thermoreg_index = LOW, disturbance_index = MED
damage_risk_cpt["LOW-MED","LOW","HIGH","LOW","MED"] <- 1
damage_risk_cpt["MED","MED","HIGH","LOW","MED"] <- 1
damage_risk_cpt["MED","HIGH","HIGH","LOW","MED"] <- 1
#table13 landscape_nutrition = LOW, thermoreg_index = MED, disturbance_index = MED
damage_risk_cpt["LOW-MED","LOW","LOW","MED","MED"] <- 1
damage_risk_cpt["LOW-MED","MED","LOW","MED","MED"] <- 1
damage_risk_cpt["MED","HIGH","LOW","MED","MED"] <- 1
#table14 landscape_nutrition = MED, thermoreg_index = MED, disturbance_index = MED
damage_risk_cpt["MED","LOW","MED","MED","MED"] <- 1
damage_risk_cpt["MED","MED","MED","MED","MED"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","MED","MED","MED"] <- 1
#table15  landscape_nutrition = HIGH, thermoreg_index = MED, disturbance_index = MED
damage_risk_cpt["MED","LOW","HIGH","MED","MED"] <- 1
damage_risk_cpt["MED-HIGH","MED","HIGH","MED","MED"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","HIGH","MED","MED"] <- 1
#table16 landscape_nutrition = LOW, thermoreg_index = HIGH, disturbance_index = MED
damage_risk_cpt["MED","LOW","LOW","HIGH","MED"] <- 1
damage_risk_cpt["MED","MED","LOW","HIGH","MED"] <- 1
damage_risk_cpt["MED","HIGH","LOW","HIGH","MED"] <- 1
#table17 landscape_nutrition = MED, thermoreg_index = HIGH, disturbance_index = MED
damage_risk_cpt["MED-HIGH","LOW","MED","HIGH","MED"] <- 1
damage_risk_cpt["MED-HIGH","MED","MED","HIGH","MED"] <- 1
damage_risk_cpt["HIGH","HIGH","MED","HIGH","MED"] <- 1
#table18 landscape_nutrition = HIGH, thermoreg_index = HIGH, disturbance_index = MED
damage_risk_cpt["MED-HIGH","LOW","HIGH","HIGH","MED"] <- 1
damage_risk_cpt["HIGH","MED","HIGH","HIGH","MED"] <- 1
damage_risk_cpt["HIGH","HIGH","HIGH","HIGH","MED"] <- 1
#table19 landscape_nutrition = LOW, thermoreg_index = LOW, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","LOW","LOW","HIGH"] <- 1
damage_risk_cpt["LOW","MED","LOW","LOW","HIGH"] <- 1
damage_risk_cpt["LOW","HIGH","LOW","LOW","HIGH"] <- 1
#table20 landscape_nutrition = MED, thermoreg_index = LOW, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","MED","LOW","HIGH"] <- 1
damage_risk_cpt["LOW","MED","MED","LOW","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","MED","LOW","HIGH"] <- 1
#table21 landscape_nutrition = HIGH, thermoreg_index = LOW, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","HIGH","LOW","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","HIGH","LOW","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","HIGH","LOW","HIGH"] <- 1
#table22 landscape_nutrition = LOW, thermoreg_index = MED, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","LOW","MED","HIGH"] <- 1
damage_risk_cpt["LOW","MED","LOW","MED","HIGH"] <- 1
damage_risk_cpt["LOW","HIGH","LOW","MED","HIGH"] <- 1
#table23 landscape_nutrition = MED, thermoreg_index = MED, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","MED","MED","HIGH"] <- 1
damage_risk_cpt["LOW","MED","MED","MED","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","MED","MED","HIGH"] <- 1
#table24 landscape_nutrition = HIGH, thermoreg_index = MED, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","HIGH","MED","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","HIGH","MED","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","HIGH","MED","HIGH"] <- 1
#table25  landscape_nutrition = LOW, thermoreg_index = HIGH, disturbance_index = HIGH
damage_risk_cpt["LOW-MED","LOW","LOW","HIGH","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","LOW","HIGH","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","LOW","HIGH","HIGH"] <- 1
#table26 landscape_nutrition = MED, thermoreg_index = HIGH, disturbance_index = HIGH
damage_risk_cpt["LOW-MED","LOW","MED","HIGH","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","MED","HIGH","HIGH"] <- 1
damage_risk_cpt["MED","HIGH","MED","HIGH","HIGH"] <- 1
#table27 landscape_nutrition = HIGH, thermoreg_index = HIGH, disturbance_index = HIGH
damage_risk_cpt["LOW-MED","LOW","HIGH","HIGH","HIGH"] <- 1
damage_risk_cpt["MED","MED","HIGH","HIGH","HIGH"] <- 1
damage_risk_cpt["MED","HIGH","HIGH","HIGH","HIGH"] <- 1

damage_risk_cpt
```

#### Create BBN structure

```{r}
# Step 1: Explicitly define the nodes in the network
nodes <- c(
  "disturb_index",
  "urban_prox",
  "nearest_road",
  "connectivity",
  "woodland_q",
  "landscape_nutritional_score",
  "NLI",
  "lf_density",
  "thermoreg_index",
  "DAMS_MAX",
  "damage_risk"
)

#empty graph
e = empty.graph(nodes)

arc.set = matrix(c("urban_prox", "disturb_index",
                     "nearest_road", "disturb_index", 
                     "connectivity", "disturb_index",
                     "lf_density","NLI",
                     "landscape_nutritional_score","NLI",
                     "woodland_q","damage_risk",
                     "DAMS_MAX","thermoreg_index",
                     "disturb_index","damage_risk",
                     "NLI","damage_risk",
                     "thermoreg_index","damage_risk"),
             ncol = 2, byrow = TRUE,
             dimnames = list(NULL, c("from", "to")))

arcs(e) <- arc.set

model_string <- modelstring(e) 

net<-model2network(model_string) 
# You dont need to repeat the names of the nodes with parents (child nodes) 

# Custom fitting network (matching up the nodes to their CPTs)
dfit = custom.fit(net, dist = list(nearest_road=nearest_road_cpt, 
                                   urban_prox=urban_prox_cpt,
                                   connectivity=connectivity_cpt,
                                   lf_density=lf_density_cpt,
                                   DAMS_MAX=dams_cpt,
                      woodland_q=pixel_nutritional_score_cpt,
                      landscape_nutritional_score=landscape_nutritional_score_cpt,
                                   disturb_index=disturb_index_cpt,
                                   NLI=NLI_cpt,
                                   thermoreg_index=thermoreg_cpt,
                                   damage_risk=damage_risk_cpt))

```

#### Plot BN structure

```{r plot the network}
#| echo: true
#| warning: false
#| message: false

graphviz.plot(net)

```

#### Check Parameters

```{r look at the dfit}
#| echo: true
#| warning: false
#| message: false
dfit
```

#### Set evidence

```{r set evidence}
#| echo: true
#| warning: false
#| message: false
cpquery(dfit, c(lf_density == "LOW"|DAMS_MAX == "HIGH"|landscape_nutritional_score == "HIGH"|connectivity=="LOW"|urban_prox=="LOW"|nearest_road == "HIGH"|DAMS_MAX=="HIGH"), (damage_risk=="HIGH"))

```

### Make predictions from BBN

```{r}
predict_dat <- df_categorical %>%st_drop_geometry() %>% #drop geometry
   mutate(across(where(is.character), toupper)) %>% #convert characters to upper case
   dplyr::select(-c("pixel_ID","x","y"))
#Ensure variable names match those in BBN
predict_dat<-predict_dat%>%rename(
                              nearest_road=nearest_road,
                              lf_density=lf,
                              urban_prox=urban_proximity,
                              landscape_nutritional_score=lc_forage_q,
                              DAMS_MAX=max.dams.1000,
                              connectivity=total_connect_log)

#Make empty columns for the latent (unobserved) variables

predict_dat$NLI <- NA
predict_dat$disturb_index <- NA
predict_dat$thermoreg_index <- NA
predict_dat$damage_risk <- NA

#Ensure all columns are factors, not characters
predict_dat <- predict_dat %>% mutate_all(as.factor)

#Ensure factor levels are in correct order
levels(predict_dat$nearest_road)<-c("LOW", "MED", "HIGH")
levels(predict_dat$landscape_nutritional_score)<-c("LOW", "MED", "HIGH")
levels(predict_dat$woodland_q)<-c("LOW", "MED", "HIGH")
levels(predict_dat$urban_prox)<-c("LOW", "MED", "HIGH")
levels(predict_dat$lf_density)<-c("LOW", "MED", "HIGH")
levels(predict_dat$DAMS_MAX)<-c("LOW", "MED", "HIGH")
levels(predict_dat$connectivity)<-c("LOW", "MED", "HIGH")
levels(predict_dat$NLI)<-c("LOW", "MED", "HIGH")
levels(predict_dat$disturb_index)<-c("LOW", "MED", "HIGH")
levels(predict_dat$thermoreg_index)<-c("LOW", "MED", "HIGH")
levels(predict_dat$damage_risk)<-c("LOW","LOW-MED","MED","MED-HIGH","HIGH")

#Ensure predict_data is a data.frame
predict_dat <- as.data.frame(predict_dat)

#Predict values for latent variables
pred_nli = predict(dfit,node="NLI",data=predict_dat, method = "bayes-lw",from=c("landscape_nutritional_score","lf_density"))
pred_thermoreg = predict(dfit,node="thermoreg_index",data=predict_dat, method = "bayes-lw", from="DAMS_MAX")
pred_disturb = predict(dfit,node="disturb_index",data=predict_dat, method = "bayes-lw",from=c("nearest_road","urban_prox","connectivity"))

#fill the columns
predict_dat$NLI <- pred_nli
predict_dat$thermoreg_index <- pred_thermoreg
predict_dat$disturb_index <- pred_disturb
  
#Predict damage
pred_damage = predict(dfit,node = "damage_risk",data = predict_dat,method = "parents")
#method = "bayes-lw"
predict_dat$damage_risk <- pred_damage

#Add damage risk back into spatial dataset

df_categorical$pred_damage <- pred_damage
df_categorical$pred_NLI <- pred_nli
df_categorical$pred_thermoreg <- pred_thermoreg
df_categorical$pred_disturb <- pred_disturb

#save
saveRDS(df_categorical, file = file.path(here("data/derived-data/Tool-extract-function-test/predicted_data_categorical_pixel_level_5km.rds")))
```

### Draw maps

```{r}
#| echo: true
#| warning: false
#| message: false


df_categorical <- readRDS(file = file.path(here("data/derived-data/Tool-extract-function-test/predicted_data_categorical_pixel_level_5km.rds")))

# Make sf object
df_sf <- st_as_sf(df_categorical, coords = c("x", "y"), crs = st_crs(bng))

# Define the value mapping for pred_damage
damage_values <- c("LOW" = 1, "LOW-MED" = 2, "MED" = 3, "MED-HIGH" = 4, "HIGH" = 5)
df_sf$value <- damage_values[df_sf$pred_damage]

# Rasterize
raster_data <- stars::st_rasterize(df_sf %>% dplyr::select(value, geometry))

# Get last 5 colours from YIOrRd color palette
color_palette <- c("LOW" = "yellow", "LOW-MED" = "#FED976", "MED" = "#FD8D3C", "MED-HIGH" = "#FC4E2A", "HIGH" = "#E31A1C")


  # Extract the i-th area
  area <- buffered_woods_7km$geometry
  
  cropped_data <- st_crop(raster_data, area)
  
  # Convert the cropped data to a data frame for ggplot2
  cropped_df <- as.data.frame(cropped_data, xy = TRUE)
  
  # Create a factor for the damage categories
  cropped_df$value <- factor(cropped_df$value, levels = 1:5, labels = names(damage_values))
  
  # Replace NAs with 0
  cropped_df <- cropped_df %>%
    filter(!is.na(value))
  
  # Plot using ggplot2
  plot_title <- buffered_woods_7km$Region
  
  plot <-ggplot(cropped_df) +
    geom_raster(aes(x = x, y = y, fill = value)) +
    scale_fill_manual(
      values = color_palette, 
      name = "Deer Impact Risk",
      breaks = names(damage_values),  # Ensure all levels are shown in the legend
      labels = names(damage_values)
    ) +
    theme_minimal() +
    coord_equal() +
    labs(
      title = plot_title,
      x = "Longitude",
      y = "Latitude"
    )

# Save each plot to a separate PDF file with 300 dpi resolution
output_dir <- here("output/figures")
dpi <- 300  # Resolution in dots per inch
width <- 6  # Width of the PDF in inches
height <- 6  # Height of the PDF in inches

  # Save the plot as a 300 dpi PDF
  ggsave(
    filename = "DI_risk", 
    plot = plot, 
    device = "pdf",  # Specify the file format
    width = width, 
    height = height, 
    dpi = dpi  # Set the resolution
  )


```

```{r}

#Adding the predicted damage data to an interactive satellite image ####
library(leaflet)

# Load the data
df_categorical <- readRDS(file = file.path(here("data/derived-data/Tool-extract-function-test/predicted_data_categorical_pixel_level_5km.rds")))

# Make sf object
df_sf <- st_as_sf(df_categorical, coords = c("x", "y"), crs = st_crs(bng))
df_sf <- st_transform(df_sf, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# Define the value mapping for pred_damage
damage_values <- c("LOW" = 1, "LOW-MED" = 2, "MED" = 3, "MED-HIGH" = 4, "HIGH" = 5)
df_sf$value <- damage_values[df_sf$pred_damage]

# Set up the colors
val <- 1:5
pal <- c("yellow", "#FED976", "#FD8D3C", "#FC4E2A", "#E31A1C")

#Convert to raster
r <- df_sf %>% dplyr::select(geometry, value) %>% stars::st_rasterize()

# Convert the raster to a data frame for ggplot
r_df <- as.data.frame(r, xy = TRUE)

# Create a factor for the value with all levels
r_df$value <- factor(r_df$value, levels = val, labels = names(damage_values))

# Add missing levels to the data
missing_levels <- data.frame(x = NA, y = NA, value = factor(val, labels = names(damage_values)))
r_df <- rbind(r_df, missing_levels)

# Plot raster using the color palette with custom labels
ggplot(r_df) +
  geom_raster(aes(x = x, y = y, fill = value)) +
  scale_fill_manual(values = pal, 
                    breaks = names(damage_values),
                    labels = names(damage_values),
                    na.value = "transparent") +
  theme_minimal() +
  labs(title = "Predicted Damage Raster",
       fill = "Damage Level") +
  guides(fill = guide_legend(override.aes = list(alpha = 1))) # Ensure all levels are shown in the legend

# Define the color palette function for leaflet
color_pal <- colorNumeric(palette = pal, domain = val, na.color = "transparent")

# Convert the stars object to a RasterLayer
raster_layer <- as(r, "Raster")

#Round all values to an integer
rounded_raster <- round(raster_layer)
rounded_raster <- raster::as.factor(rounded_raster)

#Project for leaflet map
rounded_raster_leaflet <- projectRasterForLeaflet(rounded_raster, method="ngb")

# Create a color palette for the raster
color_fact <- colorFactor(
  palette = c("yellow", "#FED976", "#FD8D3C", "#FC4E2A", "#E31A1C"),
  domain = damage_values,
na.color = "transparent"  # Set the color for NA values to transparent

)

# Create a reversed mapping for labels
damage_labels <- names(damage_values)
names(damage_labels) <- damage_values

# Create the Leaflet map
leaflet() %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addRasterImage(rounded_raster_leaflet, colors = color_fact, opacity = 0.9, project = FALSE) %>%
  addLegend(
    pal = color_fact,
    values = damage_values,  # Use the correct domain of values
    title = "Deer impact risk",
    labFormat = labelFormat(
      transform = function(x) {
        # Use the label mapping to get the labels
        damage_labels[as.character(x)]
      }
    )
  )

  
```

```{r}
#Save deer impact predictions as raster layer to export if needed

writeRaster(rounded_raster, here("FR-iDeer-tool-dev/test_5km_raster.tif"),overwrite=TRUE)

#make 3km raster for FR testing

#Buffer wood by 3km
buffered_woods_3km <- st_as_sf(st_buffer(woods, dist = 3000))
buffered_woods_3km  <- st_transform(buffered_woods_3km , crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

#Crop 5km raster to 3km

rounded_raster_3km <- crop(rounded_raster, buffered_woods_3km)
writeRaster(rounded_raster, here("FR-iDeer-tool-dev/test_3km_raster.tif"),overwrite=TRUE)

```

```{r}

#Plot the 4 incides as spatial layers to be included alongside the leaflet map ####

# Save each plot to a separate PDF file with 300 dpi resolution
output_dir <- here("output/figures")
dpi <- 300  # Resolution in dots per inch
width <- 8  # Width of the PDF in inches
height <- 6  # Height of the PDF in inches

# Load the data
df_categorical <- readRDS(file = file.path(here("data/derived-data/Tool-extract-function-test/predicted_data_categorical_pixel_level_5km.rds")))

# Make sf object
df_sf <- st_as_sf(df_categorical, coords = c("x", "y"), crs = st_crs(bng))
df_sf <- st_transform(df_sf, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

#1. DISTURBANCE INDEX

# Define the value mapping for nearest_road
disturb_values <- c("LOW" = 1, "MED" = 2, "HIGH"=3)
df_sf$value <- disturb_values[df_sf$pred_disturb]

# Set up the colors
val <- 1:3
pal <- c("lightblue", "dodgerblue", "royalblue4")

#Convert to raster
r <- df_sf %>% dplyr::select(geometry, value) %>% stars::st_rasterize()

# Convert the raster to a data frame for ggplot
r_df <- as.data.frame(r, xy = TRUE)

# Create a factor for the value with all levels
r_df$value <- factor(r_df$value, levels = val, labels = names(disturb_values))

# Add missing levels to the data
missing_levels <- data.frame(x = NA, y = NA, value = factor(val, labels = names(disturb_values)))
r_df <- rbind(r_df, missing_levels)

# Plot raster using the color palette with custom labels
disturb_plot <- ggplot(r_df) +
  geom_raster(aes(x = x, y = y, fill = value)) +
  scale_fill_manual(values = pal, 
                    breaks = names(disturb_values),
                    labels = names(disturb_values),
                    na.value = "transparent") +
  theme_minimal() +
  labs(title = NULL,
       fill = "Disturbance Index") +
  guides(fill = guide_legend(override.aes = list(alpha = 1))) # Ensure all levels are shown in the legend

# Save the plot as a 300 dpi PDF
  ggsave(
    filename = file.path(output_dir,"disturb_plot.pdf"), 
    plot = disturb_plot, 
    device = "pdf",  # Specify the file format
    width = width, 
    height = height, 
    dpi = dpi)  # Set the resolution
  
  #2. NUTRITIONAL LANDSCAPE INDEX ####

# Make sf object
df_sf <- st_as_sf(df_categorical, coords = c("x", "y"), crs = st_crs(bng))
df_sf <- st_transform(df_sf, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# Define the value mapping for nearest_road
nli_values <- c("LOW" = 1, "MED" = 2, "HIGH"=3)
df_sf$value <- disturb_values[df_sf$pred_NLI]

# Set up the colors
val <- 1:3
pal <- c("plum1", "magenta2", "purple")

#Convert to raster
r <- df_sf %>% dplyr::select(geometry, value) %>% stars::st_rasterize()

# Convert the raster to a data frame for ggplot
r_df <- as.data.frame(r, xy = TRUE)

# Create a factor for the value with all levels
r_df$value <- factor(r_df$value, levels = val, labels = names(nli_values))

# Add missing levels to the data
missing_levels <- data.frame(x = NA, y = NA, value = factor(val, labels = names(nli_values)))
r_df <- rbind(r_df, missing_levels)

# Plot raster using the color palette with custom labels
nli_plot <- ggplot(r_df) +
  geom_raster(aes(x = x, y = y, fill = value)) +
  scale_fill_manual(values = pal, 
                    breaks = names(nli_values),
                    labels = names(nli_values),
                    na.value = "transparent") +
  theme_minimal() +
  labs(title = NULL,
       fill = "Nutritional Landscape Index") +
  guides(fill = guide_legend(override.aes = list(alpha = 1))) # Ensure all levels are shown in the legend

# Save the plot as a 300 dpi PDF
  ggsave(
    filename = file.path(output_dir,"NLI_plot.pdf"), 
    plot = nli_plot, 
    device = "pdf",  # Specify the file format
    width = width, 
    height = height, 
    dpi = dpi)  # Set the resolution
  
  #3. NUTRITIONAL PIXEL INDEX ####

# Make sf object
df_sf <- st_as_sf(df_categorical, coords = c("x", "y"), crs = st_crs(bng))
df_sf <- st_transform(df_sf, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# Define the value mapping for nutritional patch index
woodland_q_values <- c("LOW" = 1, "MED" = 2, "HIGH"=3)
df_sf$value <- disturb_values[df_sf$woodland_q]

# Set up the colors
val <- 1:3
pal <- c("darkolivegreen2", "lightgreen", "forestgreen")

#Convert to raster
r <- df_sf %>% dplyr::select(geometry, value) %>% stars::st_rasterize()

# Convert the raster to a data frame for ggplot
r_df <- as.data.frame(r, xy = TRUE)

# Create a factor for the value with all levels
r_df$value <- factor(r_df$value, levels = val, labels = names(woodland_q_values))

# Add missing levels to the data
missing_levels <- data.frame(x = NA, y = NA, value = factor(val, labels = names(woodland_q_values)))
r_df <- rbind(r_df, missing_levels)

# Plot raster using the color palette with custom labels
woodland_q_plot <- ggplot(r_df) +
  geom_raster(aes(x = x, y = y, fill = value)) +
  scale_fill_manual(values = pal, 
                    breaks = names(woodland_q_values),
                    labels = names(woodland_q_values),
                    na.value = "transparent") +
  theme_minimal() +
  labs(title = NULL,
       fill = "Woodland Pixel Quality") +
  guides(fill = guide_legend(override.aes = list(alpha = 1))) # Ensure all levels are shown in the legend

# Save the plot as a 300 dpi PDF
  ggsave(
    filename = file.path(output_dir,"woodland_q_plot.pdf"), 
    plot = woodland_q_plot, 
    device = "pdf",  # Specify the file format
    width = width, 
    height = height, 
    dpi = dpi)  # Set the resolution
  
  
#4. THERMOREGULATION ####

# Make sf object
df_sf <- st_as_sf(df_categorical, coords = c("x", "y"), crs = st_crs(bng))
df_sf <- st_transform(df_sf, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# Define the value mapping for thermoregulation
thermoreg_values <- c("LOW" = 1, "MED" = 2, "HIGH"=3)
df_sf$value <- disturb_values[df_sf$pred_thermoreg]

# Set up the colors
val <- 1:3
pal <- c("coral1", "red1", "maroon")

#Convert to raster
r <- df_sf %>% dplyr::select(geometry, value) %>% stars::st_rasterize()

# Convert the raster to a data frame for ggplot
r_df <- as.data.frame(r, xy = TRUE)

# Create a factor for the value with all levels
r_df$value <- factor(r_df$value, levels = val, labels = names(thermoreg_values))

# Add missing levels to the data
missing_levels <- data.frame(x = NA, y = NA, value = factor(val, labels = names(thermoreg_values)))
r_df <- rbind(r_df, missing_levels)

# Plot raster using the color palette with custom labels
thermoreg_plot <- ggplot(r_df) +
  geom_raster(aes(x = x, y = y, fill = value)) +
  scale_fill_manual(values = pal, 
                    breaks = names(thermoreg_values),
                    labels = names(thermoreg_values),
                    na.value = "transparent") +
  theme_minimal() +
  labs(title = NULL,
       fill = "Thermoregulation Index") +
  guides(fill = guide_legend(override.aes = list(alpha = 1))) # Ensure all levels are shown in the legend

# Save the plot as a 300 dpi PDF
  ggsave(
    filename = file.path(output_dir,"thermoreg_plot.pdf"), 
    plot = thermoreg_plot, 
    device = "pdf",  # Specify the file format
    width = width, 
    height = height, 
    dpi = dpi)  # Set the resolution
  
  disturb_plot
  nli_plot
  woodland_q_plot
  thermoreg_plot
  
```

#END \####
