---
title: "iDeer Tool Function"
format: html
editor: visual

#User selects woodland
#Metrics are calculated for the focal woodland and all woodlands in surrounding 10km landscape
#Cut all rasters to 10km, buffer by 2km for spatial processes (12km total)
#Extract the metrics (roads, urban, connectivity etc)
#Run the BBN
#Loop this process for the 5 different landscapes
#Do this for the 5 DIA regions
#Write as a function
---

```{r}
library(plyr)
library(ggplot2)
library(dplyr)
library(sf)
library(raster)
library(terra)
library(here)
library(progress)
library(bnlearn)
here()
```

### Load spatial data derived for England and Wales

```{r}
bng <- 27700

#Load NFI/LCM2015 master map
map <- raster(here("data/derived-data/2015-spatial-layers/edge_core_raster_all_tiles_EW.tif"))
crs(map) <- bng

#Load main roads at 250m scale
#Summed density of A roads, B roads and motorways
all_roads <- raster(here("data/derived-data/2015-spatial-layers/all_roads_250_mosaic_all_tiles_corrected.tif"))
crs(all_roads) <- bng

#Load nearest urban
nearest_urban <- raster(here("data/derived-data/2015-spatial-layers/nearest_urban_raster_2015_all_tiles_corrected.tif"))
crs(nearest_urban) <- bng
nearest_suburb <- raster(here("data/derived-data/2015-spatial-layers/nearest_suburban_raster_2015_all_tiles_corrected.tif"))
crs(nearest_suburb) <- bng

#Load linear features at 1km scale
lf <- raster(here("data/derived-data/2015-spatial-layers/linear_features/lf_1000_mosaic_all_tiles_corrected.tif"))
crs(lf) <- bng

#Load 10k England-Wales tiles

uk10ew <- st_read(here("data/derived-data/10k_tiles_EW.shp"))

#Load MAX DAMS at 1km scale
#This is the focal.max.dams.1000.tif layer that has been cropped and masked to the 10km BNG England-Wales tiles using ArcGIS
dams <- raster(here("data/derived-data/2015-spatial-layers/Max_DAMS/max_dams_1kEW.tif"))
crs(dams) <- bng
#ggggrrr

#Read in woodland polygon dataset
#Combined NFI 2015 and LCM 2015 woodlands

hab_patches_all <- st_read(here("data/derived-data/nfi_lcm_overlaid_shapefile_woods_export.shp"))
st_crs(hab_patches_all) <- bng

head(hab_patches_all)

hab_patches_all <- hab_patches_all %>%
  rename(patch_ID = Id) %>%
  rename(patch_area = Shape_Area)

#NFI/LCM2015 master map, masked to hab_patches
#This map contains raster pixel values for woodlands only
#Made in ArcGIS
woods.r <- raster(here("data/derived-data/2015-spatial-layers/nfilcm_woods.tif"))
#woods.r <- projectRaster(woods.r, map)
```

### Choose woodlands

NEED TO CHANGE THIS TO BE ABLE TO DRAW A POLYGON INSIDE WOODLANDS SO THAT USER CAN SELECT THEIR LAND PARCEL

```{r}

#Choose five woodlands within the DIA priority areas to extract data for
#This is simulating the user selecting a woodland of interest

#Suffolk
suffolk_wood <- hab_patches_all %>% subset(patch_ID == 260977)
#Cumbria
cumbria_wood <- hab_patches_all %>% subset(patch_ID == 142765)
#South Wales 
sw_wood <- hab_patches_all %>% subset(patch_ID == 340972)
#North Wales
nw_wood <- hab_patches_all %>% subset(patch_ID == 188539)
#Oxford 
ox_wood <- hab_patches_all %>% subset(patch_ID == 370647)
#select all woodlands together
woods <- hab_patches_all %>%
  filter(patch_ID %in% c("260977", "142765", "340972", "188539", "370647"))
woods$Region <- c("Cumbria","North Wales","Suffolk","South Wales","Oxford")
woods <- st_as_sf(woods)

```

### Buffer woodlands for spatial data extraction

```{r}
#Buffer all woods by 10km
buffered_woods_10km <- st_as_sf(st_buffer(woods, dist = 10000))  # dist = 10km in meters
#Buffer all woods by 12km (10km landscape + 2km edge buffer)
buffered_woods_12km <- st_as_sf(st_buffer(woods, dist = 12000))  # dist = 10km in meters

#ensure crs of buffers is BNG
st_crs(buffered_woods_12km) <- bng
st_crs(buffered_woods_10km) <- bng

#Check points, woodlands buffers line up
plot(st_geometry(buffered_woods_12km), col = NA)
plot(st_geometry(buffered_woods_10km), col = "yellow", add = TRUE)
plot(st_geometry(woods), col = "black", add = TRUE)
#plot(st_geometry(wood_points), add = TRUE, col = "red")

```

### Extract raster variables for all woodlands in 10km buffer

Produce dataframes with pixel ID's from land cover map. This will allow calculation of pixel-level deer damage risk. Need to ensure there is a "patch_ID" column so that the connectivity values can be assigned to pixels within each patch (calculated at patch-level).

```{r}
#NEAREST URBAN EXTRACTION ####

extract_urban_rasters <- function(buffered_woods_10km, raster_layer, land_cover_raster, woodlands) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_10km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffer and crop all rasters
  for (i in seq_along(buffered_woods_10km$geometry)) {
    buffer_geometry <- buffered_woods_10km$geometry[i]
    pb$tick()  # Update the progress bar

    # Convert buffer geometry to a spatial object that raster can use
    buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
    
    # Get the cell numbers within the extent
pixel_ID <- cellsFromExtent(land_cover_raster, buffer_extent)

# Get the coordinates for these cells
cell_coords <- as.data.frame(xyFromCell(land_cover_raster, pixel_ID))

#Extract pixel values
pixel_values <- extract(raster_layer, cell_coords)

    # Create a data frame with the extracted values and cell indices
    vals <- data.frame(nearest_urban = pixel_values, 
                       patch_ID = buffered_woods_10km$patch_ID[i],
                       pixel_ID = pixel_ID,
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    vals <- vals[!is.na(vals$nearest_urban), ]
    
    # Append to the results list
    extraction_results[[i]] <- vals
  }
  
  # Combine all results into a single dataframe
  result_df <- bind_rows(extraction_results)
  
  return(result_df)
}

nearest_urban_extraction <- extract_urban_rasters(buffered_woods_10km=buffered_woods_10km, 
 raster_layer = nearest_urban,  
 woodlands = hab_patches_all,
 land_cover_raster = map)

write.csv(nearest_urban_extraction, here("data/derived-data/Tool-extract-function-test/nearest_urban_extraction_pixelIDs.csv"))
```

```{r}
#NEAREST SUBURBAN EXTRACTION ####
extract_suburban_rasters <- function(buffered_woods_10km, raster_layer, land_cover_raster, woodlands) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_10km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_10km$geometry)) {
  buffer_geometry <- buffered_woods_10km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
 # Get the cell numbers within the extent
pixel_ID <- cellsFromExtent(land_cover_raster, buffer_extent)

# Get the coordinates for these cells
cell_coords <- as.data.frame(xyFromCell(land_cover_raster, pixel_ID))

#Extract pixel values
pixel_values <- extract(raster_layer, cell_coords)

    # Create a data frame with the extracted values and cell indices
    vals <- data.frame(nearest_suburban = pixel_values, 
                       patch_ID = buffered_woods_10km$patch_ID[i],
                       pixel_ID = pixel_ID,
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    vals <- vals[!is.na(vals$nearest_suburban), ]
    
    # Append to the results list
    extraction_results[[i]] <- vals
  }
  
  # Combine all results into a single dataframe
  result_df <- bind_rows(extraction_results)
  
  return(result_df)
}

nearest_suburban_extraction <- extract_suburban_rasters(buffered_woods_10km=buffered_woods_10km, 
raster_layer=nearest_suburb,
woodlands=hab_patches_all,
land_cover_raster=map)

write.csv(nearest_suburban_extraction, here("data/derived-data/Tool-extract-function-test/nearest_suburban_extraction_pixelIDs.csv"))
```

```{r}
#MAX DAMS EXTRACTION ####
extract_dams_raster <- function(buffered_woods_10km, raster_layer, land_cover_raster, woodlands) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_10km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
    # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_10km$geometry)) {
  buffer_geometry <- buffered_woods_10km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
 # Get the cell numbers within the extent
pixel_ID <- cellsFromExtent(land_cover_raster, buffer_extent)

# Get the coordinates for these cells
cell_coords <- as.data.frame(xyFromCell(land_cover_raster, pixel_ID))

#Extract pixel values
pixel_values <- extract(raster_layer, cell_coords)

    # Create a data frame with the extracted values and cell indices
    vals <- data.frame(max_dams = pixel_values, 
                       patch_ID = buffered_woods_10km$patch_ID[i],
                       pixel_ID = pixel_ID,
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    vals <- vals[!is.na(vals$max_dams), ]
    
    # Append to the results list
    extraction_results[[i]] <- vals
  }
  
  # Combine all results into a single dataframe
  result_df <- bind_rows(extraction_results)
  
  return(result_df)
}

dams_extraction <- extract_dams_raster(
  buffered_woods_10km=buffered_woods_10km,
  raster_layer=dams, 
  woodlands=hab_patches_all,
  land_cover_raster=map)

#save as rds, too large to be saved as csv
saveRDS(dams_extraction, here("data/derived-data/Tool-extract-function-test/DAMS_MAX_extraction_1000_pixelIDs.rds"))
```

```{r}
#LINEAR FEATURES EXTRACTION ####
extract_lf_raster <- function(buffered_woods_10km, raster_layer, land_cover_raster, woodlands) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_10km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_10km$geometry)) {
  buffer_geometry <- buffered_woods_10km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
   # Get the cell numbers within the extent
pixel_ID <- cellsFromExtent(land_cover_raster, buffer_extent)

# Get the coordinates for these cells
cell_coords <- as.data.frame(xyFromCell(land_cover_raster, pixel_ID))

#Extract pixel values
pixel_values <- extract(raster_layer, cell_coords)

    # Create a data frame with the extracted values and cell indices
    vals <- data.frame(lf = pixel_values, 
                       patch_ID = buffered_woods_10km$patch_ID[i],
                       pixel_ID = pixel_ID,
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    vals <- vals[!is.na(vals$lf), ]
    
    # Append to the results list
    extraction_results[[i]] <- vals
  }
  
  # Combine all results into a single dataframe
  result_df <- bind_rows(extraction_results)
  
  return(result_df)
}

lf_extraction <- extract_lf_raster(buffered_woods_10km=buffered_woods_10km, 
raster_layer=lf, 
woodlands=hab_patches_all,
land_cover_raster=map)

write.csv(lf_extraction, here("data/derived-data/Tool-extract-function-test/linear_feature_extraction_pixelIDs_1000.csv"))
```

```{r}
#ROAD DENSITY EXTRACTION ####
extract_road_raster <- function(buffered_woods_10km, raster_layer, woodlands, land_cover_raster) {
  
  # Initialize progress bar
  pb <- progress_bar$new(
    format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
    total = nrow(buffered_woods_10km),
    clear = FALSE,   # Keep the progress bar on the screen
    width = 60
  )
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
  for (i in seq_along(buffered_woods_10km$geometry)) {
    buffer_geometry <- buffered_woods_10km$geometry[i]
    pb$tick()  # Update the progress bar

    # Convert buffer geometry to a spatial object that raster can use
    buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
    
     # Get the cell numbers within the extent
pixel_ID <- cellsFromExtent(land_cover_raster, buffer_extent)

# Get the coordinates for these cells
cell_coords <- as.data.frame(xyFromCell(land_cover_raster, pixel_ID))

#Extract pixel values
pixel_values <- extract(raster_layer, cell_coords)

    # Create a data frame with the extracted values and cell indices
    vals <- data.frame(roads = pixel_values, 
                       patch_ID = buffered_woods_10km$patch_ID[i],
                       pixel_ID = pixel_ID,
                       x = cell_coords$x,
                       y = cell_coords$y)
    
    # Filter out NA values
    vals <- vals[!is.na(vals$roads), ]
    
    # Append to the results list
    extraction_results[[i]] <- vals
  }
  
  # Combine all results into a single dataframe
  result_df <- bind_rows(extraction_results)
  
  return(result_df)
}

# Use the function and save the results
road_extraction <- extract_road_raster(buffered_woods_10km=buffered_woods_10km, 
raster_layer=all_roads, 
woodlands=hab_patches_all,
land_cover_raster=map)

write.csv(road_extraction, here("data/derived-data/Tool-extract-function-test/road_density_extraction_250_pixelIDs.csv"))

```

```{r}
#LAND COVER EXTRACTION FOR 1km AROUND WOODLANDS ####

polyFragStats <- function(POLY, veg, buffer_width = 1000) {
  # Create a buffer around the polygon
  buffer_poly <- buffer(POLY[i,], width = buffer_width, dissolve = TRUE)
  
  # Cut out the original polygon from the buffer
  buffer_area <- erase(buffer_poly, POLY[i,])
  
  # Crop land cover map by the buffer area
  fcrop <- crop(veg, buffer_area)
  
  # Mask the cropped land cover map
  fmask <- mask(fcrop, buffer_area)
  
  # Get the cell numbers within the extent
pixel_ID <- cellsFromExtent(veg, buffer_extent)

# Get the coordinates for these cells
cell_coords <- as.data.frame(xyFromCell(veg, pixel_ID))

#Extract pixel values
pixel_values <- extract(fmask, cell_coords)
  
  # Create a data frame with the extracted values and cell indices
  vals <- data.frame(pixel_ID = pixel_ID,
                     lc_class = pixel_values,
                     patch_ID = POLY@data$patch_ID[i],
                     x = cell_coords$x,
                    y = cell_coords$y)

  # Filter out NA values
  vals <- vals[!is.na(vals$lc_class), ]
  
  return(vals)
}

# Filter for woodland patches that intersect buffers
woods_buffers <- st_filter(hab_patches_all, buffered_woods_10km, .predicate = st_intersects)

# Convert to spatial object
POLY <- as(woods_buffers[1:50,], "Spatial")
veg <- map

# Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = length(POLY),
  clear = FALSE,
  width = 60
)

# Process each polygon and collect results
respoly <- ldply(1:length(POLY), function(i) {  # for each polygon
  pb$tick()  # Update the progress bar
  out <- polyFragStats(POLY[i,], veg = veg)  # compute metrics
  out$buffer_size <- "b1000"
  return(out)
})

write.csv(respoly, here("data/derived-data/Tool-extract-function-test/landscapemetrics_1000_pixelIDs.csv"))

```

```{r}
#WOODLAND PIXEL EXTRACTION ####

woodpixels <- function(POLY, veg) {
  
  # Crop land cover map by the woodland polygon
  fcrop <- crop(veg, extent(POLY[i,]))
  
  # Mask the cropped land cover map
  fmask <- mask(fcrop, POLY[i,])
  
  # Get the cell numbers within the extent
pixel_ID <- cellsFromExtent(veg, buffer_extent)

# Get the coordinates for these cells
cell_coords <- as.data.frame(xyFromCell(veg, pixel_ID))

#Extract pixel values
pixel_values <- extract(fmask, cell_coords)
  
  # Create a data frame with the extracted values and cell indices
  vals <- data.frame(pixel_ID = pixel_ID,
                     lc_class = pixel_values,
                     patch_ID = POLY@data$patch_ID[i],
                     x = cell_coords$x,
                    y = cell_coords$y)

  # Filter out NA values
  vals <- vals[!is.na(vals$lc_class), ]
  
  return(vals)
}

# Filter for woodland patches that intersect buffers
woods_buffers <- st_filter(hab_patches_all, buffered_woods_10km, .predicate = st_intersects)

# Convert to spatial object
POLY <- as(woods_buffers[1:50,], "Spatial")
veg <- map

# Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = length(POLY),
  clear = FALSE,
  width = 60
)

# Process each polygon and collect results
woodpix <- ldply(1:length(POLY), function(i) {  # for each polygon
  pb$tick()  # Update the progress bar
  out <- woodpixels(POLY[i,], veg = veg)  # compute metrics
  return(out)
})

write.csv(woodpix, here("data/derived-data/Tool-extract-function-test/woodland_type_pixelIDs.csv"))
```

### Calculate connectivity for every woodland inside the 10km buffers

```{r}
#### SET CONNECTIVITY PARAMETERS ####
# connectivity parametetrs - % of dispersers reaching a set distance #
percentage_dispersers <- 0.05 ### 5%
dispersal_distance <- 434 #Gets curve up to 1000m on x axis
dispersal_contribution <-
  -((log(1 / percentage_dispersers)) / dispersal_distance)

# Buffer distance represents the cut off - this stops the script measuring every pairwise combination
dispersal_cutoff <- 0.999 ### 99.9% cut off
#buffer_cutoff <-
  #round(log(1 / (1 - dispersal_cutoff)) / (log(1 / percentage_dispersers) /
                                             #dispersal_distance), digits = 2)
buffer_cutoff = 1000 #1km

# test & plot of dispersal contribution for sequence of distance up to buffer_cutoff
test <-
  data.frame(distance = (seq(0, buffer_cutoff, by = 0.5))) %>%
  mutate(contribution = (exp(dispersal_contribution * distance) * 100))

ggplot(data = test, aes(x = distance, y = contribution, group = 1)) +
  geom_line() +
  geom_point() +
  geom_vline(xintercept = buffer_cutoff, color = "red")

#See how buffer size looks
focal_patch_buffer <- st_buffer(hab_patches_all[1,], buffer_cutoff)
plot(st_geometry(focal_patch_buffer))
plot(st_geometry(hab_patches_all[1,]),add=TRUE)

```

```{r}

connectivity_results <- list()

  # Use the 12km buffer as the source patches area
  source_woods <- st_filter(hab_patches_all, buffered_woods_12km, .predicate = st_intersects)

  # Crop the woodlands in the original 10km square to get the focal patches
  focal_woods <- st_filter(hab_patches_all, buffered_woods_10km, .predicate = st_intersects)
  n_distinct(focal_woods$patch_ID)
  #[1] 7884

  # Calculate connectivity for each focal woodland
  Connectivity_table <- NULL
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(focal_woods),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)

  # Loop through each focal woodland to calculate connectivity
  for (j in 1:nrow(focal_woods[1:50,])) {
    
    # Focal patch in the 10km square
    focal_patch <- focal_woods[j, ]
    
    #update progress bar
    pb$tick
    
    # Buffer around the focal patch to include source patches
    focal_patch_buffer <- st_buffer(focal_patch, buffer_cutoff)
    
    # Find source patches within this buffer
    source_patches <- st_filter(source_woods, focal_patch_buffer, .predicate = st_intersects)
    
    # Calculate distances between the focal and source patches
    patch_dist <- as.vector(st_distance(focal_patch, source_patches))
    
    # Create a data frame for connectivity information
    Conn_table_site <- data.frame(
      focal_patch = focal_patch$patch_ID,
      focal_patch_area = as.numeric(focal_patch$patch_area),
      source_patch = source_patches$patch_ID,
      source_patch_area = as.numeric(source_patches$patch_area),
      distance = patch_dist,
      incoming_connect = ifelse(
        focal_patch$patch_ID == source_patches$patch_ID,
        NA,
        source_patches$patch_area * exp(dispersal_contribution * patch_dist)
      )
    )
    
    # Append to 'Connectivity_table'
    if (is.null(Connectivity_table)) {
      Connectivity_table <- Conn_table_site
    } else {
      Connectivity_table <- rbind(Connectivity_table, Conn_table_site)
    }
  
  n_distinct(Connectivity_table$focal_patch)

  # Store results for this woodland
  connectivity_results[[paste("focal_wood", focal_woods$patch_ID[i])]] <- Connectivity_table
}
```

```{r}

#Remove rows where focal_patch = source_patch
Connectivity_table_filt <- Connectivity_table %>%
  filter(focal_patch != source_patch)

# Identify missing unique patch IDs
# These patches have no source patches so will have been filtered out
# We need to put them back in so we can give them an incoming_connect value of 0
missing_patch_ids <- setdiff(
     unique(Connectivity_table$focal_patch),
     unique(Connectivity_table_filt$focal_patch)
 )
 
 print(missing_patch_ids)  # This will show which patch IDs are missing
#[1] 132853 133946 141304 260769 358349 359726
 
missing_patches <- Connectivity_table %>% filter(focal_patch %in% missing_patch_ids) %>%
  # Replace NAs in 'incoming_connect' with 0
  mutate(incoming_connect = coalesce(incoming_connect, 0))  # Replace NA with 0

#Make rows for patches with no connectivity so we don't lose any patches
Connectivity_table_filt <- rbind(Connectivity_table_filt, missing_patches)

#Ensure no duplicate rows
Connectivity_table_filt <- Connectivity_table_filt %>% distinct()

#Now no patches should be missing
n_distinct(Connectivity_table_filt$focal_patch)

#sum incoming connectivity by focal patch
incoming_connectivity_sum <- Connectivity_table_filt %>%
  dplyr::select(focal_patch, incoming_connect) %>%
  group_by(focal_patch) %>%
  dplyr::summarise(total_connect = sum(incoming_connect),
            n = n()) %>%
  mutate(n = if_else(total_connect == 0, 0, n)) %>%  # Set 'n' to 0 when 'incoming_connect' is 0
  rename(patch_ID = focal_patch)

#save incoming connectivity table

write.csv(incoming_connectivity_sum, here("data/derived-data/Tool-extract-function-test/incoming_connectivity_sum.csv"))
```

### Ready the spatial data for the Bayesian Belief Network

#### Calculate Nutritional Landscape Composition (at 1km scale)

```{r}
#CALCULATE NUTRITIONAL LANDSCAPE QUALITY ####

lcm <- respoly
lcm <- lcm%>%rename(class = lc_class)
#Load landscape nutritional scores for raster categories
#Edge forage quality scores sum of woodland score + adjacent land cover score
#Water scores retained to prevent loss of woodland edge pixels
#Water edge pixels receive one point higher than regular woodland
#Still an edge so will be highly nutritious
edge.val.types <- read.csv(file.path(here("data/raw-data/Expert_Qnaire/Expert_derived_quality_ranks_edge_types_with_water.csv")))
edge.val.types <- edge.val.types%>%dplyr::select(-"X")

#Stitch the quality scores into the lcm dataset

lc.edge.types <- lcm %>% left_join(edge.val.types, by = "class") %>% mutate(Forage_Q = ifelse(is.na(Forage_Q), 0, Forage_Q))%>%
  rename(scale = buffer_size)
#%>%filter(!is.na(Land.cover))

```

#### Assign pixel quality values

```{r}

woods <- woodpix
woods <- woods%>%rename(class = lc_class)
unique(woods$class)

#Load landscape nutritional scores for raster categories
#Edge forage quality scores sum of woodland score + adjacent land cover score
#Water scores added in to prevent loss of woodland edge pixels
#Water edge pixels receive one point higher than regular woodland
#Still an edge so will be highly nutritious
edge.val.types <- read.csv(file.path(here("data/raw-data/Expert_Qnaire/Expert_derived_quality_ranks_edge_types_with_water.csv")))
edge.val.types <- edge.val.types%>%dplyr::select(-"X")

#Check edge.val.types$class and pcm$class match in format
unique(edge.val.types$class)
class(edge.val.types$class)
edge.val.types <- edge.val.types%>%arrange(class)
unique(edge.val.types$class)

class(woods$class)
pcm <-woods%>%arrange(class)
unique(pcm$class)

# Round to 3 decimal places
#pcm$class <- round(pcm$class, 3)
#edge.val.types$class <- round(edge.val.types$class, 3)

#Stitch the quality scores into the lcm dataset

wood.edge.types <- woods %>% left_join(edge.val.types, by = c("class")) #%>%
                                          #filter(!is.na(Land.cover))

#Replace "NA" in Forage_Q column with 0
wood.edge.types <- wood.edge.types%>% 
 mutate(Forage_Q = ifelse(is.na(Forage_Q), 0, Forage_Q))%>%
  mutate(scale = "patch")

```

#### left_join the data sets together

```{r}

#Forage quality of pixels
df <- rbind(lc.edge.types, wood.edge.types)

#Connectivity ####
incoming_connectivity_sum<-incoming_connectivity_sum %>%
  mutate(n = NULL)
df <- left_join(df, incoming_connectivity_sum, by = c("patch_ID"))

#MAX DAMS ####
df <- left_join(df, dams_extraction, by=c("patch_ID","pixel_ID","x","y")) 

#Nearest urban and suburban ####
df <- left_join(df, nearest_suburban_extraction, by=c("patch_ID","pixel_ID","x","y"))

df <- left_join(df, nearest_urban_extraction, by=c("patch_ID","pixel_ID","x","y"))

#Make column for nearest urban and suburban using minimum value
df$urban_proximity <- pmin(df$nearest_suburban, df$nearest_urban)
#Drop individual columns
df <- df%>%dplyr::select(-c("nearest_suburban","nearest_urban"))

saveRDS(df, here("data/derived-data/Tool-extract-function-test/bbn_test_df_5_patches.rds"))
```

#### Convert numerical values to categorical (LOW, MED, HIGH)

```{r}

#Replace values in numerical columns with LOW, MED, HIGH


df <- readRDS(here("data/derived-data/Tool-extract-function-test/bbn_test_df_5_patches.rds"))

#Look at distribution of values across national datasets ####
#We need to decide at what point values change between categories- need to do this using whole datasets rather than localised cropped datasets.

#Roads
road_vals <- raster::values(all_roads)
road_vals <- road_vals[!is.na(road_vals)]
hist(road_vals)
df$all_roads_250 <- ifelse(df$all_roads_250 == 0, "LOW",
            ifelse(df$all_roads_250 >0 & df$all_roads_250 <= 1000, "MED",
            ifelse(df$all_roads_250> 1000, "HIGH", NA)))
unique(df$all_roads_250)

#Linear features

lf_vals <-raster::values(lf)
lf_vals <- lf_vals[!is.na(lf_vals)]
hist(lf_vals)
# Replace numeric values with categorical labels
df$lf_1000 <- ifelse(df$lf_1000<= 5000, "LOW",
            ifelse(df$lf_1000> 5000 & df$lf_1000 <= 10000, "MED",
            ifelse(df$lf_1000> 10000, "HIGH", NA)))
unique(df$lf_1000)


#nearest urban/suburban
#Proximity, therefore "HIGH" means lower value (lower distance)
hist(df$urban_proximity)
df$urban_proximity <- ifelse(df$urban_proximity <= 500, "HIGH",
                          ifelse(df$urban_proximity >500 & df$urban_proximity <= 1000, "MED",
                           ifelse(df$urban_proximity >1000,"LOW",NA)))  
unique(df$urban_proximity)

#MAX DAMS
hist(df$focal.dams.1000.max)
df$focal.dams.1000.max <- ifelse(df$focal.dams.1000.max <=10,"LOW",
                      ifelse(df$focal.dams.1000.max>10 & df$focal.dams.1000.max <= 15, "MED",
                        ifelse(df$focal.dams.1000.max >15, "HIGH",NA)))

#nutritional landscape composition score
hist(df$nli_edge_types_1000)
df$nli_edge_types_1000 <- ifelse(df$nli_edge_types_1000 <= 0, "LOW",
                          ifelse(df$nli_edge_types_1000 >0 & df$nli_edge_types_1000 <=4, "MED",
                           ifelse(df$nli_edge_types_1000 > 4, "HIGH",NA)))

#nutritional patch composition score

hist(df$npi_edge_types)

df$npi_edge_types <- ifelse(df$npi_edge_types <= 5, "LOW",
                          ifelse(df$npi_edge_types >5 & df$npi_edge_types <=10, "MED",
                           ifelse(df$npi_edge_types > 10, "HIGH",NA)))
unique(df$npi_edge_types)

#connectivity
hist(df$total_connect)
quantile(df$total_connect)

#log transform connectivity values, they are too big for easy interpretation

df$total_connect_log <- log(df$total_connect)

hist(df$total_connect_log)

df$total_connect_log <- ifelse(df$total_connect_log <=10, "LOW",
                          ifelse(df$total_connect_log >10 & df$total_connect_log  <=15, "MED",
                           ifelse(df$total_connect_log  > 15, "HIGH",NA)))

#remove original connectivity column

df <- df%>%dplyr::select(-c("total_connect","n"))

#Export

df_categorical <- df

saveRDS(df_categorical, file = file.path(here("data/derived-data/Tool-extract-function-test/tool_test_extracted_data_categorical.rds")))

```

### Create the Bayesian Belief Network

#### Initialize Conditional Probability Tables

```{r}
#CPTS for measured nodes ####
urban_prox_cpt<-matrix(c(0.333333,0.333333,0.333333), ncol=3, dimnames = list(NULL, c("LOW", "MED","HIGH")))

road_density_cpt<-matrix(c(0.333333,0.333333,0.333333), ncol=3, dimnames = list(NULL, c("LOW", "MED","HIGH")))

connectivity_cpt<-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

landscape_nutritional_score_cpt<-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

patch_nutritional_score_cpt <- matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

dams_cpt <- landscape_nutritional_score_cpt<-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

lf_density_cpt <-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))
```

```{r}
#CPT for Disturbance Index ####

disturb_index_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3, 3, 3),  # Shape of the array (3x3x3)
  dimnames = list(
    disturb_index = c("LOW", "MED", "HIGH"),
    road_density = c("LOW", "MED", "HIGH"),
    urban_prox = c("LOW","MED","HIGH"),
    connectivity = c("LOW", "MED", "HIGH")
  )
)
#order = disturb_index, road_density, urban_prox,connectivity

#table1: urban_prox = LOW, connectivity = LOW
disturb_index_cpt["LOW","LOW","LOW","LOW"] <- 1
disturb_index_cpt["MED","MED","LOW","LOW"] <- 1
disturb_index_cpt["HIGH","HIGH","LOW","LOW"] <- 1
#table2:  urban_prox = MED, connectivity = LOW
disturb_index_cpt["MED","LOW","MED","LOW"] <- 1
disturb_index_cpt["MED","MED","MED","LOW"] <- 1
disturb_index_cpt["HIGH","HIGH","MED","LOW"] <- 1
#table3: urban_prox = HIGH, connectivity = LOW
disturb_index_cpt["HIGH","LOW","HIGH","LOW"] <- 1
disturb_index_cpt["HIGH","MED","HIGH","LOW"] <- 1
disturb_index_cpt["HIGH","HIGH","HIGH","LOW"] <- 1
#table4: urban_prox = LOW, connectivity = MED
disturb_index_cpt["LOW","LOW","LOW","MED"] <- 1
disturb_index_cpt["MED","MED","LOW","MED"] <- 1
disturb_index_cpt["MED","HIGH","LOW","MED"] <- 1
#table5: urban_prox = MED, connectivity = MED
disturb_index_cpt["LOW","LOW","MED","MED"] <- 1
disturb_index_cpt["MED","MED","MED","MED"] <- 1
disturb_index_cpt["HIGH","HIGH","MED","MED"] <- 1
#table6: urban_prox = HIGH, connectivity = MED
disturb_index_cpt["HIGH","LOW","HIGH","MED"] <- 1
disturb_index_cpt["HIGH","MED","HIGH","MED"] <- 1
disturb_index_cpt["HIGH","HIGH","HIGH","MED"] <- 1
#table7: urban_prox = LOW, connectivity = HIGH
disturb_index_cpt["LOW","LOW","LOW","HIGH"] <- 1
disturb_index_cpt["LOW","MED","LOW","HIGH"] <- 1
disturb_index_cpt["MED","HIGH","LOW","HIGH"] <- 1
#table8: urban_prox = MED, connectivity = HIGH
disturb_index_cpt["MED","LOW","MED","HIGH"] <- 1
disturb_index_cpt["MED","MED","MED","HIGH"] <- 1
disturb_index_cpt["HIGH","HIGH","MED","HIGH"] <- 1
#table9: urban_prox = HIGH, connectivity = HIGH
disturb_index_cpt["MED","LOW","HIGH","HIGH"] <- 1
disturb_index_cpt["MED","MED","HIGH","HIGH"] <- 1
disturb_index_cpt["HIGH","HIGH","HIGH","HIGH"] <- 1

disturb_index_cpt

```

```{r}
#CPT for Nutritional Patch Index ####

NPI_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3),  # Shape of the array (3x3)
  dimnames = list(
    NPI = c("LOW", "MED", "HIGH"),
    patch_nutritional_score = c("LOW", "MED", "HIGH")
  )
)

NPI_cpt["LOW","LOW"] <- 1
NPI_cpt["MED","MED"] <- 1
NPI_cpt["HIGH","HIGH"] <- 1

NPI_cpt
```

```{r}
#CPT for Nutritional Landscape Index ####


NLI_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3, 3),  # Shape of the array (3x3)
  dimnames = list(
    NLI = c("LOW", "MED", "HIGH"),
    landscape_nutritional_score = c("LOW", "MED", "HIGH"),
    lf_density = c("LOW","MED","HIGH")
  )
)
#order = NLI, landscape_nutritional_score, lf_density
NLI_cpt["LOW","LOW","LOW"] <- 1
NLI_cpt["MED","MED","LOW"] <- 1
NLI_cpt["HIGH","HIGH","LOW"] <- 1
NLI_cpt["LOW","LOW","MED"] <- 1
NLI_cpt["HIGH","MED","MED"] <- 1
NLI_cpt["HIGH","HIGH","MED"] <- 1
NLI_cpt["MED","LOW","HIGH"] <- 1
NLI_cpt["HIGH","MED","HIGH"] <- 1
NLI_cpt["HIGH","HIGH","HIGH"] <- 1

NLI_cpt
```

```{r}
#CPT for Thermoregulation Index ####

thermoreg_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3),  # Shape of the array
  dimnames = list(
    thermoreg_index = c("LOW", "MED", "HIGH"),
    DAMS_MAX = c("LOW", "MED", "HIGH")
  )
)
#order = DAMS_MAX, thermoreg_index
thermoreg_cpt["LOW","LOW"] <- 1
thermoreg_cpt["MED","MED"] <- 1
thermoreg_cpt["HIGH","HIGH"] <- 1

thermoreg_cpt
```

```{r}
#Final CPT: deer damage risk

damage_risk_cpt <- array(
  0,  # Default probability for each cell
  dim = c(5, 3, 3, 3, 3),  # Shape of the array
  dimnames = list(
    damage_risk = c("LOW","LOW-MED","MED","MED-HIGH","HIGH"),
    NPI = c("LOW", "MED", "HIGH"),
    NLI = c("LOW", "MED", "HIGH"),
    thermoreg_index = c("LOW", "MED", "HIGH"),
    disturb_index = c("LOW", "MED", "HIGH")
    
  )
)
#order = damage_risk, patch_quality,landscape_nutrition,thermoreg_index,disturbance
#table1 landscape_nutrition = LOW, thermoreg_index = LOW, disturbance_index = LOW
damage_risk_cpt["LOW-MED","LOW","LOW","LOW","LOW"] <- 1
damage_risk_cpt["LOW-MED","MED","LOW","LOW","LOW"] <- 1
damage_risk_cpt["MED","HIGH","LOW","LOW","LOW"] <- 1
#table2 landscape_nutrition = MED, thermoreg_index = LOW, disturbance_index = LOW
damage_risk_cpt["LOW-MED","LOW","MED","LOW","LOW"] <- 1
damage_risk_cpt["LOW-MED","MED","MED","LOW","LOW"] <- 1
damage_risk_cpt["MED","HIGH","MED","LOW","LOW"] <- 1
#table3 landscape_nutrition = HIGH, thermoreg_index = LOW, disturbance_index = LOW
damage_risk_cpt["LOW-MED","LOW","HIGH","LOW","LOW"] <- 1
damage_risk_cpt["MED","MED","HIGH","LOW","LOW"] <- 1
damage_risk_cpt["MED","HIGH","HIGH","LOW","LOW"] <- 1
#table4 landscape_nutrition = LOW, thermoreg_index = MED, disturbance_index = LOW
damage_risk_cpt["MED","LOW","LOW","MED","LOW"] <- 1
damage_risk_cpt["MED","MED","LOW","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","LOW","MED","LOW"] <- 1
#table5 landscape_nutrition = MED, thermoreg_index = MED, disturbance_index = LOW
damage_risk_cpt["MED","LOW","MED","MED","LOW"] <- 1
damage_risk_cpt["MED","MED","MED","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","MED","MED","LOW"] <- 1
#table6 landscape_nutrition = HIGH, thermoreg_index = MED, disturbance_index = LOW
damage_risk_cpt["MED","LOW","HIGH","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","MED","HIGH","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","HIGH","MED","LOW"] <- 1
#table7  landscape_nutrition = LOW, thermoreg_index = HIGH, disturbance_index = LOW
damage_risk_cpt["MED-HIGH","LOW","LOW","HIGH","LOW"] <- 1
damage_risk_cpt["MED-HIGH","MED","LOW","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","HIGH","LOW","HIGH","LOW"] <- 1
#table8 landscape_nutrition = MED, thermoreg_index = HIGH, disturbance_index = LOW
damage_risk_cpt["MED-HIGH","LOW","MED","HIGH","LOW"] <- 1
damage_risk_cpt["MED-HIGH","MED","MED","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","HIGH","MED","HIGH","LOW"] <- 1
#table9 landscape_nutrition = HIGH, thermoreg_index = HIGH, disturbance_index = LOW
damage_risk_cpt["MED-HIGH","LOW","HIGH","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","MED","HIGH","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","HIGH","HIGH","HIGH","LOW"] <- 1
#table10 landscape_nutrition = LOW, thermoreg_index = LOW, disturbance_index = MED
damage_risk_cpt["LOW","LOW","LOW","LOW","MED"] <- 1
damage_risk_cpt["LOW","MED","LOW","LOW","MED"] <- 1
damage_risk_cpt["LOW-MED","HIGH","LOW","LOW","MED"] <- 1
#table11 landscape_nutrition = MED, thermoreg_index = LOW, disturbance_index = MED
damage_risk_cpt["LOW-MED","LOW","MED","LOW","MED"] <- 1
damage_risk_cpt["LOW-MED","MED","MED","LOW","MED"] <- 1
damage_risk_cpt["MED","HIGH","MED","LOW","MED"] <- 1
#table12 landscape_nutrition = HIGH, thermoreg_index = LOW, disturbance_index = MED
damage_risk_cpt["LOW-MED","LOW","HIGH","LOW","MED"] <- 1
damage_risk_cpt["MED","MED","HIGH","LOW","MED"] <- 1
damage_risk_cpt["MED","HIGH","HIGH","LOW","MED"] <- 1
#table13 landscape_nutrition = LOW, thermoreg_index = MED, disturbance_index = MED
damage_risk_cpt["LOW-MED","LOW","LOW","MED","MED"] <- 1
damage_risk_cpt["LOW-MED","MED","LOW","MED","MED"] <- 1
damage_risk_cpt["MED","HIGH","LOW","MED","MED"] <- 1
#table14 landscape_nutrition = MED, thermoreg_index = MED, disturbance_index = MED
damage_risk_cpt["MED","LOW","MED","MED","MED"] <- 1
damage_risk_cpt["MED","MED","MED","MED","MED"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","MED","MED","MED"] <- 1
#table15  landscape_nutrition = HIGH, thermoreg_index = MED, disturbance_index = MED
damage_risk_cpt["MED","LOW","HIGH","MED","MED"] <- 1
damage_risk_cpt["MED-HIGH","MED","HIGH","MED","MED"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","HIGH","MED","MED"] <- 1
#table16 landscape_nutrition = LOW, thermoreg_index = HIGH, disturbance_index = MED
damage_risk_cpt["MED","LOW","LOW","HIGH","MED"] <- 1
damage_risk_cpt["MED","MED","LOW","HIGH","MED"] <- 1
damage_risk_cpt["MED","HIGH","LOW","HIGH","MED"] <- 1
#table17 landscape_nutrition = MED, thermoreg_index = HIGH, disturbance_index = MED
damage_risk_cpt["MED-HIGH","LOW","MED","HIGH","MED"] <- 1
damage_risk_cpt["MED-HIGH","MED","MED","HIGH","MED"] <- 1
damage_risk_cpt["HIGH","HIGH","MED","HIGH","MED"] <- 1
#table18 landscape_nutrition = HIGH, thermoreg_index = HIGH, disturbance_index = MED
damage_risk_cpt["MED-HIGH","LOW","HIGH","HIGH","MED"] <- 1
damage_risk_cpt["HIGH","MED","HIGH","HIGH","MED"] <- 1
damage_risk_cpt["HIGH","HIGH","HIGH","HIGH","MED"] <- 1
#table19 landscape_nutrition = LOW, thermoreg_index = LOW, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","LOW","LOW","HIGH"] <- 1
damage_risk_cpt["LOW","MED","LOW","LOW","HIGH"] <- 1
damage_risk_cpt["LOW","HIGH","LOW","LOW","HIGH"] <- 1
#table20 landscape_nutrition = MED, thermoreg_index = LOW, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","MED","LOW","HIGH"] <- 1
damage_risk_cpt["LOW","MED","MED","LOW","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","MED","LOW","HIGH"] <- 1
#table21 landscape_nutrition = HIGH, thermoreg_index = LOW, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","HIGH","LOW","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","HIGH","LOW","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","HIGH","LOW","HIGH"] <- 1
#table22 landscape_nutrition = LOW, thermoreg_index = MED, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","LOW","MED","HIGH"] <- 1
damage_risk_cpt["LOW","MED","LOW","MED","HIGH"] <- 1
damage_risk_cpt["LOW","HIGH","LOW","MED","HIGH"] <- 1
#table23 landscape_nutrition = MED, thermoreg_index = MED, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","MED","MED","HIGH"] <- 1
damage_risk_cpt["LOW","MED","MED","MED","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","MED","MED","HIGH"] <- 1
#table24 landscape_nutrition = HIGH, thermoreg_index = MED, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","HIGH","MED","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","HIGH","MED","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","HIGH","MED","HIGH"] <- 1
#table25  landscape_nutrition = LOW, thermoreg_index = HIGH, disturbance_index = HIGH
damage_risk_cpt["LOW-MED","LOW","LOW","HIGH","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","LOW","HIGH","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","LOW","HIGH","HIGH"] <- 1
#table26 landscape_nutrition = MED, thermoreg_index = HIGH, disturbance_index = HIGH
damage_risk_cpt["LOW-MED","LOW","MED","HIGH","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","MED","HIGH","HIGH"] <- 1
damage_risk_cpt["MED","HIGH","MED","HIGH","HIGH"] <- 1
#table27 landscape_nutrition = HIGH, thermoreg_index = HIGH, disturbance_index = HIGH
damage_risk_cpt["LOW-MED","LOW","HIGH","HIGH","HIGH"] <- 1
damage_risk_cpt["MED","MED","HIGH","HIGH","HIGH"] <- 1
damage_risk_cpt["MED","HIGH","HIGH","HIGH","HIGH"] <- 1

damage_risk_cpt
```

#### Create BBN structure

```{r}
# Step 1: Explicitly define the nodes in the network
nodes <- c(
  "disturb_index",
  "urban_prox",
  "road_density",
  "connectivity",
  "patch_nutritional_score",
  "landscape_nutritional_score",
  "NLI",
  "lf_density",
  "thermoreg_index",
  "DAMS_MAX",
  "damage_risk",
  "NPI"
)

#empty graph
e = empty.graph(nodes)

arc.set = matrix(c("urban_prox", "disturb_index",
                     "road_density", "disturb_index", 
                     "connectivity", "disturb_index",
                     "lf_density","NLI",
                     "landscape_nutritional_score","NLI",
                     "patch_nutritional_score","NPI",
                     "DAMS_MAX","thermoreg_index",
                     "disturb_index","damage_risk",
                     "NLI","damage_risk",
                     "NPI","damage_risk",
                     "thermoreg_index","damage_risk"),
             ncol = 2, byrow = TRUE,
             dimnames = list(NULL, c("from", "to")))

arcs(e) <- arc.set

model_string <- modelstring(e) 

net<-model2network(model_string) 
# You dont need to repeat the names of the nodes with parents (child nodes) 

# Custom fitting network (matching up the nodes to their CPTs)
dfit = custom.fit(net, dist = list(road_density=road_density_cpt, 
                                   urban_prox=urban_prox_cpt,
                                   connectivity=connectivity_cpt,
                                   lf_density=lf_density_cpt,
                                   DAMS_MAX=dams_cpt,
                      patch_nutritional_score=patch_nutritional_score_cpt,
                      landscape_nutritional_score=landscape_nutritional_score_cpt, 
                                   disturb_index=disturb_index_cpt, 
                                   NPI=NPI_cpt,
                                   NLI=NLI_cpt,
                                   thermoreg_index=thermoreg_cpt,
                                   damage_risk=damage_risk_cpt))

```

#### Plot BN structure

```{r plot the network}
#| echo: true
#| warning: false
#| message: false

graphviz.plot(net)

```

#### Check Parameters

```{r look at the dfit}
#| echo: true
#| warning: false
#| message: false
dfit
```

#### Set evidence

```{r set evidence}
#| echo: true
#| warning: false
#| message: false
cpquery(dfit, c(lf_density == "LOW"|DAMS_MAX == "HIGH"|landscape_nutritional_score == "HIGH"|connectivity=="LOW"|urban_prox=="LOW"|road_density == "HIGH"|DAMS_MAX=="HIGH"), (damage_risk=="HIGH"))

```

### Make predictions from BBN

```{r}
predict_dat <- df_categorical %>%st_drop_geometry() %>% #drop geometry
   mutate(across(where(is.character), toupper)) %>% #convert to upper
   dplyr::select(-c("patch_ID","patch_area"))
#Ensure variable names match those in BBN
predict_dat<-predict_dat%>%rename(
                              road_density=all_roads_250,
                              lf_density=lf_1000,
                              urban_prox=urban_proximity,
                              landscape_nutritional_score=nli_edge_types_1000,
                              patch_nutritional_score=npi_edge_types,
                              DAMS_MAX=focal.dams.1000.max,
                              connectivity=total_connect_log)

#Make empty columns for the latent (unobserved) variables

predict_dat$NPI <- NA
predict_dat$NLI <- NA
predict_dat$disturb_index <- NA
predict_dat$thermoreg_index <- NA
predict_dat$damage_risk <- NA

#Ensure all columns are factors, not characters
predict_dat <- predict_dat %>% mutate_all(as.factor)

#Ensure factor levels are in correct order
levels(predict_dat$road_density)<-c("LOW", "MED", "HIGH")
levels(predict_dat$landscape_nutritional_score)<-c("LOW", "MED", "HIGH")
levels(predict_dat$patch_nutritional_score)<-c("LOW", "MED", "HIGH")
levels(predict_dat$urban_prox)<-c("LOW", "MED", "HIGH")
levels(predict_dat$lf_density)<-c("LOW", "MED", "HIGH")
levels(predict_dat$DAMS_MAX)<-c("LOW", "MED", "HIGH")
levels(predict_dat$connectivity)<-c("LOW", "MED", "HIGH")
levels(predict_dat$NPI)<-c("LOW", "MED", "HIGH")
levels(predict_dat$NLI)<-c("LOW", "MED", "HIGH")
levels(predict_dat$disturb_index)<-c("LOW", "MED", "HIGH")
levels(predict_dat$thermoreg_index)<-c("LOW", "MED", "HIGH")
levels(predict_dat$damage_risk)<-c("LOW","LOW-MED","MED","MED-HIGH","HIGH")

#Predict values for latent variables
pred_nli = predict(dfit,node="NLI",data=predict_dat, method = "bayes-lw")
pred_npi = predict(dfit,node="NPI",data=predict_dat, method = "bayes-lw")
pred_thermoreg = predict(dfit,node="thermoreg_index",data=predict_dat, method = "bayes-lw")
pred_disturb = predict(dfit,node="disturb_index",data=predict_dat, method = "bayes-lw")

#fill the columns
predict_dat$NLI <- pred_nli
predict_dat$NPI <- pred_npi
predict_dat$thermoreg_index <- pred_thermoreg
predict_dat$disturb_index <- pred_disturb
  
#Predict damage
pred_damage = predict(dfit,node = "damage_risk",data = predict_dat,method = "bayes-lw")

predict_dat$damage_risk <- pred_damage

#Add damage risk back into spatial dataset

df_categorical$pred_damage <- pred_damage
```

### Draw maps

```{r}
#| echo: true
#| warning: false
#| message: false

all_plotted <- plot(df_categorical["pred_damage"])

# Initialize a list to store individual plots
plots <- list()

library(RColorBrewer)
#Get last 5 colours from YIOrRd color palette
# Define the palette name
palette_name <- "YlOrRd"
#color_palette <- RColorBrewer::brewer.pal(n = 9, name = palette_name)[3:7]
color_palette <- c("yellow","#FED976","#FD8D3C", "#FC4E2A", "#E31A1C")

# Loop through each area in 'buffered_woodland_12km'
for (i in 1:nrow(buffered_woods_12km)) {
  # Extract the i-th area
  area <- buffered_woods_12km[i, ]
  
  # Crop 'df_categorical' to the extent of 'area'
  cropped_data <- st_intersection(df_categorical, area)
  
  # Create a plot for this cropped data
  plot_title <- paste(buffered_woods_12km$Region[i])
  plot <- ggplot() +
    geom_sf(data = cropped_data, aes(fill = pred_damage)) +
    ggtitle(plot_title) +
   scale_fill_manual(values = color_palette, limits = levels(df_categorical$pred_damage)) +  # Color scale from yellow to red
    theme_minimal()  # You can customize the theme and other aspects
  
  # Store the plot in the list
  plots[[i]] <- plot
}

# Display each plot
for (plot in plots) {
  print(plot)  # This will print each plot individually
}


# Save each plot to a separate PDF file with 300 dpi resolution

# Define the target directory
output_dir <- here("output", "figures")

# Define plot parameters
dpi <- 300  # Resolution in dots per inch
width <- 6  # Width of the PDF in inches
height <- 6  # Height of the PDF in inches

for (i in 1:length(plots)) {
  plot <- plots[[i]]  # Get the plot from the list
  filename <- file.path(output_dir, paste0("plot_area_", i, ".pdf"))
  
  # Save the plot as a 300 dpi PDF
  ggsave(
    filename = filename, 
    plot = plot, 
    device = "pdf",  # Specify the file format
    width = width, 
    height = height, 
    dpi = dpi  # Set the resolution
  )
}


```

### Spare code

```{r}
# Generate random points within each woodland
#This would simulate the user picking a point on a map rather than a polygon
#random_points_list <- lapply(woods$geometry, generate_random_point)  
#random_points <- as.data.frame(random_points_list)
#random_points <- random_points %>%
#  pivot_longer(cols = everything(), values_to = "geometry") #tidy object
#random_points <- random_points %>% dplyr::select(-name)
# Create an sf object with the random points and patch IDs
#wood_metadata <- st_drop_geometry(woods)
#wood_points <- cbind(wood_metadata, random_points)
#wood_points <- st_as_sf(wood_points)
```

```{r}
# Cropped rasters list to store results

#Should we set a centre point for each woodland and crop around that to make a square?
#At the moment it is buffering to the edge of the woodland, making an irregular shaped buffer.

buffered_cropped_rasters <- list()

# Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_12km$geometry)) {
  buffer_geometry <- buffered_woods_12km$geometry[i]
  
  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
  # Crop each raster to the current buffer extent
  map_cropped <- crop(map, buffer_extent)
  a_roads_cropped <- crop(a_roads, buffer_extent)
  b_roads_cropped <- crop(b_roads, buffer_extent)
  motorways_cropped <- crop(motorways, buffer_extent)
  nearest_urban_cropped <- crop(nearest_urban, buffer_extent)
  nearest_suburb_cropped <- crop(nearest_suburb, buffer_extent)
  lf_cropped <- crop(lf, buffer_extent)
  dams_cropped <- crop(dams, buffer_extent)
  
  # Store the cropped rasters in the list, using the buffer name as the key
  buffered_cropped_rasters[[paste("buffer", buffered_woods_12km$patch_ID[i])]] <- list(
    map = map_cropped,
    a_roads = a_roads_cropped,
    b_roads = b_roads_cropped,
    motorways = motorways_cropped,
    nearest_urban = nearest_urban_cropped,
    nearest_suburb = nearest_suburb_cropped,
    lf = lf_cropped,
    dams = dams_cropped
  )
}

plot(buffered_cropped_rasters[[1]]$map)
plot(buffered_woods_12km$geometry[1,], add=TRUE, border="blue", lwd=2, col=NA)
```
