---
title: "iDeer Tool Function"
format: html
editor: visual

#User selects point on map
#10km landscape around it
#Cut all raster to 10km, buffer by 2km for spatial processes (12km total)
#Extract the metrics (roads, urban, connectivity etc)
#Run the BBN
#Loop this process for the 5 different landscapes
#Do this for the 5 DIA regions
#Write as a function
---

```{r}
library(plyr)
library(ggplot2)
library(dplyr)
library(sf)
library(raster)
library(terra)
library(here)
library(progress)
here()
```

### Load data

```{r}
bng <- 27700

#Load NFI/LCM2015 master map
map <- raster(here("data/derived-data/2015-spatial-layers/edge_core_raster_all_tiles_EW.tif"))
crs(map) <- bng

#Load roads at 1km scale
a_roads <- raster(here("data/derived-data/2015-spatial-layers/Ard_1000_mosaic_all_tiles_corrected.tif"))
crs(a_roads) <- bng
b_roads <- raster(here("data/derived-data/2015-spatial-layers/Brd_1000_mosaic_all_tiles_corrected.tif"))
crs(b_roads) <- bng
motorways <- raster(here("data/derived-data/2015-spatial-layers/motorway_1000_mosaic_all_tiles_corrected.tif"))
crs(motorways) <- bng

#Load nearest urban
nearest_urban <- raster(here("data/derived-data/2015-spatial-layers/nearest_urban_raster_2015_all_tiles_corrected.tif"))
crs(nearest_urban) <- bng
nearest_suburb <- raster(here("data/derived-data/2015-spatial-layers/nearest_suburban_raster_2015_all_tiles_corrected.tif"))
crs(nearest_suburb) <- bng

#Load linear features at 1km scale
lf <- raster(here("data/derived-data/2015-spatial-layers/linear_features/lf_1000_mosaic_all_tiles_corrected.tif"))
crs(lf) <- bng

#Load MAX DAMS at 1km scale
dams <- raster(here("data/derived-data/2015-spatial-layers/Max_DAMS/focal.dams.1000.max.tif"))
crs(dams) <- bng

#Read in woodland polygon dataset
#Combined NFI 2015 and LCM 2015 woodlands

hab_patches_all <- st_read(here("data/derived-data/nfi_lcm_overlaid_shapefile_woods_export.shp"))
st_crs(hab_patches_all) <- bng

head(hab_patches_all)

hab_patches_all <- hab_patches_all %>%
  rename(patch_ID = Id) %>%
  rename(patch_area = Shape_Area)
```

### Choose woodlands

```{r}

#Choose five woodlands within the DIA priority areas to extract data for
#This is simulating the user selecting a woodland of interest

#Suffolk
suffolk_wood <- hab_patches_all %>% subset(patch_ID == 260977)
#Cumbria
cumbria_wood <- hab_patches_all %>% subset(patch_ID == 142765)
#South Wales 
sw_wood <- hab_patches_all %>% subset(patch_ID == 340972)
#North Wales
nw_wood <- hab_patches_all %>% subset(patch_ID == 188539)
#Oxford 
ox_wood <- hab_patches_all %>% subset(patch_ID == 370647)
#select all buffers together
# Select the desired buffers
woods <- hab_patches_all %>%
  filter(patch_ID %in% c("260977", "142765", "340972", "188539", "370647"))
woods <- st_as_sf(woods)


```

### Buffer woodlands for spatial data extraction

```{r}
#Buffer all woods by 10km
buffered_woods_10km <- st_as_sf(st_buffer(woods, dist = 10000))  # dist = 10km in meters
#Buffer all woods by 12km (10km landscape + 2km edge buffer)
buffered_woods_12km <- st_as_sf(st_buffer(woods, dist = 12000))  # dist = 10km in meters

#ensure crs of buffers is BNG
st_crs(buffered_woods_12km) <- bng
st_crs(buffered_woods_10km) <- bng

#Check points, woodlands buffers line up
plot(st_geometry(buffered_woods_12km), col = NA)
plot(st_geometry(woods), col = "black", add = TRUE)
#plot(st_geometry(wood_points), add = TRUE, col = "red")

```

### Extract raster variables for all woodlands in 10km buffer

```{r}
#NEAREST URBAN EXTRACTION ####

extract_urban_rasters <- function(buffered_woods_10km, raster_layer, woodlands) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_10km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_10km$geometry)) {
  buffer_geometry <- buffered_woods_10km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
  # Crop raster to the current buffer extent
  nearest_urban_cropped <- crop(nearest_urban, buffer_extent)
  
  #Convert to spatraster
  nearest_urban_spatraster <- terra::rast(nearest_urban_cropped)
  
  #Filter for woodland patches that intersect buffer
  
  woods_buffer <- st_filter(woodlands, buffered_woods_10km[i, ], .predicate = st_intersects)
  
    # Extract data for all woodland polygons within the current buffer
    extraction <- terra::extract(nearest_urban_spatraster, vect(woods_buffer), fun = mean, na.rm = TRUE)
    
    #cbind to the original woods_buffer columns
    
    extraction_cbind <- cbind(woods_buffer, extraction)
    
    # Store the extracted data in the results list
    extraction_results[[paste("Buffer", i)]] <- extraction_cbind
  }
  
  # Combine all extracted data into a single data frame
  nearest_urban_extraction_results <- bind_rows(extraction_results)  # Combines all results into one data frame
  
  return(nearest_urban_extraction_results) # Return the combined data frame
}

nearest_urban_extraction <- extract_urban_rasters(buffered_woods_10km, nearest_urban, hab_patches_all)

ggplot(nearest_urban_extraction) +
  geom_sf(aes(fill = nearest_urban_raster_2015_all_tiles_corrected))

write.csv(nearest_urban_extraction, here("data/derived-data/Tool-extract-function-test/nearest_urban_extraction.csv"))
```

```{r}
#NEAREST SUBURBAN EXTRACTION ####
extract_suburban_rasters <- function(buffered_woods_10km, raster_layer, woodlands) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_10km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_10km$geometry)) {
  buffer_geometry <- buffered_woods_10km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
  # Crop raster to the current buffer extent
  nearest_suburban_cropped <- crop(nearest_suburb, buffer_extent)
  
  #Convert to spatraster
  nearest_suburban_spatraster <- terra::rast(nearest_suburban_cropped)
  
  #Filter for woodland patches that intersect buffer
  
  woods_buffer <- st_filter(woodlands, buffered_woods_10km[i, ], .predicate = st_intersects)
  
    # Extract data for all woodland polygons within the current buffer
    extraction <- terra::extract(nearest_suburban_spatraster, vect(woods_buffer), fun = mean, na.rm = TRUE)
    
    #cbind to the original woods_buffer columns
    
    extraction_cbind <- cbind(woods_buffer, extraction)
    
    # Store the extracted data in the results list
    extraction_results[[paste("Buffer", i)]] <- extraction_cbind
  }
  
  # Combine all extracted data into a single data frame
  nearest_suburban_extraction_results <- bind_rows(extraction_results)  # Combines all results into one data frame
  
  return(nearest_suburban_extraction_results) # Return the combined data frame
}

nearest_suburban_extraction <- extract_suburban_rasters(buffered_woods_10km, nearest_suburban, hab_patches_all)

ggplot(nearest_suburban_extraction) +
  geom_sf(aes(fill = nearest_suburban_raster_2015_all_tiles_corrected))

write.csv(nearest_suburban_extraction, here("data/derived-data/Tool-extract-function-test/nearest_suburban_extraction.csv"))
```

```{r}
#MAX DAMS EXTRACTION ####
extract_dams_raster <- function(buffered_woods_10km, raster_layer, woodlands) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_10km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_10km$geometry)) {
  buffer_geometry <- buffered_woods_10km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
  # Crop raster to the current buffer extent
  dams_cropped <- crop(dams, buffer_extent)
  
  #Convert to spatraster
  dams_spatraster <- terra::rast(dams_cropped)
  
  #Filter for woodland patches that intersect buffer
  
  woods_buffer <- st_filter(woodlands, buffered_woods_10km[i, ], .predicate = st_intersects)
  
    # Extract data for all woodland polygons within the current buffer
    extraction <- terra::extract(dams_spatraster, vect(woods_buffer), fun = mean, na.rm = TRUE)
    
    #cbind to the original woods_buffer columns
    
    extraction_cbind <- cbind(woods_buffer, extraction)
    
    # Store the extracted data in the results list
    extraction_results[[paste("Buffer", i)]] <- extraction_cbind
  }
  
  # Combine all extracted data into a single data frame
 dams_extraction_results <- bind_rows(extraction_results)  # Combines all results into one data frame
  
  return(dams_extraction_results) # Return the combined data frame
}

dams_extraction <- extract_dams_raster(buffered_woods_10km, dams, hab_patches_all)

ggplot(dams_extraction) +
  geom_sf(aes(fill = focal.dams.1000.max))

write.csv(dams_extraction, here("data/derived-data/Tool-extract-function-test/DAMS_MAX_extraction_1000.csv"))
```

```{r}
#LINEAR FEATURES EXTRACTION ####
extract_lf_raster <- function(buffered_woods_10km, raster_layer, woodlands) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_10km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_10km$geometry)) {
  buffer_geometry <- buffered_woods_10km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
  # Crop raster to the current buffer extent
  lf_cropped <- crop(lf, buffer_extent)
  
  #Convert to spatraster
  lf_spatraster <- terra::rast(lf_cropped)
  
  #Filter for woodland patches that intersect buffer
  
  woods_buffer <- st_filter(woodlands, buffered_woods_10km[i, ], .predicate = st_intersects)
  
    # Extract data for all woodland polygons within the current buffer
    extraction <- terra::extract(lf_spatraster, vect(woods_buffer), fun = mean, na.rm = TRUE)
    
    #cbind to the original woods_buffer columns
    
    extraction_cbind <- cbind(woods_buffer, extraction)
    
    # Store the extracted data in the results list
    extraction_results[[paste("Buffer", i)]] <- extraction_cbind
  }
  
  # Combine all extracted data into a single data frame
lf_extraction_results <- bind_rows(extraction_results)  # Combines all results into one data frame
  
  return(lf_extraction_results) # Return the combined data frame
}

lf_extraction <- extract_lf_raster(buffered_woods_10km, lf, hab_patches_all)

ggplot(lf_extraction) +
  geom_sf(aes(fill = lf_1000_mosaic_all_tiles_corrected))

write.csv(lf_extraction, here("data/derived-data/Tool-extract-function-test/linear_feature_extraction_1000.csv"))
```

```{r}
#A ROAD DENSITY EXTRACTION ####
extract_ard_raster <- function(buffered_woods_10km, raster_layer, woodlands) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_10km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_10km$geometry)) {
  buffer_geometry <- buffered_woods_10km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
  # Crop raster to the current buffer extent
  ard_cropped <- crop(a_roads, buffer_extent)
  
  #Convert to spatraster
  ard_spatraster <- terra::rast(ard_cropped)
  
  #Filter for woodland patches that intersect buffer
  
  woods_buffer <- st_filter(woodlands, buffered_woods_10km[i, ], .predicate = st_intersects)
  
    # Extract data for all woodland polygons within the current buffer
    extraction <- terra::extract(ard_spatraster, vect(woods_buffer), fun = mean, na.rm = TRUE)
    
    #cbind to the original woods_buffer columns
    
    extraction_cbind <- cbind(woods_buffer, extraction)
    
    # Store the extracted data in the results list
    extraction_results[[paste("Buffer", i)]] <- extraction_cbind
  }
  
  # Combine all extracted data into a single data frame
ard_extraction_results <- bind_rows(extraction_results)  # Combines all results into one data frame
  
  return(ard_extraction_results) # Return the combined data frame
}

ard_extraction <- extract_ard_raster(buffered_woods_10km, a_roads, hab_patches_all)

ggplot(ard_extraction) +
  geom_sf(aes(fill = Ard_1000_mosaic_all_tiles_corrected))

write.csv(ard_extraction, here("data/derived-data/Tool-extract-function-test/Aroads_extraction_1000.csv"))
```

```{r}
#B ROAD DENSITY EXTRACTION ####
extract_brd_raster <- function(buffered_woods_10km, raster_layer, woodlands) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_10km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_10km$geometry)) {
  buffer_geometry <- buffered_woods_10km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
  # Crop raster to the current buffer extent
  brd_cropped <- crop(b_roads, buffer_extent)
  
  #Convert to spatraster
  brd_spatraster <- terra::rast(brd_cropped)
  
  #Filter for woodland patches that intersect buffer
  
  woods_buffer <- st_filter(woodlands, buffered_woods_10km[i, ], .predicate = st_intersects)
  
    # Extract data for all woodland polygons within the current buffer
    extraction <- terra::extract(brd_spatraster, vect(woods_buffer), fun = mean, na.rm = TRUE)
    
    #cbind to the original woods_buffer columns
    
    extraction_cbind <- cbind(woods_buffer, extraction)
    
    # Store the extracted data in the results list
    extraction_results[[paste("Buffer", i)]] <- extraction_cbind
  }
  
  # Combine all extracted data into a single data frame
brd_extraction_results <- bind_rows(extraction_results)  # Combines all results into one data frame
  
  return(brd_extraction_results) # Return the combined data frame
}

brd_extraction <- extract_brd_raster(buffered_woods_10km, b_roads, hab_patches_all)

ggplot(brd_extraction) +
  geom_sf(aes(fill = Brd_1000_mosaic_all_tiles_corrected))

write.csv(brd_extraction, here("data/derived-data/Tool-extract-function-test/Broads_extraction_1000.csv"))
```

```{r}
#MOTORWAY DENSITY EXTRACTION ####
extract_motorways_raster <- function(buffered_woods_10km, raster_layer, woodlands) {
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(buffered_woods_10km),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)
  
  # List to store extracted data
  extraction_results <- list()
  
  # Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_10km$geometry)) {
  buffer_geometry <- buffered_woods_10km$geometry[i]
      pb$tick()  # Update the progress bar

  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
  # Crop raster to the current buffer extent
  motorways_cropped <- crop(motorways, buffer_extent)
  
  #Convert to spatraster
  motorways_spatraster <- terra::rast(motorways_cropped)
  
  #Filter for woodland patches that intersect buffer
  
  woods_buffer <- st_filter(woodlands, buffered_woods_10km[i, ], .predicate = st_intersects)
  
    # Extract data for all woodland polygons within the current buffer
    extraction <- terra::extract(motorways_spatraster, vect(woods_buffer), fun = mean, na.rm = TRUE)
    
    #cbind to the original woods_buffer columns
    
    extraction_cbind <- cbind(woods_buffer, extraction)
    
    # Store the extracted data in the results list
    extraction_results[[paste("Buffer", i)]] <- extraction_cbind
  }
  
  # Combine all extracted data into a single data frame
motorways_extraction_results <- bind_rows(extraction_results)  # Combines all results into one data frame
  
  return(motorways_extraction_results) # Return the combined data frame
}

motorways_extraction <- extract_motorways_raster(buffered_woods_10km, motorways, hab_patches_all)

ggplot(motorways_extraction) +
  geom_sf(aes(fill = motorway_1000_mosaic_all_tiles_corrected))


write.csv(motorways_extraction, here("data/derived-data/Tool-extract-function-test/motorways_extraction_1000.csv"))
```

```{r}
#LAND COVER EXTRACTION FOR 1km AROUND WOODLANDS ####
#Function written by Becks

  polyFragStats <- function(POLY, veg) {
  # Create a buffer around the polygon
  buffer_poly <- buffer(POLY, width = 1000, dissolve = TRUE)
  
  # Cut out the original polygon from the buffer
  buffer_area <- erase(buffer_poly, POLY)
  
  # Crop land cover map by the buffer area
  fcrop <- crop(veg, buffer_area)
  
  # Mask the cropped land cover map
  fmask <- mask(fcrop, buffer_area)
  
  # Extract class statistics
  res <- raster::extract(fmask, buffer_area, method = "simple")
  
  # Convert the results to a data frame and calculate statistics
  res <- as.data.frame(table(res))
  res$patch_ID <- POLY$patch_ID
  
  res <- res %>%
    mutate(total_pix = as.numeric(sum(Freq)),
           class = res) %>%
    dplyr::select(patch_ID, class, Freq, total_pix) %>%
    mutate(perc_cov = Freq * 100 / total_pix, buff_area = area(buffer_area))
  
  return(res)
}

#Filter for woodland patches that intersect buffers
#Try with 1 buffer to check it's working OK
woods_buffers <- st_filter(hab_patches_all, buffered_woods_10km, .predicate = st_intersects)

POLY=as(woods_buffers, "Spatial")
veg = map

# Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = length(POLY),
  clear = FALSE,
  width = 60
)

respoly <- ldply(1:length(POLY), function(i) {  # for each polygon
  pb$tick()  # Update the progress bar
  out <- polyFragStats(POLY[i,],  veg=veg) # compute fragstats metrics 
  out$buffer_size <- "b1000"
  return(out)
})  

respoly

write.csv(respoly, here("data/derived-data/Tool-extract-function-test/landscapemetrics_1000.csv"))

```

```{r}
#WOODLAND PIXEL EXTRACTION ####
#This will tell us the type of woodland each polygon is, and the amount of different edges inside each one.
#This will be our patch quality indicator.

extract_pixel_frequencies <- function(woodland_polygons, land_cover_raster, polygon_id_col = "id") {

  # Initialize an empty list to store results
  results_list <- list()
  
   # Initialize a progress bar
  pb <- progress_bar$new(
    format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
    total = nrow(woodland_polygons),
    clear = FALSE,
    width = 60
  )

  # Loop over each woodland polygon to extract raster values
  for (i in 1:nrow(woodland_polygons)) {
    # Update the progress bar
    pb$tick()  # This indicates the progress as you loop through the polygons
    
    # Get the current polygon
    polygon <- woodland_polygons[i, ]

    # Convert to 'Spatial' for compatibility with 'raster' operations
    polygon_sp <- as(polygon, "Spatial")

    # Crop the raster to the polygon's extent and mask to the polygon
    raster_cropped <- crop(land_cover_raster, polygon_sp)
    raster_masked <- mask(raster_cropped, polygon_sp)

    # Extract raster values within the polygon
    pixel_values <- extract(raster_masked, polygon_sp)

    # Flatten the list of extracted values (in case there are multiple parts)
    pixel_values <- unlist(pixel_values)

    # Count the frequency of each unique raster pixel value
    value_counts <- as.data.frame(table(pixel_values))
    value_counts$pixel_values <- as.character(value_counts$pixel_values)
    value_counts$pixel_values <- as.numeric(value_counts$pixel_values)

    # Add additional information (like polygon ID)
    value_counts <- value_counts %>%
      mutate(polygon_id = polygon[[polygon_id_col]],  # Use specified column for ID
             percentage = Freq / sum(Freq) * 100)  # Calculate percentage coverage

    # Store the results
    results_list[[i]] <- value_counts
  }

  # Combine all results into a single data frame
  final_results <- bind_rows(results_list)

  return(final_results)
}

#Filter for woodland patches that intersect buffers
woods_buffers <- st_filter(hab_patches_all, buffered_woods_10km, .predicate = st_intersects)

pixel_frequencies <- extract_pixel_frequencies(woods_buffers, map, polygon_id_col = "patch_ID")
#Round all values to 3 decimal place
#pixel_freq <- pixel_frequencies
#pixel_freq$extracted_values_flat<-as.numeric(pixel_freq$extracted_values_flat)
#pixel_freq <- pixel_freq%>%mutate_if(is.numeric, ~ round(., 3))

write.csv(pixel_frequencies, here("data/derived-data/Tool-extract-function-test/woodland_patch_pixel_types.csv"))
```

### Calculate connectivity for every woodland inside the 10km buffers

```{r}
#### SET CONNECTIVITY PARAMETERS ####
# connectivity parametetrs - % of dispersers reaching a set distance #
percentage_dispersers <- 0.05 ### 5%
dispersal_distance <- 434 #Gets curve up to 1000m on x axis
dispersal_contribution <-
  -((log(1 / percentage_dispersers)) / dispersal_distance)

# Buffer distance represents the cut off - this stops the script measuring every pairwise combination
dispersal_cutoff <- 0.999 ### 99.9% cut off
#buffer_cutoff <-
  #round(log(1 / (1 - dispersal_cutoff)) / (log(1 / percentage_dispersers) /
                                             #dispersal_distance), digits = 2)
buffer_cutoff = 1000 #1km

# test & plot of dispersal contribution for sequence of distance up to buffer_cutoff
test <-
  data.frame(distance = (seq(0, buffer_cutoff, by = 0.5))) %>%
  mutate(contribution = (exp(dispersal_contribution * distance) * 100))

ggplot(data = test, aes(x = distance, y = contribution, group = 1)) +
  geom_line() +
  geom_point() +
  geom_vline(xintercept = buffer_cutoff, color = "red")

#See how buffer size looks
focal_patch_buffer <- st_buffer(hab_patches[1,], buffer_cutoff)
plot(st_geometry(focal_patch_buffer))
plot(st_geometry(hab_patches[1,]),add=TRUE)

```

```{r}

connectivity_results <- list()

  # Use the 12km buffer as the source patches area
  source_woods <- st_filter(hab_patches_all, buffered_woods_12km, .predicate = st_intersects)

  # Crop the woodlands in the original 10km square to get the focal patches
  focal_woods <- st_filter(hab_patches_all, buffered_woods_10km, .predicate = st_intersects)
  n_distinct(focal_woods$patch_ID)
  #[1] 7884

  # Calculate connectivity for each focal woodland
  Connectivity_table <- NULL
  
  # Initialize progress bar
pb <- progress_bar$new(
  format = "[:bar] :current/:total (:percent) in :elapsed, eta: :eta",
  total = nrow(focal_woods),
  clear = FALSE,   # Keep the progress bar on the screen
  width = 60
)

  # Loop through each focal woodland to calculate connectivity
  for (j in 1:nrow(focal_woods)) {
    
    # Focal patch in the 10km square
    focal_patch <- focal_woods[j, ]
    
    #update progress bar
    pb$tick
    
    # Buffer around the focal patch to include source patches
    focal_patch_buffer <- st_buffer(focal_patch, buffer_cutoff)
    
    # Find source patches within this buffer
    source_patches <- st_filter(source_woods, focal_patch_buffer, .predicate = st_intersects)
    
    # Calculate distances between the focal and source patches
    patch_dist <- as.vector(st_distance(focal_patch, source_patches))
    
    # Create a data frame for connectivity information
    Conn_table_site <- data.frame(
      focal_patch = focal_patch$patch_ID,
      focal_patch_area = as.numeric(focal_patch$patch_area),
      source_patch = source_patches$patch_ID,
      source_patch_area = as.numeric(source_patches$patch_area),
      distance = patch_dist,
      incoming_connect = ifelse(
        focal_patch$patch_ID == source_patches$patch_ID,
        NA,
        source_patches$patch_area * exp(dispersal_contribution * patch_dist)
      )
    )
    
    # Append to 'Connectivity_table'
    if (is.null(Connectivity_table)) {
      Connectivity_table <- Conn_table_site
    } else {
      Connectivity_table <- rbind(Connectivity_table, Conn_table_site)
    }
  
  n_distinct(Connectivity_table$focal_patch)

  # Store results for this woodland
  connectivity_results[[paste("focal_wood", focal_woods$patch_ID[i])]] <- Connectivity_table
}
```

```{r}

#remove rows where focal_patch = source_patch
Connectivity_table_filt <- Connectivity_table %>%
  filter(focal_patch != source_patch)

# Identify missing unique patch IDs
# These patches have no source patches so will have been filtered out
# We need to put them back in so we can give them an incoming_connect value of 0
missing_patch_ids <- setdiff(
     unique(Connectivity_table$focal_patch),
     unique(Connectivity_table_filt$focal_patch)
 )
 
 print(missing_patch_ids)  # This will show which patch IDs are missing
#[1] 132853 133946 141304 260769 358349 359726
 
missing_patches <- Connectivity_table %>% filter(focal_patch %in% missing_patch_ids) %>%
  # Replace NAs in 'incoming_connect' with 0
  mutate(incoming_connect = coalesce(incoming_connect, 0))  # Replace NA with 0

#Make rows for patches with no connectivity so we don't lose any patches
Connectivity_table_filt <- rbind(Connectivity_table_filt, missing_patches)

#Ensure no duplicate rows
Connectivity_table_filt <- Connectivity_table_filt %>% distinct()

#Now no patches should be missing
n_distinct(Connectivity_table_filt$focal_patch)

#sum incoming connectivity by focal patch
incoming_connectivity_sum <- Connectivity_table_filt %>%
  dplyr::select(focal_patch, incoming_connect) %>%
  group_by(focal_patch) %>%
  dplyr::summarise(total_connect = sum(incoming_connect),
            n = n()) %>%
  mutate(n = if_else(total_connect == 0, 0, n)) %>%  # Set 'n' to 0 when 'incoming_connect' is 0
  rename(patch_ID = focal_patch)

#save incoming connectivity table

write.csv(incoming_connectivity_sum, here("data/derived-data/Tool-extract-function-test/incoming_connectivity_sum.csv"))
```

### Ready the spatial data for the Bayesian Belief Network

#### Calculate Nutritional Landscape Composition (at 1km scale)

```{r}
#CALCULATE NUTRITIONAL LANDSCAPE QUALITY ####

lcm <- respoly
#Filter out zeros, water, rock and sediment
#lcm <- lcm %>% filter(!class %in% c(0,12,13,14,15,16,17,18))
#Filter out edge classes with .7 or .8 as these are saltwater and freshwater edges - not useful for deer
#lcm <- lcm %>% filter(!grepl(".7",class)) %>%
                                       #filter(!grepl(".8",class)) 

#Load landscape nutritional scores for raster categories
#Edge forage quality scores sum of woodland score + adjacent land cover score
#Water scores added in to prevent loss of woodland edge pixels
#Water edge pixels receive one point higher than regular woodland
#Still an edge so will be highly nutritious
edge.val.types <- read.csv(file.path(here("data/raw-data/Expert_derived_quality_ranks_edge_types_with_water.csv")))
edge.val.types$class <- as.factor(edge.val.types$class)

#Stitch the quality scores into the lcm dataset

lc.edge.types <- lcm %>% left_join(edge.val.types, by = "class") %>%
                                          filter(!is.na(Land.cover))

#Calculate Nutritional Landscape Index (NLI) for separate edges and lc ####
#Weight edges by their type: https://www.sciencedirect.com/science/article/pii/S0304380009003871 #fig1  

# Calculate the weighted score for each pixel type
lc.edge.types <- lc.edge.types %>%
  mutate(weighted_score = perc_cov * Forage_Q)

# Sum the weighted scores for each SQUID section
weighted_sum_by_section_buffer <- lc.edge.types %>%
  group_by(patch_ID, buffer_size) %>%
  summarise(total_weighted_score = sum(weighted_score),
            total_perc_cov = sum(perc_cov))

# Calculate the weighted average score for each woodland polygon
weighted_average_score_buffer <- weighted_sum_by_section_buffer %>%
  mutate(weighted_avg_score = total_weighted_score / total_perc_cov)

weighted_average_score_buffer <- weighted_average_score_buffer %>%
  group_by(patch_ID) %>%
  summarise(
    nli_edge_types_1000 = sum(weighted_avg_score[buffer_size == "b1000"], na.rm = TRUE)
  )

# Join the calculated values back to the original dataset
lc.edge.types.NLI <- left_join(woods_buffers, weighted_average_score_buffer, by = c("patch_ID"))

```

#### Calculate Nutritional Patch Composition

```{r}
#CALCULATE NUTRITIONAL PATCH QUALITY ####

pcm <- pixel_frequencies
pcm<-pcm%>%rename(class=pixel_values,
                  patch_ID=polygon_id,
                  perc_cov=percentage)
#Filter out zeros, water, rock and sediment
#pcm <- pcm %>% filter(!class %in% c(0,12,13,14,15,16,17,18))
#Filter out edge classes with .7 or .8 as these are saltwater and freshwater edges #- not useful for deer
#pcm <- pcm %>% filter(!grepl(".7",class)) %>%
                                       #filter(!grepl(".8",class)) 

#Load landscape nutritional scores for raster categories
#Edge forage quality scores sum of woodland score + adjacent land cover score
#Water scores added in to prevent loss of woodland edge pixels
#Water edge pixels receive one point higher than regular woodland
#Still an edge so will be highly nutritious
edge.val.types <- read.csv(file.path(here("data/raw-data/Expert_derived_quality_ranks_edge_types_with_water.csv")))
edge.val.types <- edge.val.types%>%dplyr::select(-"X")

#Check edge.val.types$class and pcm$class match in format
class(edge.val.types$class)
#[1] "numeric"
edge.val.types <- edge.val.types%>%arrange(class)
unique(edge.val.types$class)
class(pcm$class)
#[1] "numeric"
pcm <-pcm%>%arrange(class)
unique(pcm$class)

# Round to 3 decimal places
pcm$class <- round(pcm$class, 3)
edge.val.types$class <- round(edge.val.types$class, 3)

#Stitch the quality scores into the lcm dataset

pc.edge.types <- pcm %>% left_join(edge.val.types, by = c("class")) #%>%
                                          #filter(!is.na(Land.cover))

#Calculate Nutritional Landscape Index (NLI) for separate edges and lc ####
#Weight edges by their type: https://www.sciencedirect.com/science/article/pii/S0304380009003871 #fig1  

# Calculate the weighted score for each pixel type
pc.edge.types <- pc.edge.types %>%
  mutate(weighted_score = perc_cov * Forage_Q)

# Sum the weighted scores for each woodland polygon
weighted_sum_by_section <- pc.edge.types %>%
  group_by(patch_ID) %>%
  summarise(total_weighted_score = sum(weighted_score),
            total_perc_cov = sum(perc_cov))

# Calculate the weighted average score for each woodland polygon
weighted_average_score <- weighted_sum_by_section %>%
  mutate(weighted_avg_score = total_weighted_score / total_perc_cov)

weighted_average_score <- weighted_average_score %>%
  group_by(patch_ID) %>%
  summarise(
    npi_edge_types = sum(weighted_avg_score, na.rm = TRUE))

# Join the calculated values back to the original dataset
lc.edge.types.NPI <- left_join(woods_buffers, weighted_average_score, by = c("patch_ID"))
```

#### left_join the data sets together

```{r}

#drop geometries from datasets
ard_extraction <- st_drop_geometry(ard_extraction)
brd_extraction <-st_drop_geometry(brd_extraction)
motorways_extraction<-st_drop_geometry(motorways_extraction)
lf_extraction<-st_drop_geometry(lf_extraction)
dams_extraction<-st_drop_geometry(dams_extraction)
nearest_suburban_extraction<-st_drop_geometry(nearest_suburban_extraction)
nearest_urban_extraction<-st_drop_geometry(nearest_urban_extraction)
lc.edge.types.NPI<-st_drop_geometry(lc.edge.types.NPI)
lc.edge.types.NLI<-st_drop_geometry(lc.edge.types.NLI)

#Roads ####
df <- left_join(ard_extraction, brd_extraction, by = c("patch_ID","gridcode","Shape_Leng","patch_area","ID"))
df <- left_join(df, motorways_extraction, by=c("patch_ID","gridcode","Shape_Leng","patch_area","ID"))
#Make new column for all main roads
df$all_roads_1000 <- rowSums(df[,c("Ard_1000_mosaic_all_tiles_corrected", "Brd_1000_mosaic_all_tiles_corrected", "motorway_1000_mosaic_all_tiles_corrected")])
#Drop individual road columns
df <- df%>%dplyr::select(-c("Ard_1000_mosaic_all_tiles_corrected", "Brd_1000_mosaic_all_tiles_corrected", "motorway_1000_mosaic_all_tiles_corrected"))

#Linear features####
df <- left_join(df, lf_extraction, by=c("patch_ID","gridcode","Shape_Leng","patch_area","ID"))

#MAX DAMS ####
df <- left_join(df, dams_extraction, by=c("patch_ID","gridcode","Shape_Leng","patch_area","ID"))

#Nearest urban and suburban ####
df <- left_join(df, nearest_suburban_extraction, by=c("patch_ID","gridcode","Shape_Leng","patch_area","ID"))
df <- left_join(df, nearest_urban_extraction, by=c("patch_ID","gridcode","Shape_Leng","patch_area","ID"))
#Make column for nearest urban and suburban using minimum value
df$urban_proximity <- pmin(df$nearest_suburban_raster_2015_all_tiles_corrected, df$nearest_urban_raster_2015_all_tiles_corrected)
#Drop individual columns
df <- df%>%dplyr::select(-c("nearest_suburban_raster_2015_all_tiles_corrected","nearest_urban_raster_2015_all_tiles_corrected"))

#Nutritional Patch Index ####
df <- left_join(df, lc.edge.types.NPI, by = c("patch_ID","gridcode","Shape_Leng","patch_area"))

#Nutritional Landscape Index ####
df <- left_join(df, lc.edge.types.NLI, by = c("patch_ID","gridcode","Shape_Leng","patch_area"))

#Connectivity ####

df <- left_join(df, incoming_connectivity_sum, by = c("patch_ID"))

#Add the geometry back in
woodland_geometry <- woods_buffers%>%dplyr::select("patch_ID","geometry")

df<-left_join(df,woodland_geometry, by=c("patch_ID"))
df<-df%>%dplyr::select(-c("Shape_Leng","gridcode","ID"))
df<-df%>%rename(lf_1000 = lf_1000_mosaic_all_tiles_corrected)
df<- st_as_sf(df)

saveRDS(df, here("data/derived-data/Tool-extract-function-test/bbn_test_df_5_patches.rds"))
```

#### Convert numerical values to categorical (LOW, MED, HIGH)

```{r}

df <- readRDS(here("data/derived-data/Tool-extract-function-test/bbn_test_df_5_patches.rds"))

#NEED TO TWEAK THESE CATEGORIES
#DO WE DO THIS BASED ON THE LOCALISED DATA FOR EACH WOODLAND?
#OR USING THE RANGE OF THE WHOLE DATASETS?
#USING LOCALISED DATA WOULD KEEP IT "RELATIVE" TO THE LOCAL LANDSCAPE??

#Replace values in numerical columns with LOW, MED, HIGH

#Linear features
quantile(df$lf_1000, probs = c(0.25, 0.5,0.75))
hist(df$lf_1000)
# Replace numeric values with categorical labels
df$lf_1000 <- ifelse(df$lf_1000<= 2000, "LOW",
            ifelse(df$lf_1000> 2000 & df$lf_250 <= 4000, "MED",
            ifelse(df$lf_1000> 4000, "HIGH", NA)))

#Roads
quantile(df$all_roads_1000, probs = c(0.25, 0.5, 0.75))
hist(df$all_roads_1000)
df$all_roads_1000 <- ifelse(df$all_roads_1000 == 0, "LOW",
            ifelse(df$all_roads_1000 >0 & df$all_roads_1000 <= 2000, "MED",
            ifelse(df$all_roads_1000> 2000, "HIGH", NA)))

#nearest urban/suburban
#Proximity, therefore "HIGH" means closer
hist(df$urban_proximity)
df$urban_proximity <- ifelse(df$urban_proximity <= 250, "HIGH",
                          ifelse(df$urban_proximity >250 & df$urban_proximity <= 500, "MED",
                           ifelse(df$urban_proximity >500,"LOW",NA)))  

#MAX DAMS
quantile(df$focal.dams.250, probs=c(0.25,0.5,0.75))
hist(df$focal.dams.250)
df$focal.dams.250 <- ifelse(df$focal.dams.250 <=10,"LOW",
                      ifelse(df$focal.dams.250 >10 & df$focal.dams.250 <= 15, "MED",
                        ifelse(df$focal.dams.250 >15, "HIGH",NA)))

#nutritional landscape composition score
quantile(df$nli_edge_types_250)
hist(df$nli_edge_types_250)
df$nli_edge_types_250 <- ifelse(df$nli_edge_types_250 <= 0, "LOW",
                          ifelse(df$nli_edge_types_250 >0 & df$nli_edge_types_250 <=5, "MED",
                           ifelse(df$nli_edge_types_250 > 5, "HIGH",NA)))

#total landscape edges

quantile(df$total_edge_pix_b250)
hist(df$total_edge_pix_b250)
df$total_edge_pix_b250 <- ifelse(df$total_edge_pix_b250 <= 500, "LOW",
                            ifelse(df$total_edge_pix_b250 > 372 & df$total_edge_pix_b250 <= 750, "MED",
                              ifelse(df$total_edge_pix_b250 > 750, "HIGH",NA)))

#Export

write.csv(df, file = file.path(here("data/derived-data/nfi_2015_extracted_data_categorical.csv")))

```

### Create the Bayesian Belief Network

#### Initialize Conditional Probability Tables

```{r}
#CPTS for measured nodes ####
urban_prox_cpt<-matrix(c(0.333333,0.333333,0.333333), ncol=3, dimnames = list(NULL, c("LOW", "MED","HIGH")))

road_density_cpt<-matrix(c(0.333333,0.333333,0.333333), ncol=3, dimnames = list(NULL, c("LOW", "MED","HIGH")))

connectivity_cpt<-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

landscape_nutritional_score_cpt<-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

patch_nutritional_score_cpt <- matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

dams_cpt <- landscape_nutritional_score_cpt<-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))

lf_density_cpt <-matrix(c(0.333333,0.333333,0.333333),ncol = 3,dimnames = list(NULL, c("LOW", "MED","HIGH")))
```

```{r}
#CPT for Disturbance Index ####

disturb_index_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3, 3, 3),  # Shape of the array (3x3x3)
  dimnames = list(
    disturb_index = c("LOW", "MED", "HIGH"),
    road_density = c("LOW", "MED", "HIGH"),
    urban_prox = c("LOW","MED","HIGH"),
    connectivity = c("LOW", "MED", "HIGH")
  )
)
#order = disturb_index, road_density, urban_prox,connectivity
#table1: urban_prox = LOW, connectivity = LOW
disturb_index_cpt["LOW","LOW","LOW","LOW"] <- 1
disturb_index_cpt["MED","MED","LOW","LOW"] <- 1
disturb_index_cpt["HIGH","HIGH","LOW","LOW"] <- 1
#table2:  urban_prox = MED, connectivity = LOW
disturb_index_cpt["MED","LOW","MED","LOW"] <- 1
disturb_index_cpt["MED","MED","MED","LOW"] <- 1
disturb_index_cpt["HIGH","HIGH","MED","LOW"] <- 1
#table3: urban_prox = HIGH, connectivity = LOW
disturb_index_cpt["HIGH","LOW","HIGH","LOW"] <- 1
disturb_index_cpt["HIGH","MED","HIGH","LOW"] <- 1
disturb_index_cpt["HIGH","HIGH","HIGH","LOW"] <- 1
#table4: urban_prox = LOW, connectivity = MED
disturb_index_cpt["LOW","LOW","LOW","MED"] <- 1
disturb_index_cpt["MED","MED","LOW","MED"] <- 1
disturb_index_cpt["MED","HIGH","LOW","MED"] <- 1
#table5: urban_prox = MED, connectivity = MED
disturb_index_cpt["LOW","LOW","MED","MED"] <- 1
disturb_index_cpt["MED","MED","MED","MED"] <- 1
disturb_index_cpt["HIGH","HIGH","MED","MED"] <- 1
#table6: urban_prox = HIGH, connectivity = MED
disturb_index_cpt["HIGH","LOW","HIGH","MED"] <- 1
disturb_index_cpt["HIGH","MED","HIGH","MED"] <- 1
disturb_index_cpt["HIGH","HIGH","HIGH","MED"] <- 1
#table7: urban_prox = LOW, connectivity = HIGH
disturb_index_cpt["LOW","LOW","LOW","HIGH"] <- 1
disturb_index_cpt["LOW","MED","LOW","HIGH"] <- 1
disturb_index_cpt["MED","HIGH","LOW","HIGH"] <- 1
#table8: urban_prox = MED, connectivity = HIGH
disturb_index_cpt["MED","LOW","MED","HIGH"] <- 1
disturb_index_cpt["MED","MED","MED","HIGH"] <- 1
disturb_index_cpt["HIGH","HIGH","MED","HIGH"] <- 1
#table9: urban_prox = HIGH, connectivity = HIGH
disturb_index_cpt["MED","LOW","HIGH","HIGH"] <- 1
disturb_index_cpt["MED","MED","HIGH","HIGH"] <- 1
disturb_index_cpt["HIGH","HIGH","HIGH","HIGH"] <- 1

disturb_index_cpt

```

```{r}
#CPT for Nutritional Patch Index ####

NPI_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3),  # Shape of the array (3x3)
  dimnames = list(
    NPI = c("LOW", "MED", "HIGH"),
    patch_nutritional_score = c("LOW", "MED", "HIGH")
  )
)

NPI_cpt["LOW","LOW"] <- 1
NPI_cpt["MED","MED"] <- 1
NPI_cpt["HIGH","HIGH"] <- 1

NPI_cpt
```

```{r}
#CPT for Nutritional Landscape Index ####


NLI_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3, 3),  # Shape of the array (3x3)
  dimnames = list(
    NLI = c("LOW", "MED", "HIGH"),
    landscape_nutritional_score = c("LOW", "MED", "HIGH"),
    lf_density = c("LOW","MED","HIGH")
  )
)
#order = landscape_quality_index, nutritional_comp, lf_density
NLI_cpt["LOW","LOW","LOW"] <- 1
NLI_cpt["MED","MED","LOW"] <- 1
NLI_cpt["HIGH","HIGH","LOW"] <- 1
NLI_cpt["LOW","LOW","MED"] <- 1
NLI_cpt["HIGH","MED","MED"] <- 1
NLI_cpt["HIGH","HIGH","MED"] <- 1
NLI_cpt["MED","LOW","HIGH"] <- 1
NLI_cpt["HIGH","MED","HIGH"] <- 1
NLI_cpt["HIGH","HIGH","HIGH"] <- 1

NLI_cpt
```

```{r}
#CPT for Thermoregulation Index ####

thermoreg_cpt <- array(
  0,  # Default probability for each cell
  dim = c(3, 3),  # Shape of the array
  dimnames = list(
    thermoreg_index = c("LOW", "MED", "HIGH"),
    DAMS_MAX = c("LOW", "MED", "HIGH")
  )
)
#order = DAMS_MAX, thermoreg_index
thermoreg_cpt["LOW","LOW"] <- 1
thermoreg_cpt["MED","MED"] <- 1
thermoreg_cpt["HIGH","HIGH"] <- 1

thermoreg_cpt
```

```{r}
#Final CPT: deer damage risk

damage_risk_cpt <- array(
  0,  # Default probability for each cell
  dim = c(5, 3, 3, 3, 3),  # Shape of the array
  dimnames = list(
    damage_risk = c("LOW","LOW-MED","MED","MED-HIGH","HIGH"),
    NPI = c("LOW", "MED", "HIGH"),
    NLI = c("LOW", "MED", "HIGH"),
    thermoreg_index = c("LOW", "MED", "HIGH"),
    disturb_index = c("LOW", "MED", "HIGH")
    
  )
)
#order = damage_risk, patch_quality,landscape_nutrition,thermoreg_index,disturbance
#table1 landscape_nutrition = LOW, thermoreg_index = LOW, disturbance_index = LOW
damage_risk_cpt["LOW-MED","LOW","LOW","LOW","LOW"] <- 1
damage_risk_cpt["LOW-MED","MED","LOW","LOW","LOW"] <- 1
damage_risk_cpt["MED","HIGH","LOW","LOW","LOW"] <- 1
#table2 landscape_nutrition = MED, thermoreg_index = LOW, disturbance_index = LOW
damage_risk_cpt["LOW-MED","LOW","MED","LOW","LOW"] <- 1
damage_risk_cpt["LOW-MED","MED","MED","LOW","LOW"] <- 1
damage_risk_cpt["MED","HIGH","MED","LOW","LOW"] <- 1
#table3 landscape_nutrition = HIGH, thermoreg_index = LOW, disturbance_index = LOW
damage_risk_cpt["LOW-MED","LOW","HIGH","LOW","LOW"] <- 1
damage_risk_cpt["MED","MED","HIGH","LOW","LOW"] <- 1
damage_risk_cpt["MED","HIGH","HIGH","LOW","LOW"] <- 1
#table4 landscape_nutrition = LOW, thermoreg_index = MED, disturbance_index = LOW
damage_risk_cpt["MED","LOW","LOW","MED","LOW"] <- 1
damage_risk_cpt["MED","MED","LOW","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","LOW","MED","LOW"] <- 1
#table5 landscape_nutrition = MED, thermoreg_index = MED, disturbance_index = LOW
damage_risk_cpt["MED","LOW","MED","MED","LOW"] <- 1
damage_risk_cpt["MED","MED","MED","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","MED","MED","LOW"] <- 1
#table6 landscape_nutrition = HIGH, thermoreg_index = MED, disturbance_index = LOW
damage_risk_cpt["MED","LOW","HIGH","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","MED","HIGH","MED","LOW"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","HIGH","MED","LOW"] <- 1
#table7  landscape_nutrition = LOW, thermoreg_index = HIGH, disturbance_index = LOW
damage_risk_cpt["MED-HIGH","LOW","LOW","HIGH","LOW"] <- 1
damage_risk_cpt["MED-HIGH","MED","LOW","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","HIGH","LOW","HIGH","LOW"] <- 1
#table8 landscape_nutrition = MED, thermoreg_index = HIGH, disturbance_index = LOW
damage_risk_cpt["MED-HIGH","LOW","MED","HIGH","LOW"] <- 1
damage_risk_cpt["MED-HIGH","MED","MED","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","HIGH","MED","HIGH","LOW"] <- 1
#table9 landscape_nutrition = HIGH, thermoreg_index = HIGH, disturbance_index = LOW
damage_risk_cpt["MED-HIGH","LOW","HIGH","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","MED","HIGH","HIGH","LOW"] <- 1
damage_risk_cpt["HIGH","HIGH","HIGH","HIGH","LOW"] <- 1
#table10 landscape_nutrition = LOW, thermoreg_index = LOW, disturbance_index = MED
damage_risk_cpt["LOW","LOW","LOW","LOW","MED"] <- 1
damage_risk_cpt["LOW","MED","LOW","LOW","MED"] <- 1
damage_risk_cpt["LOW-MED","HIGH","LOW","LOW","MED"] <- 1
#table11 landscape_nutrition = MED, thermoreg_index = LOW, disturbance_index = MED
damage_risk_cpt["LOW-MED","LOW","MED","LOW","MED"] <- 1
damage_risk_cpt["LOW-MED","MED","MED","LOW","MED"] <- 1
damage_risk_cpt["MED","HIGH","MED","LOW","MED"] <- 1
#table12 landscape_nutrition = HIGH, thermoreg_index = LOW, disturbance_index = MED
damage_risk_cpt["LOW-MED","LOW","HIGH","LOW","MED"] <- 1
damage_risk_cpt["MED","MED","HIGH","LOW","MED"] <- 1
damage_risk_cpt["MED","HIGH","HIGH","LOW","MED"] <- 1
#table13 landscape_nutrition = LOW, thermoreg_index = MED, disturbance_index = MED
damage_risk_cpt["LOW-MED","LOW","LOW","MED","MED"] <- 1
damage_risk_cpt["LOW-MED","MED","LOW","MED","MED"] <- 1
damage_risk_cpt["MED","HIGH","LOW","MED","MED"] <- 1
#table14 landscape_nutrition = MED, thermoreg_index = MED, disturbance_index = MED
damage_risk_cpt["MED","LOW","MED","MED","MED"] <- 1
damage_risk_cpt["MED","MED","MED","MED","MED"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","MED","MED","MED"] <- 1
#table15  landscape_nutrition = HIGH, thermoreg_index = MED, disturbance_index = MED
damage_risk_cpt["MED","LOW","HIGH","MED","MED"] <- 1
damage_risk_cpt["MED-HIGH","MED","HIGH","MED","MED"] <- 1
damage_risk_cpt["MED-HIGH","HIGH","HIGH","MED","MED"] <- 1
#table16 landscape_nutrition = LOW, thermoreg_index = HIGH, disturbance_index = MED
damage_risk_cpt["MED","LOW","LOW","HIGH","MED"] <- 1
damage_risk_cpt["MED","MED","LOW","HIGH","MED"] <- 1
damage_risk_cpt["MED","HIGH","LOW","HIGH","MED"] <- 1
#table17 landscape_nutrition = MED, thermoreg_index = HIGH, disturbance_index = MED
damage_risk_cpt["MED-HIGH","LOW","MED","HIGH","MED"] <- 1
damage_risk_cpt["MED-HIGH","MED","MED","HIGH","MED"] <- 1
damage_risk_cpt["HIGH","HIGH","MED","HIGH","MED"] <- 1
#table18 landscape_nutrition = HIGH, thermoreg_index = HIGH, disturbance_index = MED
damage_risk_cpt["MED-HIGH","LOW","HIGH","HIGH","MED"] <- 1
damage_risk_cpt["HIGH","MED","HIGH","HIGH","MED"] <- 1
damage_risk_cpt["HIGH","HIGH","HIGH","HIGH","MED"] <- 1
#table19 landscape_nutrition = LOW, thermoreg_index = LOW, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","LOW","LOW","HIGH"] <- 1
damage_risk_cpt["LOW","MED","LOW","LOW","HIGH"] <- 1
damage_risk_cpt["LOW","HIGH","LOW","LOW","HIGH"] <- 1
#table20 landscape_nutrition = MED, thermoreg_index = LOW, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","MED","LOW","HIGH"] <- 1
damage_risk_cpt["LOW","MED","MED","LOW","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","MED","LOW","HIGH"] <- 1
#table21 landscape_nutrition = HIGH, thermoreg_index = LOW, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","HIGH","LOW","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","HIGH","LOW","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","HIGH","LOW","HIGH"] <- 1
#table22 landscape_nutrition = LOW, thermoreg_index = MED, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","LOW","MED","HIGH"] <- 1
damage_risk_cpt["LOW","MED","LOW","MED","HIGH"] <- 1
damage_risk_cpt["LOW","HIGH","LOW","MED","HIGH"] <- 1
#table23 landscape_nutrition = MED, thermoreg_index = MED, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","MED","MED","HIGH"] <- 1
damage_risk_cpt["LOW","MED","MED","MED","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","MED","MED","HIGH"] <- 1
#table24 landscape_nutrition = HIGH, thermoreg_index = MED, disturbance_index = HIGH
damage_risk_cpt["LOW","LOW","HIGH","MED","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","HIGH","MED","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","HIGH","MED","HIGH"] <- 1
#table25  landscape_nutrition = LOW, thermoreg_index = HIGH, disturbance_index = HIGH
damage_risk_cpt["LOW-MED","LOW","LOW","HIGH","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","LOW","HIGH","HIGH"] <- 1
damage_risk_cpt["LOW-MED","HIGH","LOW","HIGH","HIGH"] <- 1
#table26 landscape_nutrition = MED, thermoreg_index = HIGH, disturbance_index = HIGH
damage_risk_cpt["LOW-MED","LOW","MED","HIGH","HIGH"] <- 1
damage_risk_cpt["LOW-MED","MED","MED","HIGH","HIGH"] <- 1
damage_risk_cpt["MED","HIGH","MED","HIGH","HIGH"] <- 1
#table27 landscape_nutrition = HIGH, thermoreg_index = HIGH, disturbance_index = HIGH
damage_risk_cpt["LOW-MED","LOW","HIGH","HIGH","HIGH"] <- 1
damage_risk_cpt["MED","MED","HIGH","HIGH","HIGH"] <- 1
damage_risk_cpt["MED","HIGH","HIGH","HIGH","HIGH"] <- 1

damage_risk_cpt
```

#### Create BBN structure

```{r}
# Step 1: Explicitly define the nodes in the network
nodes <- c(
  "disturb_index",
  "urban_prox",
  "road_density",
  "connectivity",
  "patch_nutritional_score",
  "landscape_nutritional_score",
  "NLI",
  "lf_density",
  "woodland_type",
  "thermoreg_index",
  "DAMS_MAX",
  "damage_risk",
  "NPI"
)

#empty graph
e = empty.graph(nodes)

arc.set = matrix(c("urban_prox", "disturb_index",
                     "road_density", "disturb_index", 
                     "connectivity", "disturb_index",
                     "lf_density","landscape_nutrition",
                     "landscape_nutritional_score","NLI",
                     "patch_nutritional_score","NPI",
                     "DAMS_MAX","thermoreg_index",
                     "disturb_index","damage_risk",
                     "NLI","damage_risk",
                     "NPI","damage_risk",
                     "thermoreg_index","damage_risk"),
             ncol = 2, byrow = TRUE,
             dimnames = list(NULL, c("from", "to")))

arcs(e) <- arc.set

model_string <- modelstring(e) 

net<-model2network(model_string) 
# You dont need to repeat the names of the nodes with parents (child nodes) 

# Custom fitting network (matching up the nodes to their CPTs)
dfit = custom.fit(net, dist = list(road_density=road_density_cpt, 
                                   urban_prox=urban_prox_cpt,
                                   connectivity=connectivity_cpt,
                                   lf_density=lf_density_cpt,
                                   DAMS_MAX=dams_cpt,
                                   patch_nutritional_score=patch_nutritional_score_cpt,
                                   landscape_nutritional_score=landscape_nutritional_score_cpt, 
                                   disturb_index=disturb_index_cpt, 
                                   NPI=NPI_cpt,
                                   NLI=NLI_cpt,
                                   thermoreg_index=thermoreg_cpt,
                                   damage_risk=damage_risk_cpt))

```

#### Look at the BN

```{r plot the network}
#| echo: true
#| warning: false
#| message: false

graphviz.plot(net)

```

#### Parameters

```{r look at the dfit}
#| echo: true
#| warning: false
#| message: false
dfit
```

#### Set evidence

```{r set evidence}
#| echo: true
#| warning: false
#| message: false
cpquery(dfit, c(patch_quality == "LOW"|DAMS_MAX == "HIGH"|woodland_type=="BROADLEAVED"), (damage_risk=="LOW"))

```

### Make predictions from BBN

```{r}
predict_dat <- df %>%st_drop_geometry() %>% #drop geometry
   mutate(across(where(is.character), toupper)) %>% #convert to upper
   select(-c("patch_ID","patch_area"))
#Ensure variable names match those in BBN
predict_dat<-predict_dat%>%rename(
                              road_density=all_roads_1000,
                              lf_density=lf_1000,
                              urban_prox=urban_proximity,
                              NLI=nli_edge_types_1000,
                              NPI=npi_edge_types,
                              DAMS_MAX=focal.dams.1000.max,
                              connectivity=incoming_connect_sum)

#Ensure factor levels are in correct order
levels(predict_dat$road_density)<-c("LOW", "MED", "HIGH")
levels(predict_dat$NLI)<-c("LOW", "MED", "HIGH")
levels(predict_dat$NPI)<-c("LOW", "MED", "HIGH")
levels(predict_dat$urban_prox)<-c("LOW", "MED", "HIGH")
levels(predict_dat$road_density)<-c("LOW", "MED", "HIGH")
levels(predict_dat$lf_density)<-c("LOW", "MED", "HIGH")
levels(predict_dat$DAMS_MAX)<-c("LOW", "MED", "HIGH")

#Predict deer damage from data
predict_dat <- predict_dat
pred = predict(dfit,node = "damage_risk",data = predict_dat,method = "bayes-lw")

df$pred<-pred
```

### Draw maps

```{r}
#| echo: true
#| warning: false
#| message: false

plot(predict_dat["pred"])

#Make a plot for each of the 5 selected woodlands to show Matt and Becks
```

### Spare code

```{r}
# Generate random points within each woodland
#This would simulate the user picking a point on a map rather than a polygon
#random_points_list <- lapply(woods$geometry, generate_random_point)  
#random_points <- as.data.frame(random_points_list)
#random_points <- random_points %>%
#  pivot_longer(cols = everything(), values_to = "geometry") #tidy object
#random_points <- random_points %>% dplyr::select(-name)
# Create an sf object with the random points and patch IDs
#wood_metadata <- st_drop_geometry(woods)
#wood_points <- cbind(wood_metadata, random_points)
#wood_points <- st_as_sf(wood_points)
```

```{r}
# Cropped rasters list to store results

#Should we set a centre point for each woodland and crop around that to make a square?
#At the moment it is buffering to the edge of the woodland, making an irregular shaped buffer.

buffered_cropped_rasters <- list()

# Loop through each buffered buffer and crop all rasters
for (i in seq_along(buffered_woods_12km$geometry)) {
  buffer_geometry <- buffered_woods_12km$geometry[i]
  
  # Convert buffer geometry to a spatial object that raster can use
  buffer_extent <- as(extent(st_bbox(buffer_geometry)), "Extent")
  
  # Crop each raster to the current buffer extent
  map_cropped <- crop(map, buffer_extent)
  a_roads_cropped <- crop(a_roads, buffer_extent)
  b_roads_cropped <- crop(b_roads, buffer_extent)
  motorways_cropped <- crop(motorways, buffer_extent)
  nearest_urban_cropped <- crop(nearest_urban, buffer_extent)
  nearest_suburb_cropped <- crop(nearest_suburb, buffer_extent)
  lf_cropped <- crop(lf, buffer_extent)
  dams_cropped <- crop(dams, buffer_extent)
  
  # Store the cropped rasters in the list, using the buffer name as the key
  buffered_cropped_rasters[[paste("buffer", buffered_woods_12km$patch_ID[i])]] <- list(
    map = map_cropped,
    a_roads = a_roads_cropped,
    b_roads = b_roads_cropped,
    motorways = motorways_cropped,
    nearest_urban = nearest_urban_cropped,
    nearest_suburb = nearest_suburb_cropped,
    lf = lf_cropped,
    dams = dams_cropped
  )
}

plot(buffered_cropped_rasters[[1]]$map)
plot(buffered_woods_12km$geometry[1,], add=TRUE, border="blue", lwd=2, col=NA)
```
